### 2021-08-05 Seldin Report.Rmd
Initial- transfer to work computer

---
title: "Seldin Report: Ssec Scores"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Match top Ssec values to gene symbols
  c. Figure out threshold/significance test
  d. Functional Enrichment

### 2021-08-05 Tissue Specific Report.Rmd
Initial- transfer to work computer

---
title: "Report for Tissue-Specific Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report results of tissue-specific proteins data

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")

############## Calculate R squared for tissue specific and not tissue specific########################
fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")
fin.r.squared_liver_sec = read.csv("R_Squared_Liver_Sec.csv")
fin.r.squared_brain_notsec = read.csv("R_Squared_Brain_NotSec.csv")
fin.r.squared_liver_notsec = read.csv("R_Squared_Liver_NotSec.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
ts_protein_brain = fix.read.csv(ts_protein_brain)

protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
ts_protein_liver = fix.read.csv(ts_protein_liver)

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)
fin.r.squared_liver_sec = fix.read.csv(fin.r.squared_liver_sec)
fin.r.squared_brain_notsec = fix.read.csv(fin.r.squared_brain_notsec)
fin.r.squared_liver_notsec = fix.read.csv(fin.r.squared_liver_notsec)

```

```{r, echo = FALSE, message = FALSE}
# table comparing number of genes
ngenes = ncol(protein_brain)
ngenes_sec = ncol(secreted_protein_brain)
ngenes_ts = ncol(ts_protein_brain)

ngenes.l = ncol(protein_liver)
ngenes_sec.l = ncol(secreted_protein_liver)
ngenes_ts.l = ncol(ts_protein_liver)
tot_sec = 1760 # from secreted uniprot data
tot_ts = "WHAT"

num = c(ngenes, ngenes_sec, ngenes_ts, ngenes.l, ngenes_sec.l, ngenes_ts.l, tot_sec, tot_ts)

names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Brain: Tissue-Specific Data", "Liver: Full Data", "Liver: Secreted Data", "Liver: Tissue-Specific Data", "Total Secreted Rat Proteins", "Total Tissue-Specific Rat Proteins")

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

# compare shared genes between secreted brain and secreted liver

gee = intersect(colnames(ts_protein_brain), colnames(ts_protein_liver))

int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
names(int) = c("Tissue-Specific Proteins Brain", "Tissue-Specific Proteins Shared by Brain and Liver", "Tissue-Specific Proteins Liver")

knitr::kable(int)

## THis isn't right...

```

### 2021-08-05 Tissue Specific Report.Rmd
Code finished - needs writing

@@
 secreted_protein_brain = read.csv("Secreted_Brain.csv")
-ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
+#ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
 protein_liver = read.csv("Liver_WGCNA.csv")
 secreted_protein_liver = read.csv("Secreted_Liver.csv")
-ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
+#ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
 
-############## Calculate R squared for tissue specific and not tissue specific########################
 fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
-fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")
 fin.r.squared_liver = read.csv("R_Squared_Liver.csv")
-fin.r.squared_liver_sec = read.csv("R_Squared_Liver_Sec.csv")
-fin.r.squared_brain_notsec = read.csv("R_Squared_Brain_NotSec.csv")
-fin.r.squared_liver_notsec = read.csv("R_Squared_Liver_NotSec.csv")
 
@@
 secreted_protein_brain = fix.read.csv(secreted_protein_brain)
-ts_protein_brain = fix.read.csv(ts_protein_brain)
+#ts_protein_brain = fix.read.csv(ts_protein_brain)
 
@@
 secreted_protein_liver = fix.read.csv(secreted_protein_liver)
-ts_protein_liver = fix.read.csv(ts_protein_liver)
+#ts_protein_liver = fix.read.csv(ts_protein_liver)
 
 fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
-fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
 fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)
-fin.r.squared_liver_sec = fix.read.csv(fin.r.squared_liver_sec)
-fin.r.squared_brain_notsec = fix.read.csv(fin.r.squared_brain_notsec)
-fin.r.squared_liver_notsec = fix.read.csv(fin.r.squared_liver_notsec)
+
+```
+
+```{r}
+# looking for unique genes
+
+genes_brain = colnames(secreted_protein_brain)
+genes_liver = colnames(secreted_protein_liver)
+
+gee = genes_brain %in% genes_liver
+gee = gee == FALSE
+only_brain = genes_brain[gee]
+
+wee = genes_liver %in% genes_brain
+wee = wee == FALSE
+only_liver = genes_liver[wee]
+
+ts_protein_brain = secreted_protein_brain[, colnames(secreted_protein_brain) %in% only_brain]
+ts_protein_liver = secreted_protein_liver[, colnames(secreted_protein_liver) %in% only_liver]
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(ts_protein_brain, "TissueSpecific2_Protein_Brain.csv")
+write.csv(ts_protein_liver, "TissueSpecific2_Protein_Liver.csv")
 
@@
 tot_sec = 1760 # from secreted uniprot data
-tot_ts = "WHAT"
+tot_ts = ngenes_ts + ngenes_ts.l
 
@@
 
-## THis isn't right...
+```
+
+```{r}
+# R squared
+
+rsq.brain = sub(" ~ strain_factor_brain.*", "", fin.r.squared_brain$names_brain)
+rsq.liver = sub(" ~ strain_factor_liver.*", "", fin.r.squared_liver$names_liver)
+
+fin.r.squared_brain[,1] = rsq.brain
+fin.r.squared_liver[,1] = rsq.liver
+
+r.squared_brain_ts = fin.r.squared_brain[fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain),]
+r.squared_liver_ts = fin.r.squared_liver[fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver),]
+
+r.squared_brain_notts = fin.r.squared_brain[!(fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain)),]
+r.squared_liver_notts = fin.r.squared_liver[!(fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver)),]
+
+#brain
+
+gee = cbind(rep("not ts", times = dim(r.squared_brain_notts)[1]), r.squared_brain_notts[,2])
+gee2 = cbind(rep("ts", times = dim(r.squared_brain_ts)[1]), r.squared_brain_ts[,2])
+
+wil.b = rbind(gee, gee2)
+wil.b = as.data.frame(wil.b)
+colnames(wil.b) = c("trt", "r.sq")
+
+wil.b$trt = as.factor(wil.b$trt)
+wil.b$r.sq = as.numeric(wil.b$r.sq)
+
+wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b)
+
+# liver
+
+gee = cbind(rep("not ts", times = dim(r.squared_liver_notts)[1]), r.squared_liver_notts[,2])
+gee2 = cbind(rep("ts", times = dim(r.squared_liver_ts)[1]), r.squared_liver_ts[,2])
+
+wil.l = rbind(gee, gee2)
+wil.l = as.data.frame(wil.l)
+colnames(wil.l) = c("trt", "r.sq")
+
+wil.l$trt = as.factor(wil.l$trt)
+wil.l$r.sq = as.numeric(wil.l$r.sq)
+
+wil.liver = coin::wilcox_test(r.sq ~ trt, wil.l)
 
 ```
+
+```{r}
+
+# Density R Squared
+
+ggplot(wil.b, aes(x = r.sq, color = trt)) + 
+  geom_density() +
+  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")
+
+ggplot(wil.l, aes(x = r.sq, color = trt)) + 
+  geom_density() +
+  labs(title="R Squared Density Curve: Liver",x="R Squared", y = "Density")
+
+# Box R Squared
+
+boxplot(fin.r.squared_brain[,2], fin.r.squared_brain_sec[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
+boxplot(fin.r.squared_liver[,2], fin.r.squared_liver_sec[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
+
+```

### 2021-08-07 Seldin Report.Rmd
Permuted wrong

@@
   d. Functional Enrichment
+
+* Using exclusive ts proteins unless we decide later to change back to Uniprot TS
+
+```{r, echo = FALSE, message = FALSE}
+
+library(tidyverse)
+library(reshape2)
+library(WGCNA)
+#library(reshape)
+#library(ggplot2)
+
+# Data Input
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+protein_liver = read.csv("Liver_WGCNA.csv")
+protein_brain = read.csv("Brain_WGCNA.csv")
+secreted_protein_brain = read.csv("Secreted_Brain.csv")
+secreted_protein_liver = read.csv("Secreted_Liver.csv")
+sec_annot = read.csv("Sec_Annotation.csv")
+tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
+tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
+ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
+ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
+
+fix.read.csv = function(x)
+{
+  names = x[,1]
+  gee = x[,2:dim(x)[2]]
+  rownames(gee) = names
+  return(gee)
+}
+
+protein_liver = fix.read.csv(protein_liver)
+protein_brain = fix.read.csv(protein_brain)
+secreted_protein_brain = fix.read.csv(secreted_protein_brain)
+secreted_protein_liver = fix.read.csv(secreted_protein_liver)
+sec_annot = fix.read.csv(sec_annot)
+tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
+tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
+ts_protein_brain = fix.read.csv(ts_protein_brain)
+ts_protein_liver = fix.read.csv(ts_protein_liver)
+
+```
+Calculating Ssec values
+```{r, echo = FALSE, message = FALSE}
+
+#### BRAIN ####
+
+brain.liv = bicorAndPvalue(secreted_protein_brain, protein_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+## tissue-specific genes
+
+ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
+top_brain = head(ts_Ssec_brain[,1])
+
+#### LIVER ####
+
+liv.brain = bicorAndPvalue(secreted_protein_liver, protein_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain$p))
+
+Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+## tissue-specific genes
+
+ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
+top_liver = head(ts_Ssec_liver[,1])
+bot_liver = tail(ts_Ssec_liver[,1])
+
+```
+
+PROCEED WITH CAUTION
+```{r}
+####################################
+# Calc Ssec for all proteins# Don't do this !!
+
+brain.liv = bicorAndPvalue(protein_brain, protein_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+## tissue-specific genes
+
+top_brain = head(Ssec_brain[,1])
+```
+
+```{r}
+# Visualizations
+hist(Ssec_brain$Ssec)
+hist(ts_Ssec_brain$Ssec)
+
+h1 = cbind(rep("brain", times = length(Ssec_brain$Ssec)), Ssec_brain$Ssec)
+h2 = cbind(rep("TS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")
+
+plotting = melt(liv.brain$p, id.vars = "refseq")
+limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Proteins", x = "P-Value", y = "Density")
+
+```
+NEXT: permute null distribution of Ssec; Calculate Ssec with cutoff; maybe figure out how to run Ssec on non-secreted proteins; null distribution of p-values?
+
+We permuted the target tissue strains and recalculated Ssec to
+construct the null distribution and maintain the gene-gene correlation structures in each tissue. These tests typically required
+1,000–1,000,000 permutations, listed as p values in Table S1.
+The permutation tests were designed to specifically to allow generation of a statistic while maintaining correlation structure to
+limit the influence of inflation of cross-tissue gene correlations
+(discussed below). We note that in these tests the p values
+used reflect a placeholder for the correlation coefficient, which
+does not require independence.
+
+```{r}
+### This is no good
+
+brain.liv_p = brain.liv$p
+
+brain.liv_perm = apply(brain.liv_p, 2, sample, size = dim(brain.liv_p)[1], replace = F)
+rownames(brain.liv_perm) = as.character(1:dim(brain.liv_perm)[1])
+
+# calc Ssec
+
+scores = rowSums(-log(brain.liv_perm))
+
+Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+hist(Ssec_brain$Ssec)
+hist(Ssec_brain_perm$Ssec)
+
+```
+```{r}
+
+#permute strain names in target tissue array
+# bicor> pval> Ssec val
+# repeat 1000- compare observed value to mean of permuted values
+
+```

### 2021-08-09 Tissue Specific Report.Rmd
final TS report

@@
 1. Report results of tissue-specific proteins data
+
+The ensembl rat liver and brain RNAseq data was downloaded from BioMart database (accessed 2021).
+Inbred strains were excluded, leaving only recombinant inbred strains.
+Enrichment data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set.
+This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.
+
+From there, the data was reduced to secreted proteins, then to tissue-specific secreted proteins.
+Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021).
+The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets.
+From the secreted proteins datasets, we then determined the proteins in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset).
+This resulted in two tissue-specific secreted protein datasets, with 174 in the brain dataset, and 35 in the liver dataset.
+
+The following report was completed using `r R.Version()$version.string`.
+Additionally, the packages WGCNA (version 1.70.3), broom (version 0.7.7), tidyverse (version `r packageVersion("tidyverse")`), and knitr (version `r packageVersion("knitr")`).
 
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 # looking for unique genes
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 # R squared
@@
 
-```{r}
+The R-squared density curves for the brain data below show similar peaks.
+The curves for the liver data indicate a greater right skew in the tissue-specific proteins.
+For both sets of data, the distinction is not as clear as the secreted proteins dataset, likely due to small sample size.
+
+```{r, echo = FALSE, message = FALSE}
 
@@
 
-boxplot(fin.r.squared_brain[,2], fin.r.squared_brain_sec[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
-boxplot(fin.r.squared_liver[,2], fin.r.squared_liver_sec[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
+boxplot(fin.r.squared_brain[,2], r.squared_brain_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
+boxplot(fin.r.squared_liver[,2], r.squared_liver_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
 
 ```
+
+We tested whether the heritability of the non-tissue-specific and tissue specific datasets were significantly different using the Wilcoxon-Mann-Whitney Test.
+Both tests showed a significant difference at alpha = 0.05, however the p-value for the brain data was very close to 0.05.
+
+Non-Tissue-Specific Brain Dataset v. Tissue-Specific Brain Dataset:
+
+```{r, echo = FALSE, message = FALSE}
+wil.brain
+
+```
+
+Non-Tissue-Specific Liver Dataset v. Tissue-Specific Liver Dataset:
+
+```{r, echo = FALSE, message = FALSE}
+
+wil.liver
+
+```

### 2021-08-09 Seldin Report.Rmd
rough

@@
   a. Find distribution of Ssec values
-  b. Match top Ssec values to gene symbols
-  c. Figure out threshold/significance test
-  d. Functional Enrichment
+  b. Find distribution of correlation p-values
 
@@
 top_brain = head(ts_Ssec_brain[,1])
+bot_brain = tail(ts_Ssec_brain[,1])
 
@@
 PROCEED WITH CAUTION
-```{r}
+```{r, eval = FALSE}
 ####################################
@@
 # Visualizations
-hist(Ssec_brain$Ssec)
-hist(ts_Ssec_brain$Ssec)
+#hist(Ssec_brain$Ssec)
+#hist(ts_Ssec_brain$Ssec)
 
+#Brain
 h1 = cbind(rep("brain", times = length(Ssec_brain$Ssec)), Ssec_brain$Ssec)
@@
   geom_density() +
-  labs(title="Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")
+  labs(title="Brain Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")
 
+# Liver
+h1 = cbind(rep("liver", times = length(Ssec_liver$Ssec)), Ssec_liver$Ssec)
+h2 = cbind(rep("TS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")
+
+```
+
+The above graphs show the density distribution of Ssec scores of the secreted proteins and the tissue-specific proteins (a subset of secreted proteins).
+The shape of the distributions are not very different, indicating that tissue-specificity does not make it more likely that there will be a significant relationship with target-tissue proteins.
+Consider the relationship between the non-tissue-specific proteins and the target tissue expression.
+
+```{r}
+##
+# brain
+plotting = melt(brain.liv$p, id.vars = "refseq")
+limit = plotting[as.character(plotting$Var1) %in% c(top_brain, bot_brain),]
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Proteins", x = "P-Value", y = "Density")
+
+# liver
 plotting = melt(liv.brain$p, id.vars = "refseq")
@@
 ```
+
+The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores.
+We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores.
+Are there a high number of low but not significantly low p-values that could inflate these?
+Are we discounting some relationships because there aren't a lot of connections, but a small number of strong connections?
+The log transformation of the pvalues should weigh the lowest p-values much higher.
+Don't want to entirely discount pvalues just under a certain threshold.
+
+Secreted v TS- don't change shape as you go from secreted to tissue specific.
+Does that mean tissue specific not more likely to have a relationship with target tissue?
+See if correlation with proteins across tissues is more significant than the same correlation with the same proteins within the tissue....
+Would secreted proteins have an effect on its own tissue?
+Perhaps some kind of relationship
+Pval. Do we care at all about the magnitude of the correlation? I know we're not interested in direction.
+
+```{r}
+# brain
+magn = brain.liv$bicor
+
+plotting = melt(magn, id.vars = "refseq")
+limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Proteins", x = "Correlation", y = "Density")
+
+# liver
+magn = liv.brain$bicor
+plotting = melt(magn, id.vars = "refseq")
+limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Proteins", x = "Correlation", y = "Density")
+
+```
+
+Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation.
+The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients.
+
 NEXT: permute null distribution of Ssec; Calculate Ssec with cutoff; maybe figure out how to run Ssec on non-secreted proteins; null distribution of p-values?
@@
 
-```{r}
-### This is no good
+```{r, eval = FALSE}
 
-brain.liv_p = brain.liv$p
+#permute strain names in target tissue array
+# bicor> pval> Ssec val
+# repeat 1000- compare observed value to mean of permuted values
+target_protein = protein_liver
+perm_test = matrix(NA, ncol = 1000, nrow = 174)
+perm_test = as.data.frame(perm_test)
+rownames(perm_test) = ts_Ssec_brain[,1]
 
-brain.liv_perm = apply(brain.liv_p, 2, sample, size = dim(brain.liv_p)[1], replace = F)
-rownames(brain.liv_perm) = as.character(1:dim(brain.liv_perm)[1])
+for (i in 1:1000)
+{
 
-# calc Ssec
+perm = sample(rownames(target_protein), size = length(rownames(target_protein)), replace = FALSE)
+target_protein = target_protein[perm,]
 
-scores = rowSums(-log(brain.liv_perm))
+brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')
+
+scores = rowSums(-log(brain.liv$p))
 
@@
 
-hist(Ssec_brain$Ssec)
-hist(Ssec_brain_perm$Ssec)
+perm_test[,i] = Ssec_brain_perm[,2]
+}
+
+## i dont get how this works...
+
+pee = perm_test[rownames(perm_test) == top_brain[2],]
+hist(as.numeric(pee))
+Ssec_brain[Ssec_brain$Gene_symbol == top_brain[2],]
+# ok all of the true values are inside the null distribution. Did i do something wrong??
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(perm_test, "Permutation_Seldin_Report.csv")
 
 ```
-```{r}
-
-#permute strain names in target tissue array
-# bicor> pval> Ssec val
-# repeat 1000- compare observed value to mean of permuted values
-
-```

### 2021-08-11 Seldin Report.Rmd
incorporating tsns

--- 2021-08-05 Seldin Report.Rmd
@@
   a. Find distribution of Ssec values
-  b. Match top Ssec values to gene symbols
-  c. Figure out threshold/significance test
-  d. Functional Enrichment
+  b. Find distribution of correlation p-values
+  c. Find a null distribution of Ssec values
+
+```{r, echo = FALSE, message = FALSE}
+
+library(tidyverse)
+library(reshape2)
+library(WGCNA)
+#library(reshape)
+#library(ggplot2)
+
+# Data Input
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+protein_liver = read.csv("Liver_WGCNA.csv")
+protein_brain = read.csv("Brain_WGCNA.csv")
+secreted_protein_brain = read.csv("Secreted_Brain.csv")
+secreted_protein_liver = read.csv("Secreted_Liver.csv")
+sec_annot = read.csv("Sec_Annotation.csv")
+tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
+tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
+ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
+ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
+tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
+tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")
+
+fix.read.csv = function(x)
+{
+  names = x[,1]
+  gee = x[,2:dim(x)[2]]
+  rownames(gee) = names
+  return(gee)
+}
+
+protein_liver = fix.read.csv(protein_liver)
+protein_brain = fix.read.csv(protein_brain)
+secreted_protein_brain = fix.read.csv(secreted_protein_brain)
+secreted_protein_liver = fix.read.csv(secreted_protein_liver)
+sec_annot = fix.read.csv(sec_annot)
+tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
+tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
+ts_protein_brain = fix.read.csv(ts_protein_brain)
+ts_protein_liver = fix.read.csv(ts_protein_liver)
+tsns_protein_brain = fix.read.csv(tsns_protein_brain)
+tsns_protein_liver = fix.read.csv(tsns_protein_liver)
+
+```
+Tissue-Specific Secreted > all target proteins
+Secreted > all target proteins
+Tissue-Specific Secreted > Tissue-Specific Non-Secreted
+Tissue-Specific Non-Secreted > all target proteins
+Tissue-Specific Non-Secreted > Tissue-Specific Non-Secreted
+TRYING all origin proteins > all target proteins- SERVER
+Maybe try doing tissue specific non-secreted and tissue specific secreted as target
+
+```{r, echo = FALSE, message = FALSE}
+# Calculate Ssec values: tissue-specific -> ns tissue-specific
+#### TS ####
+## BRAIN ##
+
+brain.liv = bicorAndPvalue(ts_protein_brain, tsns_protein_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+ts_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(tsns_protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
+top_brain = head(ts_Ssec_brain[,1])
+bot_brain = tail(ts_Ssec_brain[,1])
+
+## LIVER ##
+
+liv.brain = bicorAndPvalue(ts_protein_liver, tsns_protein_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain$p))
+
+ts_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(tsns_protein_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
+top_liver = head(ts_Ssec_liver[,1])
+bot_liver = tail(ts_Ssec_liver[,1])
+
+#### TSNS ####
+# ns tissue-specific -> ns tissue-specific
+## BRAIN ##
+
+brain.liv = bicorAndPvalue(tsns_protein_brain, tsns_protein_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+tsns_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(tsns_protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
+#top_brain = head(tsns_Ssec_brain[,1])
+#bot_brain = tail(ts_Ssec_brain[,1])
+
+## LIVER ##
+
+liv.brain = bicorAndPvalue(tsns_protein_liver, tsns_protein_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain$p))
+
+tsns_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(tsns_protein_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
+#top_liver = head(ts_Ssec_liver[,1])
+#bot_liver = tail(ts_Ssec_liver[,1])
+
+```
+
+```{r}
+# this isn't working great- probably server
+ints = c(1, 1500, 1501, 3000, 3001, 4500, 4501, 6000, 6001, 7500, 7501, 9000, 9001, 10500, 10501, 12000, 12001, 13500, 13501, 15000, 15001, 16457)
+
+bee = matrix(data = NA, nrow = 0, ncol = 2)
+bee = as.data.frame(bee)
+count = 1
+
+for (i in 1:11)
+{
+  interval = protein_brain[, ints[count]:ints[count +1]]
+brain.liv = bicorAndPvalue(interval, protein_liver, use = 'pairwise.complete.obs')
+scores = rowSums(-log(liv.brain$p))
+
+full_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+bee = rbind(bee, full_Ssec_brain)
+count = count + 2
+}
+
+```
+
+```{r, echo = FALSE}
+# Visualizations
+#hist(Ssec_brain$Ssec)
+#hist(ts_Ssec_brain$Ssec)
+
+# Ssec distribtions TSNS v. TS, brain and liver
+#Brain
+h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
+h2 = cbind(rep("TS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Brain Ssec Distribution: TSNS Proteins v. TS Proteins", x = "Ssec", y = "Density")
+
+# Liver
+h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
+h2 = cbind(rep("TS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: TSNS Proteins v. TS Proteins", x = "Ssec", y = "Density")
+
+```
+
+# interpretation
+
+```{r, echo = FALSE}
+## Pvalue for top and bottom Ssec scores, liver and brain
+# brain
+plotting = melt(brain.liv$p, id.vars = "refseq")
+limit = plotting[as.character(plotting$Var1) %in% c(top_brain, bot_brain),]
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Proteins", x = "P-Value", y = "Density")
+
+# liver
+plotting = melt(liv.brain$p, id.vars = "refseq")
+limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Proteins", x = "P-Value", y = "Density")
+
+```
+
+The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores.
+We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores.
+Are there a high number of low but not significantly low p-values that could inflate these?
+Are we discounting some relationships because there aren't a lot of connections, but a small number of strong connections?
+The log transformation of the pvalues should weigh the lowest p-values much higher.
+Don't want to entirely discount pvalues just under a certain threshold.
+
+```{r, echo = FALSE}
+# Correlation coefficients top and bottom Ssec scores, liver and brain
+# brain
+magn = brain.liv$bicor
+
+plotting = melt(magn, id.vars = "refseq")
+limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Proteins", x = "Correlation", y = "Density")
+
+# liver
+magn = liv.brain$bicor
+plotting = melt(magn, id.vars = "refseq")
+limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])
+
+ggplot(limit, aes(value, color = Var1)) + 
+  geom_density() +
+  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Proteins", x = "Correlation", y = "Density")
+
+```
+
+Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation.
+The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients.
+
+PERMUTATION
+
+We permuted the target tissue strains and recalculated Ssec to
+construct the null distribution and maintain the gene-gene correlation structures in each tissue. These tests typically required
+1,000–1,000,000 permutations, listed as p values in Table S1.
+The permutation tests were designed to specifically to allow generation of a statistic while maintaining correlation structure to
+limit the influence of inflation of cross-tissue gene correlations
+(discussed below). We note that in these tests the p values
+used reflect a placeholder for the correlation coefficient, which
+does not require independence.
+
+```{r, eval = FALSE}
+
+#permute strain names in target tissue array
+# bicor> pval> Ssec val
+# repeat 1000- compare observed value to mean of permuted values
+target_protein = protein_liver
+perm_test = matrix(NA, ncol = 1000, nrow = 174)
+perm_test = as.data.frame(perm_test)
+rownames(perm_test) = ts_Ssec_brain[,1]
+
+for (i in 1:1000)
+{
+
+perm = sample(rownames(target_protein), size = length(rownames(target_protein)), replace = FALSE)
+target_protein = target_protein[perm,]
+
+brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')
+
+scores = rowSums(-log(brain.liv$p))
+
+Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+perm_test[,i] = Ssec_brain_perm[,2]
+}
+
+## i dont get how this works...
+
+pee = perm_test[rownames(perm_test) == top_brain[2],]
+hist(as.numeric(pee))
+Ssec_brain[Ssec_brain$Gene_symbol == top_brain[2],]
+# ok all of the true values are inside the null distribution. Did i do something wrong??
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(perm_test, "Permutation_Seldin_Report.csv")
+
+```

### 2021-08-11 Tissue Specific Report.Rmd
incorporating tsns

@@
 write.csv(ts_protein_liver, "TissueSpecific2_Protein_Liver.csv")
+
+## Tissue Specific Non-Secreted
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+notsec_brain = read.csv("NotSecreted_Brain.csv")
+notsec_liver = read.csv("NotSecreted_Liver.csv")
+notsec_liver = fix.read.csv(notsec_liver)
+notsec_brain = fix.read.csv(notsec_brain)
+
+genes_brain = colnames(notsec_brain)
+genes_liver = colnames(notsec_liver)
+
+gee = genes_brain %in% genes_liver
+gee = gee == FALSE
+only_brain = genes_brain[gee]
+
+wee = genes_liver %in% genes_brain
+wee = wee == FALSE
+only_liver = genes_liver[wee]
+
+tsns_protein_brain = notsec_brain[, colnames(notsec_brain) %in% only_brain]
+tsns_protein_liver = notsec_liver[, colnames(notsec_liver) %in% only_liver]
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(tsns_protein_brain, "NSTissueSpecific_Protein_Brain.csv")
+write.csv(tsns_protein_liver, "NSTissueSpecific_Protein_Liver.csv")
+
+#sum(colnames(ts_protein_liver) %in% colnames(tsns_protein_liver))
 

### 2021-08-12 Tissue Specific Report.Rmd
comparing to full dataset instead of non-secreted tissue-specific

--- 2021-08-09 Tissue Specific Report.Rmd
@@
 
+## Tissue Specific Non-Secreted
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#notsec_brain = read.csv("NotSecreted_Brain.csv")
+#notsec_liver = read.csv("NotSecreted_Liver.csv")
+#notsec_liver = fix.read.csv(notsec_liver)
+#notsec_brain = fix.read.csv(notsec_brain)
+
+#genes_brain = colnames(notsec_brain)
+#genes_liver = colnames(notsec_liver)
+
+#gee = genes_brain %in% genes_liver
+#gee = gee == FALSE
+#only_brain = genes_brain[gee]
+
+#wee = genes_liver %in% genes_brain
+#wee = wee == FALSE
+#only_liver = genes_liver[wee]
+
+#tsns_protein_brain = notsec_brain[, colnames(notsec_brain) %in% only_brain]
+#tsns_protein_liver = notsec_liver[, colnames(notsec_liver) %in% only_liver]
+
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(tsns_protein_brain, "NSTissueSpecific_Protein_Brain.csv")
+#write.csv(tsns_protein_liver, "NSTissueSpecific_Protein_Liver.csv")
+
+#sum(colnames(ts_protein_liver) %in% colnames(tsns_protein_liver))
+
 ```
@@
 
-r.squared_brain_notts = fin.r.squared_brain[!(fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain)),]
-r.squared_liver_notts = fin.r.squared_liver[!(fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver)),]
+#r.squared_brain_notts = fin.r.squared_brain[!(fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain)),]
+#r.squared_liver_notts = fin.r.squared_liver[!(fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver)),]
 
@@
 
-gee = cbind(rep("not ts", times = dim(r.squared_brain_notts)[1]), r.squared_brain_notts[,2])
+gee = cbind(rep("full", times = dim(fin.r.squared_brain)[1]), fin.r.squared_brain[,2])
 gee2 = cbind(rep("ts", times = dim(r.squared_brain_ts)[1]), r.squared_brain_ts[,2])
@@
 
-gee = cbind(rep("not ts", times = dim(r.squared_liver_notts)[1]), r.squared_liver_notts[,2])
+gee = cbind(rep("full", times = dim(fin.r.squared_liver)[1]), fin.r.squared_liver[,2])
 gee2 = cbind(rep("ts", times = dim(r.squared_liver_ts)[1]), r.squared_liver_ts[,2])
@@
 
-boxplot(fin.r.squared_brain[,2], r.squared_brain_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
-boxplot(fin.r.squared_liver[,2], r.squared_liver_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
+boxplot(fin.r.squared_brain[,2], r.squared_brain_ts[,2], names = c("Full", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
+boxplot(fin.r.squared_liver[,2], r.squared_liver_ts[,2], names = c("Full", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
 

### 2021-08-27 Tissue Specific Report.Rmd
fixed a lot of problems with TS report

--- 2021-08-09 Tissue Specific Report.Rmd
@@
 
-The ensembl rat liver and brain RNAseq data was downloaded from BioMart database (accessed 2021).
+The ensembl rat liver and brain RNAseq data was acquired (get more info from papers) downloaded from PhenoGen.
 Inbred strains were excluded, leaving only recombinant inbred strains.
-Enrichment data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set.
+Annontation data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set.
 This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.
 
-From there, the data was reduced to secreted proteins, then to tissue-specific secreted proteins.
+From there, our datasets were segregated by terms secreted and tissue-specific.
 Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021).
 The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets.
-From the secreted proteins datasets, we then determined the proteins in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset).
-This resulted in two tissue-specific secreted protein datasets, with 174 in the brain dataset, and 35 in the liver dataset.
+Tissue-specific status was determined by finding the genes in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset).
+From this we created our dataset of investigation: tissue-specific secreted gene datasets for liver (with 35 genes) and brain (with 174 genes) tissues.
+This dataset was compared to a comparable dataset of tissue-specific non-secreted genes, with _ liver genes and _ brain genes.
 
@@
 # Data Input
-#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 protein_brain = read.csv("Brain_WGCNA.csv")
 secreted_protein_brain = read.csv("Secreted_Brain.csv")
-#ts_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
+
 protein_liver = read.csv("Liver_WGCNA.csv")
 secreted_protein_liver = read.csv("Secreted_Liver.csv")
-#ts_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
 
@@
 secreted_protein_brain = fix.read.csv(secreted_protein_brain)
-#ts_protein_brain = fix.read.csv(ts_protein_brain)
 
@@
 secreted_protein_liver = fix.read.csv(secreted_protein_liver)
-#ts_protein_liver = fix.read.csv(ts_protein_liver)
 
@@
 
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
-write.csv(ts_protein_brain, "TissueSpecific2_Protein_Brain.csv")
-write.csv(ts_protein_liver, "TissueSpecific2_Protein_Liver.csv")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(ts_protein_brain, "TissueSpecific2_Protein_Brain.csv")
+#write.csv(ts_protein_liver, "TissueSpecific2_Protein_Liver.csv")
+
+## Tissue Specific Non-Secreted
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+notsec_brain = read.csv("NotSecreted_Brain.csv")
+notsec_liver = read.csv("NotSecreted_Liver.csv")
+notsec_liver = fix.read.csv(notsec_liver)
+notsec_brain = fix.read.csv(notsec_brain)
+
+genes_brain = colnames(notsec_brain)
+genes_liver = colnames(notsec_liver)
+
+gee = genes_brain %in% genes_liver
+gee = gee == FALSE
+only_brain = genes_brain[gee]
+
+wee = genes_liver %in% genes_brain
+wee = wee == FALSE
+only_liver = genes_liver[wee]
+
+tsns_protein_brain = notsec_brain[, colnames(notsec_brain) %in% only_brain]
+tsns_protein_liver = notsec_liver[, colnames(notsec_liver) %in% only_liver]
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(tsns_protein_brain, "NSTissueSpecific_Protein_Brain.csv")
+#write.csv(tsns_protein_liver, "NSTissueSpecific_Protein_Liver.csv")
+
+#sum(colnames(ts_protein_liver) %in% colnames(tsns_protein_liver))
 
@@
 ngenes = ncol(protein_brain)
-ngenes_sec = ncol(secreted_protein_brain)
+ngenes_tsns = ncol(tsns_protein_brain)
 ngenes_ts = ncol(ts_protein_brain)
@@
 ngenes.l = ncol(protein_liver)
-ngenes_sec.l = ncol(secreted_protein_liver)
+ngenes_tsns.l = ncol(tsns_protein_liver)
 ngenes_ts.l = ncol(ts_protein_liver)
-tot_sec = 1760 # from secreted uniprot data
+tot_tsns = ngenes_tsns + ngenes_tsns.l # from secreted uniprot data
 tot_ts = ngenes_ts + ngenes_ts.l
 
-num = c(ngenes, ngenes_sec, ngenes_ts, ngenes.l, ngenes_sec.l, ngenes_ts.l, tot_sec, tot_ts)
+num = c(ngenes, ngenes_tsns, ngenes_ts, ngenes.l, ngenes_tsns.l, ngenes_ts.l, tot_tsns, tot_ts)
 
-names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Brain: Tissue-Specific Data", "Liver: Full Data", "Liver: Secreted Data", "Liver: Tissue-Specific Data", "Total Secreted Rat Proteins", "Total Tissue-Specific Rat Proteins")
+names(num) = c("Brain: Full Data", "Brain: Tissue-Specific Non-Secreted Data", "Brain: Tissue-Specific Secreted Data", "Liver: Full Data", "Liver: Tissue-Specific Non-Secreted Data", "Liver: Tissue-Specific Secreted Data", "Total Tissue-Specific Non-Secreted Rat Genes", "Total Tissue-Specific Secreted Rat Genes")
 
@@
 
-# compare shared genes between secreted brain and secreted liver
+# compare shared genes between brain and liver
 
@@
 int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
-names(int) = c("Tissue-Specific Proteins Brain", "Tissue-Specific Proteins Shared by Brain and Liver", "Tissue-Specific Proteins Liver")
+names(int) = c("Tissue-Specific Secreted Genes Brain", "Tissue-Specific Secreted Genes Shared by Brain and Liver", "Tissue-Specific Secreted Genes Liver")
 
 knitr::kable(int)
+
+hee = intersect(colnames(tsns_protein_brain), colnames(tsns_protein_liver))
+
+int2 = c(length(colnames(tsns_protein_brain)), length(hee), length(colnames(tsns_protein_liver)))
+names(int2) = c("Tissue-Specific Non-Secreted Genes Brain", "Tissue-Specific Non-Secreted Genes Shared by Brain and Liver", "Tissue-Specific Non-Secreted Genes Liver")
+
+knitr::kable(int2)
 
@@
 
-r.squared_brain_notts = fin.r.squared_brain[!(fin.r.squared_brain$names_brain %in% colnames(ts_protein_brain)),]
-r.squared_liver_notts = fin.r.squared_liver[!(fin.r.squared_liver$names_liver %in% colnames(ts_protein_liver)),]
+r.squared_brain_tsns = fin.r.squared_brain[fin.r.squared_brain$names_brain %in% colnames(tsns_protein_brain),]
+r.squared_liver_tsns = fin.r.squared_liver[fin.r.squared_liver$names_liver %in% colnames(tsns_protein_liver),]
 
@@
 
-gee = cbind(rep("not ts", times = dim(r.squared_brain_notts)[1]), r.squared_brain_notts[,2])
-gee2 = cbind(rep("ts", times = dim(r.squared_brain_ts)[1]), r.squared_brain_ts[,2])
+gee = cbind(rep("Tissue-Specific Non-Secreted", times = dim(r.squared_brain_tsns)[1]), r.squared_brain_tsns[,2])
+gee2 = cbind(rep("Tissue-Specific Secreted", times = dim(r.squared_brain_ts)[1]), r.squared_brain_ts[,2])
 
@@
 
-gee = cbind(rep("not ts", times = dim(r.squared_liver_notts)[1]), r.squared_liver_notts[,2])
-gee2 = cbind(rep("ts", times = dim(r.squared_liver_ts)[1]), r.squared_liver_ts[,2])
+gee = cbind(rep("Tissue-Specific Non-Secreted", times = dim(r.squared_brain_tsns)[1]), r.squared_brain_tsns[,2])
+gee2 = cbind(rep("Tissue-Specific Secreted", times = dim(r.squared_liver_ts)[1]), r.squared_liver_ts[,2])
 
@@
 
-boxplot(fin.r.squared_brain[,2], r.squared_brain_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Brain Data")
-boxplot(fin.r.squared_liver[,2], r.squared_liver_ts[,2], names = c("Not TS", "TS"), xlab = "R Squared", main = "R Squared Liver Data")
+boxplot(r.squared_brain_tsns[,2], r.squared_brain_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Brain Data")
+boxplot(r.squared_liver_tsns[,2], r.squared_liver_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Liver Data")
 
@@
 
-We tested whether the heritability of the non-tissue-specific and tissue specific datasets were significantly different using the Wilcoxon-Mann-Whitney Test.
-Both tests showed a significant difference at alpha = 0.05, however the p-value for the brain data was very close to 0.05.
+We tested whether the heritability of the Tissue-Specific Non-Secreted and Tissue-Specific Secreted datasets were significantly different using the Wilcoxon-Mann-Whitney Test.
+Only the liver dataset showed a significant difference betweeen the R Squared of the two groups.
 
-Non-Tissue-Specific Brain Dataset v. Tissue-Specific Brain Dataset:
+Tissue-Specific Non-Secreted Brain Dataset v. Tissue-Specific Secreted Brain Dataset:
 
@@
 
-Non-Tissue-Specific Liver Dataset v. Tissue-Specific Liver Dataset:
+Tissue-Specific Non-Secreted Liver Dataset v. Tissue-Specific Secreted Liver Dataset:
 

### 2021-09-14 SeldinReport.Rmd
fixed plot issues, small errors

---
title: "Seldin Report: Ssec Scores"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Find distribution of correlation p-values
  c. Find a null distribution of Ssec values

```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)
#library(reshape)
#library(ggplot2)

# Data Input

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)

```

Previously compared Secreted Genes to Tissue-Specific Secreted Genes. For both of these categories, they were compared to the full set of target tissue genes. There was not a very significant difference between the two peaks.  
Below are the graphs for Tissue-Specific Secreted Genes and Tissue-Specific Non-Secreted Genes. These were both compared to full set of target genes. These also do not show a significant difference between the Ssec scores of Tissue-Specific Secreted and Tissue-Specific Non-Secreted.

## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group All Target Genes

```{r, echo = FALSE, message = FALSE}
# Calculate Ssec values: tissue-specific -> full
#### TS ####
## BRAIN ##

brain.liv = bicorAndPvalue(ts_protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

ts_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(ts_Ssec_brain[,1])
bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain = bicorAndPvalue(ts_protein_liver, protein_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

ts_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(ts_Ssec_liver[,1])
bot_liver = tail(ts_Ssec_liver[,1])

#### TSNS ####
# ns tissue-specific -> ns tissue-specific
## BRAIN ##

brain.liv = bicorAndPvalue(tsns_protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

tsns_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain = bicorAndPvalue(tsns_protein_liver, protein_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

tsns_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
#top_liver = head(ts_Ssec_liver[,1])
#bot_liver = tail(ts_Ssec_liver[,1])

```

```{r, echo = FALSE}
# Visualizations
#hist(Ssec_brain$Ssec)
#hist(ts_Ssec_brain$Ssec)

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

```

The graphs below modify the comparison slightly. Instead of comparing the tissue specific secreted and tissue specific nonsecreted to the full dataset as target genes, our target genes are ALL tissue specific genes in the target tissue (secreted and non-secreted). This hopefully eliminates the overlap with the non-tissue-specific genes and doesn't limit the association to just secreted genes. There is a very slight difference in the density graphs for these. The Tissue-Specific Secreted have a VERY slightly higher Ssec score on average than the Tissue-Specific Non-Secreted.

## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group Tissue-Specific Target Genes

```{r, echo = FALSE, message = FALSE}
# Calculate Ssec values: tissue-specific -> ns tissue-specific
#### TS ####
## BRAIN ##
target_liver = cbind(ts_protein_liver, tsns_protein_liver)

brain.liv = bicorAndPvalue(ts_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

ts_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(ts_Ssec_brain[,1])
bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

target_brain = cbind(ts_protein_brain, tsns_protein_brain)

liv.brain = bicorAndPvalue(ts_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

ts_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(ts_Ssec_liver[,1])
bot_liver = tail(ts_Ssec_liver[,1])

#### TSNS ####
# ns tissue-specific -> ns tissue-specific
## BRAIN ##

brain.liv.c = bicorAndPvalue(tsns_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

tsns_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain.c = bicorAndPvalue(tsns_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

tsns_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
#top_liver = head(ts_Ssec_liver[,1])
#bot_liver = tail(ts_Ssec_liver[,1])

```

```{r, echo = FALSE}
# Visualizations
#hist(Ssec_brain$Ssec)
#hist(ts_Ssec_brain$Ssec)

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

compare.b = rbind(h1, h2)
colnames(compare.b) = c("trt", "val")
compare.b = as.data.frame(compare.b)
compare.b$trt = as.factor(compare.b$trt)
compare.b$val = as.numeric(compare.b$val)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

compare.l = rbind(h1, h2)
colnames(compare.l) = c("trt", "val")
compare.l = as.data.frame(compare.l)
compare.l$trt = as.factor(compare.l$trt)
compare.l$val = as.numeric(compare.l$val)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

```

```{r, echo = FALSE, message = FALSE}

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)

```

A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
We can try comparing the Tissue-Specific Secreted to the Ssec values of all of the origin tissue proteins. I need to create that dataset in the server; attempts to complete in chunks have been unsuccessful.

```{r, echo = FALSE}
## Pvalue for top and bottom Ssec scores, liver and brain
# brain
plotting = melt(brain.liv$p, id.vars = "refseq")
cheese =  as.character(plotting$Var1) %in% c(top_brain, bot_brain)
limit = plotting[cheese, ]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Genes", x = "P-Value", y = "Density")

# liver
plotting = melt(liv.brain$p, id.vars = "refseq")
limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Genes", x = "P-Value", y = "Density")

```

The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores. We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores. These graphs do not appear as well-defined as the graphs from previous reports where the all target tissue genes were used as the reference. 
[Attempted to put pvalues on log scale but lost the distinction between top and bottom Ssec scores; Tried to include a line for a null distribution, but was having trouble with ggplot]

```{r, echo = FALSE}
# Correlation coefficients top and bottom Ssec scores, liver and brain
# brain
magn = brain.liv$bicor

plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Genes", x = "Correlation", y = "Density")

# liver
magn = liv.brain$bicor
plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Genes", x = "Correlation", y = "Density")

```

Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation. The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients.

Below, we have used methods outlined in Seldin (2018) to complete a permutation test for the Ssec values of particular proteins. We can see with the significant protein below, the observed value is far above the 95% confidence interval of the null distribution. Below that, a protein that had the lowest Ssec value is far BELOW the 95% confidence interval (0.974 to 0.996). What this means has yet to be determined.

```{r, eval = FALSE, echo = FALSE, warning = FALSE}

#permute strain names in target tissue array
# bicor> pval> Ssec val
# repeat 1000- compare observed value to mean of permuted values
target_protein = protein_liver
perm_test = matrix(NA, ncol = 1000, nrow = 174)
perm_test = as.data.frame(perm_test)
rownames(perm_test) = ts_Ssec_brain[,1]

for (i in 1:1000)
{

perm = sample(rownames(target_protein), size = length(rownames(target_protein)), replace = FALSE)
target_protein = target_protein[perm,]

brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')

scores = rowSums(-log(brain.liv$p))

Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score)
  #arrange(desc(Ssec))

perm_test[,i] = Ssec_brain_perm[,2]
} 
```

```{r, echo = FALSE, warning = FALSE}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
perm_test = read.csv("Permutation_Seldin_Report.csv")
perm_test= fix.read.csv(perm_test)

test_top_b = perm_test[rownames(perm_test) == top_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == top_brain[1],]
hist(as.numeric(test_top_b), main = "Null Ssec Distribution for Top Gene: ENSRNOG00000021020", xlab = "Ssec Score")
abline(v = true, col = "red")

test_bot_b = perm_test[rownames(perm_test) == bot_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == bot_brain[1],]
hist(as.numeric(test_bot_b), main = "Null Ssec Distribution for Bottom Gene: ENSRNOG00000026419", xlab = "Ssec Score")
abline(v = true, col = "red")

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(perm_test, "Permutation_Seldin_Report.csv")

```

# still want to compare protein_brain, protein_liver- server

### 2021-10-07 Checking Secreted.Rmd
havent committed in a while

---
title: "Seldin Report: Secreted excluding Extracellular Matrix"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. See if significance improves after removing secreted proteins that are annotated for extracellular matrix

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)
#library(reshape)
#library(ggplot2)

# Data Input

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
notsecreted_protein_brain = read.csv("NotSecreted_Brain.csv")
notsecreted_protein_liver = read.csv("NotSecreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
#tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
#tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

#new_sec = read.delim("uniprot-rat+secreted+locations_(location__Secreted+[SL-0243]_+type--.tab.gz")
extracell = read.delim("uniprot-rat+extracellular+matrix+organism__Rattus+norvegicus+(Rat)+[10--.tab")
brain_uniprot = read.csv("Brain_Uniprot.csv")
liver_uniprot = read.csv("Liver_Uniprot.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
notsecreted_protein_brain = fix.read.csv(notsecreted_protein_brain)
notsecreted_protein_liver = fix.read.csv(notsecreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
#tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
#tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)
brain_uniprot = fix.read.csv(brain_uniprot)
liver_uniprot = fix.read.csv(liver_uniprot)

```

```{r, echo = FALSE}

# get ensembl ids for extracellular

`%!in%` <- Negate(`%in%`)

ens_brain = unique(brain_uniprot$ensembl_gene_id[brain_uniprot$uniprot_gn_id %in% extracell$Entry])
ens_liver = unique(liver_uniprot$ensembl_gene_id[liver_uniprot$uniprot_gn_id %in% extracell$Entry])

secreted_ens_brain = unique(colnames(secreted_protein_brain)[colnames(secreted_protein_brain) %in% ens_brain])
secreted_ens_liver = unique(colnames(secreted_protein_liver)[colnames(secreted_protein_liver) %in% ens_liver])

secnoextra_protein_brain = secreted_protein_brain[, colnames(secreted_protein_brain) %!in% secreted_ens_brain]
secnoextra_protein_liver = secreted_protein_liver[, colnames(secreted_protein_liver) %!in% secreted_ens_liver]

```

There is still no significant difference between secreted and non-secreted Ssec scores.

```{r, echo = FALSE}

notsecreted.1_protein_brain = cbind(notsecreted_protein_brain, protein_brain[colnames(protein_brain) %in% secreted_ens_brain])
notsecreted.1_protein_liver = cbind(notsecreted_protein_liver, protein_brain[colnames(protein_liver) %in% secreted_ens_liver])

#### Secreted ####
## BRAIN ##
target_liver = cbind(ts_protein_liver, tsns_protein_liver)

brain.liv = bicorAndPvalue(secnoextra_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

## LIVER ##

target_brain = cbind(ts_protein_brain, tsns_protein_brain)

liv.brain = bicorAndPvalue(secnoextra_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##

brain.liv.c = bicorAndPvalue(notsecreted.1_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain.c = bicorAndPvalue(notsecreted.1_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
#top_liver = head(ts_Ssec_liver[,1])
#bot_liver = tail(ts_Ssec_liver[,1])

####### Visualizations

#Brain
h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)

compare.b = rbind(h1, h2)
colnames(compare.b) = c("trt", "val")
compare.b = as.data.frame(compare.b)
compare.b$trt = as.factor(compare.b$trt)
compare.b$val = as.numeric(compare.b$val)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)

compare.l = rbind(h1, h2)
colnames(compare.l) = c("trt", "val")
compare.l = as.data.frame(compare.l)
compare.l$trt = as.factor(compare.l$trt)
compare.l$val = as.numeric(compare.l$val)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)
wil.brain

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)
wil.liver

# :'(

```

### 2021-10-07 SeldinReport.Rmd
havent committed in a while

@@
 
-```{r, echo = FALSE, message = FALSE}
+```{r echo=FALSE, message=FALSE}
 
@@
 secreted_protein_liver = read.csv("Secreted_Liver.csv")
+notsecreted_protein_brain = read.csv("NotSecreted_Brain.csv")
+notsecreted_protein_liver = read.csv("NotSecreted_Liver.csv")
 sec_annot = read.csv("Sec_Annotation.csv")
@@
 secreted_protein_liver = fix.read.csv(secreted_protein_liver)
+notsecreted_protein_brain = fix.read.csv(notsecreted_protein_brain)
+notsecreted_protein_liver = fix.read.csv(notsecreted_protein_liver)
 sec_annot = fix.read.csv(sec_annot)
@@
 
+```{r}
+# checking secreted
+
+checking = sec_annot[sec_annot$ensembl_gene_id == top_brain[4],] # extracellular matrix
+
+# ENSRNOG00000005770: "The WNT signalling modulator, Wise, is expressed in an interaction-dependent manner during hair-follicle cycling."
+
+```
 A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
@@
 
+```{r}
+# annotation
+
+anno_brain = sec_annot[sec_annot$ensembl_gene_id %in% top_brain,]
+anno_liver = sec_annot[sec_annot$ensembl_gene_id %in% top_liver,]
+
+```
 ```{r, eval = FALSE, echo = FALSE, warning = FALSE}
@@
 # still want to compare protein_brain, protein_liver- server
+
+```{r, echo = FALSE}
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+full_brain = read.csv("Full_Brain_asso_TS.TSNS_Ssec.csv")
+full_liver = read.csv("Full_Liver_asso_TS.TSNS_Ssec.csv")
+
+# Visualizations
+#hist(Ssec_brain$Ssec)
+#hist(ts_Ssec_brain$Ssec)
+
+# Ssec distribtions TSNS v. TS, brain and liver
+#Brain
+h1 = cbind(rep("Full brain", times = length(full_brain$Ssec)), full_brain$Ssec)
+h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
+
+compare.b = rbind(h1, h2)
+colnames(compare.b) = c("trt", "val")
+compare.b = as.data.frame(compare.b)
+compare.b$trt = as.factor(compare.b$trt)
+compare.b$val = as.numeric(compare.b$val)
+
+ggplot(compare.b, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Brain Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+# Liver
+h1 = cbind(rep("Full liver", times = length(full_liver$Ssec)), full_liver$Ssec)
+h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
+
+compare.l = rbind(h1, h2)
+colnames(compare.l) = c("trt", "val")
+compare.l = as.data.frame(compare.l)
+compare.l$trt = as.factor(compare.l$trt)
+compare.l$val = as.numeric(compare.l$val)
+
+ggplot(compare.l, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+```
+# compare secreted to non-secreted
+
+```{r}
+
+#### Secreted ####
+## BRAIN ##
+target_liver = cbind(ts_protein_liver, tsns_protein_liver)
+
+brain.liv = bicorAndPvalue(secreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
+top_brain = head(sec_Ssec_brain[,1])
+bot_brain = tail(sec_Ssec_brain[,1])
+
+## LIVER ##
+
+target_brain = cbind(ts_protein_brain, tsns_protein_brain)
+
+liv.brain = bicorAndPvalue(secreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain$p))
+
+sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
+top_liver = head(sec_Ssec_liver[,1])
+bot_liver = tail(sec_Ssec_liver[,1])
+
+#### Not Secreted ####
+
+## BRAIN ##
+
+brain.liv.c = bicorAndPvalue(notsecreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv.c$p))
+
+nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
+#top_brain = head(tsns_Ssec_brain[,1])
+#bot_brain = tail(ts_Ssec_brain[,1])
+
+## LIVER ##
+
+liv.brain.c = bicorAndPvalue(notsecreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain.c$p))
+
+nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
+#top_liver = head(ts_Ssec_liver[,1])
+#bot_liver = tail(ts_Ssec_liver[,1])
+
+####### Visualizations
+
+#Brain
+h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
+h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)
+
+compare.b = rbind(h1, h2)
+colnames(compare.b) = c("trt", "val")
+compare.b = as.data.frame(compare.b)
+compare.b$trt = as.factor(compare.b$trt)
+compare.b$val = as.numeric(compare.b$val)
+
+ggplot(compare.b, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+# Liver
+h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
+h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)
+
+compare.l = rbind(h1, h2)
+colnames(compare.l) = c("trt", "val")
+compare.l = as.data.frame(compare.l)
+compare.l$trt = as.factor(compare.l$trt)
+compare.l$val = as.numeric(compare.l$val)
+
+ggplot(compare.l, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+#brain
+
+wil.brain = coin::wilcox_test(val ~ trt, compare.b)
+wil.brain
+
+# liver
+
+wil.liver = coin::wilcox_test(val ~ trt, compare.l)
+wil.liver
+
+# :'(
+
+```
+
+```{r}
+
+# random brain secreted genes
+
+sample(secreted_protein_brain, size = 4, replace = FALSE)
+
+#ENSRNOG00000012655: extracellular matrix
+# ENSRNOG00000053272: endo reticulum, extracellular space, cytoplasm, perinuclear region
+
+checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000012655",]
+checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000001806",]
+checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000008519",]
+
+checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000021020",]
+
+# random liber secreted genes
+
+sample(secreted_protein_liver, size = 4, replace = FALSE)
+
+# ENSRNOG00000058444: nucleus, chromosome, secreted
+# ENSRNOG00000048642: same
+
+```

### 2021-10-07 Tissue Specific Report.Rmd
havent committed in a while

@@
 
-The ensembl rat liver and brain RNAseq data was acquired (get more info from papers) downloaded from PhenoGen.
+The ensembl rat liver and brain RNAseq data was produced by our lab from tissues imported from the Czech Academy of Sciences laboratory.
 Inbred strains were excluded, leaving only recombinant inbred strains.
+The means across strains were calculated for the RNA expression data.
 Annontation data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set.
@@
 
-From there, our datasets were segregated by terms secreted and tissue-specific.
-Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021).
+From there, our datasets were segregated by terms "secreted" and "tissue-specific." Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021).
 The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets.
@@
 From this we created our dataset of investigation: tissue-specific secreted gene datasets for liver (with 35 genes) and brain (with 174 genes) tissues.
-This dataset was compared to a comparable dataset of tissue-specific non-secreted genes, with _ liver genes and _ brain genes.
+This dataset was compared to a comparable dataset of tissue-specific non-secreted genes, with 368 liver genes and 2,577 brain genes.
 
@@
 # Data Input
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 protein_brain = read.csv("Brain_WGCNA.csv")
@@
 ```{r, echo = FALSE, message = FALSE}
+
+# flow chart
+# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
+library(DiagrammeR)
+
+grViz(diagram = "digraph flowchart { 
+      node [fontname = Arial, shape = oval, color = Lavender, style = filled]
+      tab1 [label = '@@1', fontsize = 100]
+      tab2 [label = '@@2', fontsize = 100]
+      tab3 [label = '@@3', fontsize = 100]
+      tab4 [label = '@@4', fontsize = 100]
+      tab5 [label = '@@5', fontsize = 100]
+      tab6 [label = '@@6', fontsize = 100]
+      tab7 [label = '@@7', fontsize = 100]
+      tab8 [label = '@@8', fontsize = 100]
+
+      edge[minlen = 4, penwidth = 8, arrowsize = 4]
+      tab1 -> tab2;
+      tab2 -> tab3;
+      tab3 -> tab4;
+      tab4 -> tab5;
+      tab5 -> tab6;
+      tab4 -> tab7;
+      tab7 -> tab8}
+
+      [1]: 'Full Cleaned Data: Brain (132 individuals, 18,385 genes), Liver (132 individuals, 15,365 genes)'
+      [2]: 'Means of Strains: Brain (45 strains, 18,385 genes), Liver (45 strains, 15,365 genes)'
+      [3]: 'RI Strains Only: Brain (30 strains, 18,385 genes), Liver (30 strains, 15,365 genes)'
+      [4]: 'Protein-Coding Genes Only: Brain (30 strains, 16,457 genes), Liver (30 strains, 14,109 genes)'
+      [5]: 'Secreted Protein-Coding Genes: B(30, 657), L(30, 818)'
+      [6]: 'Tissue-Specific Secreted Protein-Coding Genes: B(30, 174), L(30, 35)'
+      [7]: 'Non-Secreted Protein-Coding Genes: B(30, 15,800), L(30, 13,591)'
+      [8]: 'Tissue-Specific Non-Secreted Protein-Coding Genes: B(30, 2,577), L(30, 368)' 
+      ", width = 1000, height = 400)
+
+```
+
+```{r, echo = FALSE, message = FALSE}
 # looking for unique genes
@@
 ## Tissue Specific Non-Secreted
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 notsec_brain = read.csv("NotSecreted_Brain.csv")
@@
 
-knitr::kable(num, caption = "Number of Genes", col.names = NA, )
-
-# compare shared genes between brain and liver
-
-gee = intersect(colnames(ts_protein_brain), colnames(ts_protein_liver))
-
-int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
-names(int) = c("Tissue-Specific Secreted Genes Brain", "Tissue-Specific Secreted Genes Shared by Brain and Liver", "Tissue-Specific Secreted Genes Liver")
-
-knitr::kable(int)
-
-hee = intersect(colnames(tsns_protein_brain), colnames(tsns_protein_liver))
-
-int2 = c(length(colnames(tsns_protein_brain)), length(hee), length(colnames(tsns_protein_liver)))
-names(int2) = c("Tissue-Specific Non-Secreted Genes Brain", "Tissue-Specific Non-Secreted Genes Shared by Brain and Liver", "Tissue-Specific Non-Secreted Genes Liver")
-
-knitr::kable(int2)
+knitr::kable(num, col.names = "Number of Ensembl Genes")
 
@@
 wil.b = as.data.frame(wil.b)
-colnames(wil.b) = c("trt", "r.sq")
+colnames(wil.b) = c("Dataset", "r.sq")
 
-wil.b$trt = as.factor(wil.b$trt)
+wil.b$Dataset = as.factor(wil.b$Dataset)
 wil.b$r.sq = as.numeric(wil.b$r.sq)
 
-wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b)
+wil.brain = coin::wilcox_test(r.sq ~ Dataset, wil.b)
 
@@
 wil.l = as.data.frame(wil.l)
-colnames(wil.l) = c("trt", "r.sq")
+colnames(wil.l) = c("Dataset", "r.sq")
 
-wil.l$trt = as.factor(wil.l$trt)
+wil.l$Dataset = as.factor(wil.l$Dataset)
 wil.l$r.sq = as.numeric(wil.l$r.sq)
 
-wil.liver = coin::wilcox_test(r.sq ~ trt, wil.l)
+wil.liver = coin::wilcox_test(r.sq ~ Dataset, wil.l)
 
@@
 
-ggplot(wil.b, aes(x = r.sq, color = trt)) + 
+ggplot(wil.b, aes(x = r.sq, color = Dataset)) + 
   geom_density() +
-  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")
+  labs(title ="Broad Sense Heritability Density Curve: Brain",x = "R Squared", y = "Density")
 
-ggplot(wil.l, aes(x = r.sq, color = trt)) + 
+ggplot(wil.l, aes(x = r.sq, color = Dataset)) + 
   geom_density() +
-  labs(title="R Squared Density Curve: Liver",x="R Squared", y = "Density")
+  labs(title ="Broad Sense Heritability Density Curve: Liver",x = "R Squared", y = "Density")
 
@@
 
-boxplot(r.squared_brain_tsns[,2], r.squared_brain_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Brain Data")
-boxplot(r.squared_liver_tsns[,2], r.squared_liver_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Liver Data")
+boxplot(r.squared_brain_tsns[,2], r.squared_brain_ts[,2], names = c("Brain-Specific Non-Secreted", "Brain-Specific Secreted"), main = "Broad Sense Heritability: Brain", ylab = "R Squared")
+boxplot(r.squared_liver_tsns[,2], r.squared_liver_ts[,2], names = c("Liver-Specific Non-Secreted", "Liver-Specific Secreted"), main = "Broad Sense Heritability: Liver", ylab = "R Squared")
 
@@
 Only the liver dataset showed a significant difference betweeen the R Squared of the two groups.
-
-Tissue-Specific Non-Secreted Brain Dataset v. Tissue-Specific Secreted Brain Dataset:
-
-```{r, echo = FALSE, message = FALSE}
-wil.brain
-
-```
-
-Tissue-Specific Non-Secreted Liver Dataset v. Tissue-Specific Secreted Liver Dataset:
-
-```{r, echo = FALSE, message = FALSE}
-
-wil.liver
-
-```
+The test comparing the Brain-Specific Non-Secreted and Brain-Specific Secreted resulted in a p-value of `r round(coin::pvalue(wil.brain), 2)`.
+The test comparing the Liver-Specific Non-Secreted and Liver-Specific Secreted resulted in a p-value of `r format(coin::pvalue(wil.liver), scientific = TRUE, digits = 3)`.

### 2021-10-12 New Tissue Specific.Rmd
deleting some stuff

---
title: "Report for Uniprot-Identified Tissue-Specific Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Summarize the genes identified as tissue-specific by uniprot
  a. Look at the criteria for tissue-specificity within uniprot
2. Run Ssec code with these genes
  a. Uniprot-identified Tissue-Specific and exclusive to the tissues
  b. Uniprot-identified Tissue-Specific, without excluding those not exclusive to the tissues
3. Determine the characteristics of those that are labeled as tissue-specific but are not exclusive to one tissue

Here we are restricting our protein-coding RNA-seq data to those genes that are labeled as "tissue-specific" on the uniprot database. [Seldin method]. [Definition of tissue-specific on uniprot]. We will calculate Ssec scores for those labeled as tissue-specific on uniprot, comparing secreted to non-secreted. These will be labeled ts uniprot. Additionally, we will calculate these same scores for genes that are labeled tissue-specific AND are only found in the corresponding dataset (brain genes are not found in the liver dataset and vice versa). These will be labeled ts uniprot exclusive. 
Seldin: The Ssec rankings were then filtered ensuring that genes were
primarily expressed by the origin tissue. This was accomplished
by manual inspection of each gene in multi-tissue expression
panels available via BioGPS (Su et al., 2002; Wu et al., 2009) in
both mouse and humans (representative examples using
BioGPS arrays shown in Figure S3). Genes not detected in
BioGPS arrays were investigated for tissue specificity using the
UniProt literature query for ‘‘tissue specificity.’’
This was done after calculation of Ssec matrices. Manual inspection of top genes
For this particular test, it does not appear that they focused on tissue-specific genes at all. Solely focused on the categories secreted and nonsecreted.They tested this with QQPlots
What are BioGPS Arrays?
Would it be useful at all to look at individuals instead of strain means?

## CHECK INF!!!
Do I want to try full dataset as target?

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)
#library(reshape)
#library(ggplot2)

# Data Input

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
notsecreted_protein_brain = read.csv("NotSecreted_Brain.csv")
notsecreted_protein_liver = read.csv("NotSecreted_Liver.csv")
brain_uniprot = read.csv("Brain_Uniprot.csv")
liver_uniprot = read.csv("Liver_Uniprot.csv")
#sec_annot = read.csv("Sec_Annotation.csv")
#tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
#tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
#ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
#ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
#tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
#tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

uniprot.tissue = read.delim("uniprot-annotation_(type__tissue+specificity_)-filtered-organism_%--.tab")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
notsecreted_protein_brain = fix.read.csv(notsecreted_protein_brain)
notsecreted_protein_liver = fix.read.csv(notsecreted_protein_liver)
#sec_annot = fix.read.csv(sec_annot)
#tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
#tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
#ts_protein_brain = fix.read.csv(ts_protein_brain)
#ts_protein_liver = fix.read.csv(ts_protein_liver)
#tsns_protein_brain = fix.read.csv(tsns_protein_brain)
#tsns_protein_liver = fix.read.csv(tsns_protein_liver)
brain_uniprot = fix.read.csv(brain_uniprot)
liver_uniprot = fix.read.csv(liver_uniprot)

```

Below is a table of the number of genes in the labeled categories. We can see the sample size go down dramatically with the conditioning on tissue exclusivity. The first Ssec score analysis below will be using the Secreted TS Uniprot Exclusive brain and liver datasets (the last two rows below). The second Ssec score analysis will be using the Secreted TS Uniprot brain and liver datasets (the fifth and sixth rows).

```{r, echo = FALSE}
# find brain and liver tissue-specific

these_brain = brain_uniprot$ensembl_gene_id[brain_uniprot$uniprot_gn_id %in% uniprot.tissue$Entry]
these_liver = liver_uniprot$ensembl_gene_id[liver_uniprot$uniprot_gn_id %in% uniprot.tissue$Entry]

ts_uni_protein_brain = protein_brain[, colnames(protein_brain) %in% these_brain]
ts_uni_protein_liver = protein_liver[, colnames(protein_liver) %in% these_liver]

tssec_uni_protein_brain = secreted_protein_brain[, colnames(secreted_protein_brain) %in% these_brain]
tssec_uni_protein_liver = secreted_protein_liver[, colnames(secreted_protein_liver) %in% these_liver]

`%!in%` <- Negate(`%in%`)

tsns_uni_protein_brain = ts_uni_protein_brain[, colnames(ts_uni_protein_brain) %!in% colnames(tssec_uni_protein_brain)]
tsns_uni_protein_liver = ts_uni_protein_liver[, colnames(ts_uni_protein_liver) %!in% colnames(tssec_uni_protein_liver)]

# exclude the ones that are shared

ts.ts_uni_protein_brain = ts_uni_protein_brain[, colnames(ts_uni_protein_brain) %!in% colnames(ts_uni_protein_liver)]
ts.ts_uni_protein_liver = ts_uni_protein_liver[, colnames(ts_uni_protein_liver) %!in% colnames(ts_uni_protein_brain)]

ts.tssec_uni_protein_brain = tssec_uni_protein_brain[, colnames(tssec_uni_protein_brain) %!in% colnames(tssec_uni_protein_liver)]
ts.tssec_uni_protein_liver = tssec_uni_protein_liver[, colnames(tssec_uni_protein_liver) %!in% colnames(tssec_uni_protein_brain)]

ts.tsns_uni_protein_brain = tsns_uni_protein_brain[, colnames(tsns_uni_protein_brain) %!in% colnames(tsns_uni_protein_liver)]
ts.tsns_uni_protein_liver = tsns_uni_protein_liver[, colnames(tsns_uni_protein_liver) %!in% colnames(tsns_uni_protein_brain)]

numgenes = c(dim(ts_uni_protein_brain)[2], dim(ts_uni_protein_liver)[2], dim(ts.ts_uni_protein_brain)[2], dim(ts.ts_uni_protein_liver)[2], dim(tssec_uni_protein_brain)[2], dim(tssec_uni_protein_liver)[2], dim(ts.tssec_uni_protein_brain)[2], dim(ts.tssec_uni_protein_liver)[2])

names(numgenes) = c("TS Uniprot Brain", "TS Uniprot Liver", "TS Uniprot Exclusive Brain", "TS Uniprot Exclusive Liver", "Secreted TS Uniprot Brain", "Secreted TS Uniprot Liver", "Secreted TS Uniprot Exclusive Brain", "Secreted TS Uniprot Exclusive Liver")

knitr::kable(numgenes, caption = "Number of Genes")

```

## Secreted TS Uniprot Exclusive

The following analysis is comparing the Ssec distribution between secreted and non-secreted genes. The Wilcoxon-Mann-Whitney test shows a very significant difference in the liver dataset, though with the very small sample size of 8 genes, we cannot be sure of the accuracy of our calculations. The brain dataset continues to not be significantly different between secreted and non-secreted.

```{r, echo = FALSE}

#### Secreted ####
## BRAIN ##
target_liver = ts.ts_uni_protein_liver

brain.liv = bicorAndPvalue(ts.tssec_uni_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

## LIVER ##

target_brain = ts.ts_uni_protein_brain

liv.brain = bicorAndPvalue(ts.tssec_uni_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##

brain.liv.c = bicorAndPvalue(ts.tsns_uni_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain.c = bicorAndPvalue(ts.tsns_uni_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
#top_liver = head(ts_Ssec_liver[,1])
#bot_liver = tail(ts_Ssec_liver[,1])

####### Visualizations

#Brain
h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)

compare.b = rbind(h1, h2)
colnames(compare.b) = c("trt", "val")
compare.b = as.data.frame(compare.b)
compare.b$trt = as.factor(compare.b$trt)
compare.b$val = as.numeric(compare.b$val)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)

compare.l = rbind(h1, h2)
colnames(compare.l) = c("trt", "val")
compare.l = as.data.frame(compare.l)
compare.l$trt = as.factor(compare.l$trt)
compare.l$val = as.numeric(compare.l$val)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)
wil.brain

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)
wil.liver

# :'(

```

## Secreted TS Uniprot

The following analysis is comparing the Ssec distribution between secreted and non-secreted genes. The Wilcoxon-Mann-Whitney test shows a highly significant difference between secreted and non-secreted for the liver dataset. The brain dataset, while not showing a significant p-value, is the closest to significance found in our analyses.

```{r, echo = FALSE}

#### Secreted ####
## BRAIN ##
target_liver = ts_uni_protein_liver

brain.liv = bicorAndPvalue(tssec_uni_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

## LIVER ##

target_brain = ts_uni_protein_brain

liv.brain = bicorAndPvalue(tssec_uni_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##

brain.liv.c = bicorAndPvalue(tsns_uni_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

liv.brain.c = bicorAndPvalue(tsns_uni_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
#top_liver = head(ts_Ssec_liver[,1])
#bot_liver = tail(ts_Ssec_liver[,1])

####### Visualizations

#Brain
h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)

compare.b = rbind(h1, h2)
colnames(compare.b) = c("trt", "val")
compare.b = as.data.frame(compare.b)
compare.b$trt = as.factor(compare.b$trt)
compare.b$val = as.numeric(compare.b$val)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)

compare.l = rbind(h1, h2)
colnames(compare.l) = c("trt", "val")
compare.l = as.data.frame(compare.l)
compare.l$trt = as.factor(compare.l$trt)
compare.l$val = as.numeric(compare.l$val)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)
wil.brain

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)
wil.liver

# :'(

```

## Looking

```{r}

`%!in%` <- Negate(`%in%`)

not_ex_brain = colnames(ts_uni_protein_brain)[colnames(ts_uni_protein_brain) %!in% colnames(ts.ts_uni_protein_brain)]
not_ex_liver = colnames(ts_uni_protein_liver)[colnames(ts_uni_protein_liver) %!in% colnames(ts.ts_uni_protein_liver)]

# checking the reads of the genes in not_ex_brain in brain and liver (i did not code this incorrectly)
not_ex_protein_brain = protein_brain[,colnames(protein_brain) %in% not_ex_brain]
not_ex_protein_liver = protein_liver[,colnames(protein_liver) %in% not_ex_brain]

#sum(colnames(not_ex_protein_brain) != colnames(not_ex_protein_liver))

avg_reads_not_ex_brain = apply(not_ex_protein_brain, 2, mean)
avg_reads_not_ex_liver = apply(not_ex_protein_liver, 2, mean)
label = c(rep("brain", times = length(avg_reads_not_ex_brain)), rep("liver", times = length(avg_reads_not_ex_liver)))
effort = cbind(label, c(avg_reads_not_ex_brain, avg_reads_not_ex_liver))
colnames(effort) = c("trt", "val")
effort = as.data.frame(effort)
effort$trt = factor(effort$trt, levels = c("brain", "liver"))
effort$val = as.numeric(effort$val)

ggplot(effort, aes(val, color = trt)) + 
  geom_density() +
  labs(title = "Average Reads of Brain and Liver Samples that are Tissue-Specific but NOT Exclusive to Brain/Liver", x = "Ssec", y = "Density")

wil.effort = coin::wilcox_test(val ~ trt, effort)
wil.effort

```
It seems that in the uniprot database, tissue-specicificity does not indicate that a protein is only found in a single tissue, rather that there is SOME KIND of specificity in where it is found.

ENSRNOG00000003732: from top_brain; "Detected in brain (at protein level)"
ENSRNOG00000023337: from top_brain; "Expressed in the dorsal root ganglia"
ENSRNOG00000018962: from top_brain; "Expressed in the ventricle and atrium of adult rats. Also detected in the lung, kidney, liver, skeletal muscle, stomach and urinary bladder. Not detected in brain, colon, testis, spleen or thymus. Overexpressed in the ventricles in the case of hypertension and hypertrophy." in the ensembl database, they say there is expression in brain with an expression score of 56.67, compared to 77.93 in the liver

ENSRNOG00000002811: from the Secreted TS Uniprot Exclusive Brain dataset; "Expressed in olfactory bulb, cerebellum and neuronal layers in hippocampus."

ENSRNOG00000050761: from the Secreted TS Uniprot Exclusive Liver dataset; "Expressed by the liver and secreted in plasma."

ENSRNOG00000012196: from the Secreted TS Uniprot Liver dataset; "Highly expressed in brain, kidney and heart (PubMed:11328816). Expressed at lower level in other tissues such as liver (PubMed:11328816). Expressed in intestine, kidney and liver (at protein level) (PubMed:11328816, PubMed:11330410). Localizes in the epithelia of the jejunum and ileum (PubMed:11330410)."

ENSRNOG00000020889: not exclusive but tissue specific brain; "Minor constituent of the mineralized matrix of bone. Detected in calvaria, rib cartilage, liver, kidney, spleen, brain, lung, skeletal and heart muscle with highest expression in calvaria and approximately half the amount in kidney, liver and brain."

### 2022-05-23 Data_Summary_Overview.Rmd
havent committed in a while pt1

---
title: "Data Visualizations for HRDP RNA Data from the Brain and Liver"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Range of genes in histogram
2. Strain means via RLE
3. Broad Sense Heritability
4. Hierarchical Clustering Plot
5. PCA plot

```{r, echo = FALSE}
# Data Import

# set working directory
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")

# download brain and liver datasets
brain = read.csv("PhenoGen.HRDP.v5.totalRNA.Brain.gene.ensembl96.txt", header = T, sep = "\t")
liver = read.csv("PhenoGen.HRDP.v5.totalRNA.Liver.gene.ensembl96.txt", header = T, sep = "\t")

# download strain means dataset
storagematrix_brain = read.csv("Means_strainXgene_brain.csv")
storagematrix_liver = read.csv("Means_strainXgene_liver.csv")

# download BSH vectors
fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")

# download RLE matrices
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
RLE_brain = read.csv("RLE_Brain.csv")
RLE_liver = read.csv("RLE_Liver.csv")
RLE_brain = RLE_brain[,2:dim(RLE_brain)[2]]
RLE_liver = RLE_liver[,2:dim(RLE_liver)[2]]
```

### Range Histograms

The following are histograms of the range of values for each of the datasets. We can see in the first two graphs that there are some pretty extreme outliers in each of the datasets. The brain dataset shows values greater than 15, and the liver dataset shows values greater than 80. The zoomed in histograms are controlled by x-axis and bar width. We can see that the values are generally concentrated at the lower end with a right skew. Brain appears to have greater concentration at the lower end than liver. It seems there is more variability in the liver dataset.

```{r, echo = FALSE}

# Range of Genes

# distribution of gene ranges
range_brain = numeric(dim(brain)[1])
for (i in 1:dim(brain)[1])
{
  r = range(brain[i,])
  m = r[2] - r[1]
  range_brain[i] = m
}

range_liver = numeric(dim(liver)[1])
for (i in 1:dim(liver)[1])
{
  r = range(liver[i,])
  m = r[2] - r[1]
  range_liver[i] = m
}

hist(range_brain, main = "Range by Gene: Brain", xlab = "Range")
hist(range_liver, main = "Range by Gene: Liver", xlab = "Range")

hist(range_brain, main = "Range by Gene: Brain, Zoom", xlab = "Range", xlim = range(0:15))
hist(range_liver, main = "Range by Gene: Liver, Zoom", xlab = "Range", xlim = range(0:15), breaks = 80)

boxplot(range_brain, main = "Range by Gene: Brain, Zoom", ylim = range(0:10))
boxplot(range_liver, main = "Range by Gene: Liver, Zoom", ylim = range(0:10))

```

### RLE Plots

Below are the Relative Log Odds plots for the brain and liver data, including zoomed plots. We can see that with a couple exceptions in the liver data, the data is pretty evenly distributed around 0. Should I identify the outliers in the liver dataset?

```{r, echo = FALSE}

# RLE

boxplot(x = t(RLE_brain), main = "RLE plot: Brain dataset", xlab = "Strain")
boxplot(x = t(RLE_liver), main = "RLE plot: Liver dataset", xlab = "Strain")

boxplot(x = t(RLE_brain), main = "RLE plot: Brain dataset ZOOM", xlab = "Strain", ylim = range(-5:5))
boxplot(x = t(RLE_liver), main = "RLE plot: Liver dataset ZOOM", xlab = "Strain", ylim = range(-5:5))

```

### Broad Sense Heritability

The Broad Sense Heritability for the brain and liver datasets is very similar to each other and close to normally distributed.

```{r, echo = FALSE}

#Broad Sense Heritability

hist(as.numeric(fin.r.squared_brain[,3]), main = "Broad Sense Heritability: Brain", xlab = "R-Squared Value")
hist(as.numeric(fin.r.squared_liver[,3]), main = "Broad Sense Heritability: Liver", xlab = "R-Squared Value")
```

### Hierarchical Clustering

Below we have Hierarchical Clustering Dendrograms that are color coded by strain, batch, and recombinant v. inbred. For the dendrogram coded by batch, we can see some clusters of similar colors, but not complete separation. For recombinant v. inbred, there is significant clustering, as expected.

```{r, echo = FALSE, message = FALSE}
library(WGCNA)

#Hierarchical Clustering

for_clustering_brain = as.dist(1 - abs(cor(x = brain, method = "spearman")))
for_clustering_liver = as.dist(1 - abs(cor(x = liver, method = "spearman")))

h_clustering_brain = hclust(for_clustering_brain)
h_clustering_liver = hclust(for_clustering_liver)

# By Strain

rats_brain = sub("_.*", "", colnames(brain))
rats_liver = sub("_.*", "", colnames(liver))

color_brain = labels2colors(rats_brain)
color_liver = labels2colors(rats_liver)

#plotDendroAndColors(h_clustering_brain, color_brain, main = "Brain by Strain", cex.dendroLabels = 0.4)
#plotDendroAndColors(h_clustering_liver, color_liver, main = "Liver by Strain", cex.dendroLabels = 0.4)

# By Batch

batch_brain = as.numeric(sub(".*batch", "", colnames(brain)))
batch_liver = as.numeric(sub(".*batch", "", colnames(liver)))

color_batch_brain = labels2colors(batch_brain)
color_batch_liver = labels2colors(batch_liver)

#plotDendroAndColors(h_clustering_brain, color_batch_brain, main = "Brain by Batch")
#plotDendroAndColors(h_clustering_liver, color_batch_liver, main = "Liver by Batch")

# By Recombinant v. Inbred

hxb_brain = grep("HXB", colnames(brain))
bxh_brain = grep("BXH", colnames(brain))
recombinant_brain = sort(c(hxb_brain,bxh_brain))
recomblog_brain = 1:length(colnames(brain))
recomblog_brain[recombinant_brain] = 0
recomblog_brain = as.logical(recomblog_brain)

hxb_liver = grep("HXB", colnames(liver))
bxh_liver = grep("BXH", colnames(liver))
recombinant_liver = sort(c(hxb_liver,bxh_liver))
recomblog_liver = 1:length(colnames(liver))
recomblog_liver[recombinant_liver] = 0
recomblog_liver = as.logical(recomblog_liver)

color_recomb_brain = labels2colors(recomblog_brain)
color_recomb_liver = labels2colors(recomblog_liver)

totcolors_brain = cbind(color_brain, color_batch_brain, color_recomb_brain)
colnames(totcolors_brain) = c("Strain", "Batch", "Recombinant v. Inbred")
totcolors_liver = cbind(color_liver, color_batch_liver, color_recomb_liver)
colnames(totcolors_liver) = c("Strain", "Batch", "Recombinant v. Inbred")

plotDendroAndColors(h_clustering_brain, colors = totcolors_brain, main = "Brain", cex.dendroLabels = 0.4)
plotDendroAndColors(h_clustering_liver, colors = totcolors_liver, main = "Liver", cex.dendroLabels = 0.4)

```

### PCA Plot

Here we have the PCA plots, coded by recombinant v. inbred and batch. Again, there is pretty distinct separation of color for the recombinant v. inbred, as expected. Batch is somewhat unreadable

```{r, echo = FALSE}

# PCA Plot

# PCA
library(PCAtools)

pca_brain <- pca(brain, center=TRUE, scale=TRUE)
pca_liver = pca(liver, center = TRUE, scale = TRUE)

#biplot(pca_brain)
#biplot(pca_liver)

# by recombinant v. inbred

mat = matrix(recomblog_brain, nrow = length(recomblog_brain), ncol = 1)
rownames(mat) = colnames(brain)
colnames(mat) = c("col")

mat2 = matrix(recomblog_liver, nrow = length(recomblog_liver), ncol = 1)
rownames(mat2) = colnames(liver)
colnames(mat2) = c("col")

pca_brain = pca(brain, meta = mat, center=TRUE, scale=TRUE)
pca_liver = pca(liver, meta = mat2, center = TRUE, scale = TRUE)

biplot(pca_brain, colby = "col", title = "Brain by Recombinant v. Inbred", lab = NULL)
biplot(pca_liver, colby = "col", title = "Liver by Recombinant v. Inbred", lab = NULL)

# by batch
color_pca_batchbrain = labels2colors(batch_brain)
mat = matrix(color_pca_batchbrain, nrow = length(batch_brain), ncol = 1)
rownames(mat) = colnames(brain)
colnames(mat) = c("col")

color_pca_batchliver = labels2colors(batch_liver)
mat2 = matrix(color_pca_batchliver, nrow = length(batch_liver), ncol = 1)
rownames(mat2) = colnames(liver)
colnames(mat2) = c("col")

pca_brain = pca(brain, meta = mat, center=TRUE, scale=TRUE)
pca_liver = pca(liver, meta = mat2, center = TRUE, scale = TRUE)

biplot(pca_brain, colby = "col", title = "Brain by Batch", lab = NULL)
biplot(pca_liver, colby = "col", title = "Liver by Batch", lab = NULL)
```

### 2022-05-23 Intro_Summary.Rmd
havent committed in a while pt1

# Data Summary of HRDP RNA Data from the Brain and Liver
## Angela Yoder

Goals: Download Phenogen HRDP total RNA ensemble data for Brain and Liver and complete a data summary

```{r, echo = FALSE}
# set working directory
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

# download brain and liver datasets
brain = read.csv("PhenoGen.HRDP.v5.totalRNA.Brain.gene.ensembl96.txt", header = T, sep = "\t")
liver = read.csv("PhenoGen.HRDP.v5.totalRNA.Liver.gene.ensembl96.txt", header = T, sep = "\t")

# number of samples
numsamples_brain = ncol(brain)
numsamples_liver = ncol(liver)
```

The brain dataset contains `r numsamples_brain` samples. The liver dataset contains `r numsamples_liver` samples.

```{r, echo = FALSE}

# number of strains

  # extracting strain from colname
rats_brain = sub("_.*", "", colnames(brain))
rats_liver = sub("_.*", "", colnames(liver))

  # counting unique strains
numstrains_brain = length(unique(rats_brain))
numstrains_liver = length(unique(rats_liver))

# number of rats in each strain

tbl_brain = table(rats_brain)
tbl_liver = table(rats_liver)

# avg number of rats in each strain

avgsample_brain = mean(tbl_brain)
avgsample_liver = mean(tbl_liver)

ex.avgsample_brain = format(round(avgsample_brain, 2), nsmall = 2)
ex.avgsample_liver = format(round(avgsample_liver, 2), nsmall = 2)

```

There are `r numstrains_brain` strains of rat in the brain dataset, and `r numstrains_liver` in the liver dataset.
There are an average of `r ex.avgsample_brain` samples per strain in the brain dataset, and `r ex.avgsample_liver` in the liver dataset.

```{r, echo = FALSE}

# num genes in each dataset

numgene_brain = nrow(brain)
numgene_liver = nrow(liver)

# number of overlapping genes between datasets

genes_brain = rownames(brain)
genes_liver = rownames(liver)

overlap = match(genes_brain, genes_liver)

overlap = na.omit(overlap)
numoverlap = length(overlap)

ex.numgene_brain = formatC(numgene_brain, big.mark = ",")
ex.numgene_liver = formatC(numgene_liver, big.mark = ",")
ex.numoverlap = formatC(numoverlap, big.mark = ",")

```

There are `r ex.numgene_brain` genes in the brain dataset, and `r ex.numgene_liver` genes in the liver dataset.
There are `r ex.numoverlap` genes that are shared between the two datasets.

```{r, echo = FALSE}

# distribution of gene ranges
range_ay = function(x)
{
  r = range(x)
  m = r[2] - r[1]
  out = 2^m
  return(out)
}

range_brain = apply(brain, 1, range_ay)
range_liver = apply(liver, 1, range_ay)

#range_brain = numeric(dim(brain)[1])
#for (i in 1:dim(brain)[1])
#{
#  r = range(brain[i,])
#  m = r[2] - r[1]
#  range_brain[i] = 2^m
#}

#range_liver = numeric(dim(liver)[1])
#for (i in 1:dim(liver)[1])
#{                                                   
#  r = range(liver[i,])
#  m = r[2] - r[1]
#  range_liver[i] = 2^m
#}

meanrange_liver = mean(range_liver)
meanrange_brain = mean(range_brain)

### testing ###
 #liver[which.max(range_liver), grep('LEW', colnames(liver))]

 outlier_liver = which(range_liver > 32)
 outlier_brain = which(range_brain > 32)

 #find rat names that have outliers
 beep = lapply(brain[outlier_brain,], which.max)
 beepbeep = names(beep)
 #table(beepbeep) All the rats have outliers
 choo = lapply(liver[outlier_liver,], which.max)
 choochoo = names(choo)
 #table(choochoo) All the rats have outliers

 #liver[as.numeric(choo), grep('ACI', colnames(liver))]
 #brain[as.numeric(beep), grep("SS.JrHsd", colnames(brain))]

 #double checking outlier ranges
 max_liver = as.numeric(lapply(liver[outlier_liver,], max))
 min_liver = as.numeric(lapply(liver[outlier_liver,], min))
 checking_liver = max_liver - min_liver

 max_brain = as.numeric(lapply(brain[outlier_brain,], max))
 min_brain = as.numeric(lapply(brain[outlier_brain,], min))
 checking_brain = max_brain - min_brain

# find the largest values for range
sort_range_liver = sort(range_liver)
sort_range_brain = sort(range_brain)

# pull out outliers from liver ranges
outliers = tail(sort_range_liver, n = 3)

# logical vector to weed out outliers
outlier_logical_liver = range_liver > 32
outlier_logical_brain = range_brain > 32

# find indices and names of genes that are outliers
index = c()
name = c()
count = 1
for (i in outlier_logical_liver)
{
  if (i == 0)
  {
    index = c(index, count)
    name = c(name, rownames(liver[count,]))
  }
  count = count + 1
}

# summ is number of values < 5 per bad genes
# geene is the index of the rats that have the values that are too high
new_outlier_liver = liver < 5

#geene = c()
#for (i in index)
#{
#  for (j in 1:length(new_outlier_liver[i,]))
#  {
#    if (new_outlier_liver[i,j] == 0)
#    {
#      geene = c(geene, j)
#    }
#  }
#}

summ = apply(new_outlier_liver[index,], 1, sum)
#apply(new_outlier_liver[index,], 2, grep, pattern = 0)
#apply(new_outlier_liver[index,], 2, )

# max_rat is the maximum value for each rat
#max_rat = c()
#max_index = c()
#numrats = length(liver[1,])
#numgenes = length(liver[,1])
#for (i in 1:numrats)
#{
#  max_rat = c(max_rat, max(liver[,i]))
#  for (j in 1: numgenes)
#  {
#    if (liver[j,i] == max(liver[,i]))
#    {
#      max_index = c(max_index, j)
#    }
#  }
#}

max_rat = apply(liver, 2, max)

#max_gene is the max value for each gene
#max_gene = c()
#for (i in 1: numgenes)
#{
#  max_gene = c(max_gene, max(liver[i,]))
#}

max_gene = apply(liver, 1, max)

log_max_gene = max_gene > 15

# find proportion of range vectors that are outliers
prop_outlier_brain = sum(outlier_logical_brain) - length(outlier_logical_brain)
prop_outlier_liver = sum(outlier_logical_liver) - length(outlier_logical_liver)

# create new vectors with outliers removed with logical vector from before
new_range_liver = outlier_logical_liver * range_liver
new_range_brain = outlier_logical_brain * range_brain

# histograms of new vectors
hist(new_range_brain[new_range_brain != 0], main = "Range of Genes with Outliers Excluded, Brain", xlab = "Range")
hist(new_range_liver[new_range_liver != 0], main = "Range of Genes with Outliers Excluded, Liver", xlab = "Range")
```

outliers greater than 32 (2^5). There were 191 outliers from the brain dataset 228 outliers in the liver dataset

The average range of reads per gene for the brain dataset is `r meanrange_brain` and `r meanrange_liver` for the liver dataset. Above are this histograms for ranges of reads per gene.

```{r, echo = FALSE}

#hist(range_liver, main = "Range of Number of Reads for RNA found in the Liver", xlab = "Number of Reads")
#hist(range_brain, main = "Range of Number of Reads for RNA found in the Brain", xlab = "Number of Reads")

```

### 2022-05-23 QQ Plots.Rmd
havent committed in a while pt1

---
title: "Seldin Report: Ssec QQ Plots"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Calculate Ssec scores for Secreted and Non-Secreted proteins v. full datasets
  a. Plot Ssec scores in a QQ plot
2. Calculate Ssec scores for Tissue-Exclusive Secreted and Tissue-Exclusive Non-Secreted proteins v. full datasets
  b. Plot Ssec scores in a QQ plot

Following the procedure laid out in Seldin, we are comparing secreted proteins to non-secreted proteins from the origin tissue, without the designation of tissue-specific. This was compared to the full protein-coding dataset for the target tissue.

### Secreted v. Non-Secreted

```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
sec_Ssec_brain = read.csv("Sec-Full-Ssec_brain.csv")
sec_Ssec_liver = read.csv("Sec-Full-Ssec_liver.csv")
nsec_Ssec_brain = read.csv("NSec-Full-Ssec_brain.csv")
nsec_Ssec_liver = read.csv("NSec-Full-Ssec_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

sec_Ssec_brain = fix.read.csv(sec_Ssec_brain)
sec_Ssec_liver = fix.read.csv(sec_Ssec_liver)
nsec_Ssec_brain = fix.read.csv(nsec_Ssec_brain)
nsec_Ssec_liver = fix.read.csv(nsec_Ssec_liver)
```

Below are the QQ plots for the Ssec values comparing secreted to non-secreted. Ideally, Ssec values would have higher values in secreted proteins, showing that the Ssec measure was quantifying communication between tissues. However, the graphs show the opposite, with non-secreted proteins having higher correlations to the target tissue proteins than the secreted proteins.

```{r, echo = FALSE}

## BRAIN

sec = sec_Ssec_brain$Ssec
nsec = nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = sec_Ssec_liver$Ssec
nsec = nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
abline(a = 0, b = 1)

```

We can confirm by looking at the maximum values for each designation.

```{r, echo = FALSE}

top_sec_brain = head(sec_Ssec_brain)
top_nsec_brain = head(nsec_Ssec_brain)
top_sec_liver = head(sec_Ssec_liver)
top_nsec_liver = head(nsec_Ssec_liver)

out = rbind(top_sec_brain$Ssec, top_nsec_brain$Ssec, top_sec_liver$Ssec, top_nsec_liver$Ssec)
rownames(out) = c("Secreted Brain", "Non-Secreted Brain", "Secreted Liver", "Non-Secreted Liver")

knitr::kable(out, caption = "Top Ssec Scores by Category")

```

### Tissue-Exclusive Secreted v. Tissue-Exclusive Non-Secreted

Below is the same procedure, but with only RNA sequences that are exclusive to the tissue of interest. Liver appears to favor secreted proteins a bit more, albeit with a much lower sample size. Brain, however, still favors non-secreted protein associations. We can try to repeat this procedure, changing target genes to also be tissue-exclusive, instead of all protein-coding genes.

```{r, echo = FALSE}

`%!in%` <- Negate(`%in%`)

te.sec_Ssec_brain = sec_Ssec_brain[sec_Ssec_brain$Gene_symbol %!in% sec_Ssec_liver$Gene_symbol,]
te.sec_Ssec_liver = sec_Ssec_liver[sec_Ssec_liver$Gene_symbol %!in% sec_Ssec_brain$Gene_symbol,]
te.nsec_Ssec_brain = nsec_Ssec_brain[nsec_Ssec_brain$Gene_symbol %!in% nsec_Ssec_liver$Gene_symbol,]
te.nsec_Ssec_liver = nsec_Ssec_liver[nsec_Ssec_liver$Gene_symbol %!in% nsec_Ssec_brain$Gene_symbol,]

## BRAIN

sec = te.sec_Ssec_brain$Ssec
nsec = te.nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = te.sec_Ssec_liver$Ssec
nsec = te.nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Liver Proteins"))
abline(a = 0, b = 1)

```

```{r, echo = FALSE, eval = FALSE}

# server code for sec TE v. nsec TE

#### Secreted ####
## BRAIN ##
te.sec_brain = protein_brain[, colnames(protein_brain) %in% rownames(te.sec_Ssec_brain)]
target_liver = protein_liver

brain.liv = bicorAndPvalue(te.sec_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

## LIVER ##

target_brain = protein_brain
te.sec_liver = protein_liver[, colnames(protein_liver) %in% rownames(te.sec_Ssec_liver)]

liv.brain = bicorAndPvalue(te.sec_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##
te.nsec_brain = protein_brain[, colnames(protein_brain) %in% rownames(te.nsec_Ssec_brain)]

brain.liv.c = bicorAndPvalue(te.nsec_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])

## LIVER ##

te.nsec_liver = protein_liver[, colnames(protein_liver) %in% rownames(te.nsec_Ssec_liver)]

liv.brain.c = bicorAndPvalue(te.nsec_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

```

```{r, echo = FALSE, eval = FALSE}
# double checking
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
te.sec_Ssec_brain = read.csv("TE.Sec_Ssec_brain.csv")
te.sec_Ssec_liver = read.csv("TE.Sec_Ssec_liver.csv")
te.nsec_Ssec_brain = read.csv("TE.NSec_Ssec_brain.csv")
te.nsec_Ssec_liver = read.csv("TE.NSec_Ssec_liver.csv")

te.sec_Ssec_brain = fix.read.csv(te.sec_Ssec_brain)
te.sec_Ssec_liver = fix.read.csv(te.sec_Ssec_liver)
te.nsec_Ssec_brain = fix.read.csv(te.nsec_Ssec_brain)
te.nsec_Ssec_liver = fix.read.csv(te.nsec_Ssec_liver)

## BRAIN

sec = te.sec_Ssec_brain$Ssec
nsec = te.nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = te.sec_Ssec_liver$Ssec
nsec = te.nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Liver Proteins"))
abline(a = 0, b = 1)

```

```{r}
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
associate_brain = read.csv("Associations_Brain.csv")
associate_liver = read.csv("Associations_Liver.csv")

top_brain = head(te.sec_Ssec_brain[,1])
bot_brain = tail(te.sec_Ssec_brain[,1])

assoc_top_brain = associate_brain[associate_brain$Var1 %in% top_brain,]
top_assoc_top_brain = assoc_top_brain[assoc_top_brain$value < 0.05,]
one = top_assoc_top_brain[top_assoc_top_brain$Var1 == top_brain[1],]

top_liver = head(te.sec_Ssec_liver[,1])
bot_liver = tail(te.sec_Ssec_liver[,1])

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(one, "OneSecretedProtein_LessThan5Per.csv")

```

### checking target shit

```{r, echo = FALSE}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

ts_Ssec_target_brain = read.csv("Target_TESec_Ssec_brain.csv")
ts_Ssec_target_liver = read.csv("Target_TESec_Ssec_liver.csv")
tsns_Ssec_target_brain = read.csv("Target_TENSec_Ssec_brain.csv")
tsns_Ssec_target_liver = read.csv("Target_TENSec_Ssec_liver.csv")
## BRAIN

sec = ts_Ssec_target_brain$Ssec
nsec = tsns_Ssec_target_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = ts_Ssec_target_liver$Ssec
nsec = tsns_Ssec_target_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
abline(a = 0, b = 1)

```

### 2022-05-23 Secreted Proteins Report.1.Rmd
havent committed in a while pt1

---
title: "Report for Secreted Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report results of secreted proteins data

The original data of ensembl rat liver and brain RNAseq data was downloaded from BioMart database (accessed 2021). Inbred strains were excluded, leaving only recombinant inbred strains. Enrichment data was retrieved from BiomaRt (version number `r packageVersion("biomaRT")`), which was used to exclude non-protein-coding genes from the data set. This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.

From there, the data was further reduced to only secreted proteins. This was done using a list of secreted proteins from the uniprot database (accessed 2021). Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted." This resulted in a list of 1,772 secreted proteins attributed to Rattus norvegicus. The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets. This comparison in 657 secreted proteins from the brain dataset, and 518 secreted proteins from the liver dataset.

The following report was completed using `r R.Version()$version.string`. Additionally, the packages WGCNA (version `r packageVersion("WGCNA")`), broom (version `r packageVersion("broom")`), tidyverse (version `r packageVersion("tidyverse")`), and knitr (version `r packageVersion("knitr")`).

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
#notsec_protein_brain = read.csv("NotSecreted_Brain.csv")
#notsec_protein_liver = read.csv("NotSecreted_Liver.csv")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")
fin.r.squared_liver_sec = read.csv("R_Squared_Liver_Sec.csv")
fin.r.squared_brain_notsec = read.csv("R_Squared_Brain_NotSec.csv")
fin.r.squared_liver_notsec = read.csv("R_Squared_Liver_NotSec.csv")

#funct_brain = read.csv("Function_Annotation_Brain.csv")
#funct_brain_sec = read.csv("Function_Annotation_Brain_Sec.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
#funct_brain = fix.read.csv(funct_brain)
#funct_brain_sec = fix.read.csv(funct_brain_sec)
protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
#notsec_protein_brain = fix.read.csv(notsec_protein_brain)
#notsec_protein_liver = fix.read.csv(notsec_protein_liver)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)
fin.r.squared_liver_sec = fix.read.csv(fin.r.squared_liver_sec)
fin.r.squared_brain_notsec = fix.read.csv(fin.r.squared_brain_notsec)
fin.r.squared_liver_notsec = fix.read.csv(fin.r.squared_liver_notsec)

```

The following tables show the number of proteins for each dataset. Here we can see that secreted proteins make up a small proportion of genes expressed in the liver and brain tissues. Of the brain proteins, secreted proteins make up approximately 4%; of the liver proteins, secreted proteins make up 3.8%. Additionally, of all secreted proteins found in the rat proteome, those produced by the brain and liver are 37% and 29% respectively. There is also a significant overlap in secreted proteins in the brain and liver. In the second table, we can see that 74% of proteins secreted by the brain are also secreted by the liver, and 93% vice versa.

```{r, echo = FALSE, message = FALSE}
# table comparing number of genes
ngenes = ncol(protein_brain)
ngenes_sec = ncol(secreted_protein_brain)

ngenes.l = ncol(protein_liver)
ngenes_sec.l = ncol(secreted_protein_liver)
tot_sec = 1760 # from secreted uniprot data

num = c(ngenes, ngenes_sec, ngenes.l, ngenes_sec.l, tot_sec)

names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Liver: Full Data", "Liver: Secreted Data", "Total Secreted Rat Proteins")

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

# compare shared genes between secreted brain and secreted liver

gee = intersect(colnames(secreted_protein_brain), colnames(secreted_protein_liver))

int = c(length(colnames(secreted_protein_brain)), length(gee), length(colnames(secreted_protein_liver)))
names(int) = c("Secreted Proteins Brain", "Secreted Proteins Shared by Brain and Liver", "Secreted Proteins Liver")

knitr::kable(int)

```

```{r, echo = FALSE, message = FALSE}
# hypothesis test wilcoxon mann whitney

#brain

gee = cbind(rep("not secreted", times = dim(fin.r.squared_brain_notsec)[1]), fin.r.squared_brain_notsec[,2])
gee2 = cbind(rep("secreted", times = dim(fin.r.squared_brain_sec)[1]), fin.r.squared_brain_sec[,2])

wil.b = rbind(gee, gee2)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("trt", "r.sq")

wil.b$trt = as.factor(wil.b$trt)
wil.b$r.sq = as.numeric(wil.b$r.sq)

wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b)

# liver

gee = cbind(rep("not secreted", times = dim(fin.r.squared_liver_notsec)[1]), fin.r.squared_liver_notsec[,2])
gee2 = cbind(rep("secreted", times = dim(fin.r.squared_liver_sec)[1]), fin.r.squared_liver_sec[,2])

wil.l = rbind(gee, gee2)
wil.l = as.data.frame(wil.l)
colnames(wil.l) = c("trt", "r.sq")

wil.l$trt = as.factor(wil.l$trt)
wil.l$r.sq = as.numeric(wil.l$r.sq)

wil.liver = coin::wilcox_test(r.sq ~ trt, wil.l)

```

The following boxplots and density plots show an increase in median value and a decrease in right skew that may indicate that the secreted dataset has higher relative heritability.

```{r, echo = FALSE, message = FALSE}
# Hist R Squared

#hist(as.numeric(fin.r.squared_brain_sec[,2]), main = "Broad Sense Heritability: Secreted Brain", xlab = "R-Squared Value")
#hist(as.numeric(fin.r.squared_brain[,2]), main = "Broad Sense Heritability: Full Brain", xlab = "R-Squared Value")

#hist(as.numeric(fin.r.squared_liver_sec[,2]), main = "Broad Sense Heritability: Secreted Liver", xlab = "R-Squared Value")
#hist(as.numeric(fin.r.squared_liver[,2]), main = "Broad Sense Heritability: Full Liver", xlab = "R-Squared Value")

# Density R Squared

ggplot(wil.b, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")

ggplot(wil.l, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Liver",x="R Squared", y = "Density")

# Box R Squared

boxplot(fin.r.squared_brain[,2], fin.r.squared_brain_sec[,2], names = c("Not Secreted", "Secreted"), xlab = "R Squared", main = "R Squared Brain Data")
boxplot(fin.r.squared_liver[,2], fin.r.squared_liver_sec[,2], names = c("Not Secreted", "Secreted"), xlab = "R Squared", main = "R Squared Liver Data")

```

We tested whether the heritability of the non-secreted and secreted datasets were significantly different using the Wilcoxon-Mann-Whitney Test. Both the tests of the brain and liver data showed that the two datasets were significantly different at an alpha = 0.05 level.

```{r, echo = FALSE, message = FALSE}
print("Non-Secreted Brain Dataset v. Secreted Brain Dataset")
wil.brain

print("Non-Secreted Liver Dataset v. Secreted Liver Dataset")
wil.liver

R.Version()$version.string

```

### 2022-05-23 Secreted Proteins Report.Rmd
havent committed in a while pt1

---
title: "Report for Secreted Proteins"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report results from Data Cleaning.Rmd and Annotation.Rmd
Total secreted in rat genome
Introduction on secreted proteins. Where got them from. How found them in my data. 
Could make a series of reports of the two Rmd
Overlap between the two tissues

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")

funct_brain = read.csv("Function_Annotation_Brain.csv")
funct_brain_sec = read.csv("Function_Annotation_Brain_Sec.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
funct_brain = fix.read.csv(funct_brain)
funct_brain_sec = fix.read.csv(funct_brain_sec)
protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)

```

```{r, echo = FALSE, message = FALSE}
# table comparing number of genes
ngenes = ncol(protein_brain)
ngenes_sec = ncol(secreted_protein_brain)

ngenes.l = ncol(protein_liver)
ngenes_sec.l = ncol(secreted_protein_liver)

num = c(ngenes, ngenes_sec, ngenes.l, ngenes_sec.l)

names(num) = c("Brain: Full Data", "Brain: Secreted Data", "Liver: Full Data", "Liver: Secreted Data")

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

```

Density plots and overlaying. Compare medians, quartiles- boxplot- ttest - might not show biological relavence. - use confidence intervals nonparametric test
```{r, echo = FALSE, message = FALSE}
# Hist R Squared

hist(as.numeric(fin.r.squared_brain_sec[,2]), main = "Broad Sense Heritability: Secreted Brain", xlab = "R-Squared Value")
hist(as.numeric(fin.r.squared_brain[,2]), main = "Broad Sense Heritability: Full Brain", xlab = "R-Squared Value")

```

exclude cellular components - no good- enrichment more useful later on
eqtl : find point in geneome changing expression; usually near it- transcription factor binding site; P30 webinars- intro to expression QTL and QTL; go to every snp; ttest of quant value at each snp; looking for really low pvalue. rpackage called qtl- need to get data in right format- tutorials- email; genotype dataset in email (full_hrdp_geno)- BIGGER project
```{r, echo = FALSE, message = FALSE}
# Annotation Comparison

most = slice_max(funct_brain_sec, order_by = funct_brain_sec$Freq, n = 30)
murst = slice_max(funct_brain, order_by = funct_brain$Freq, n = 30)
most = most[1:30,]

most = most[order(most[,3], decreasing = TRUE),]
murst = murst[order(murst[,3], decreasing = TRUE),]

fin = cbind(as.character(most[,1]), round(most[,3], digits = 3), as.character(murst[,1]), round(murst[,3], digits = 3))
colnames(fin) = c("Full: Protein Function", "Full: Proportion", "Secreted: Protein Function", "Secreted: Proportion")

knitr::kable(fin)

```

### 2022-05-23 SeldinReport.Rmd
haven't committed in a while pt2

@@
   c. Find a null distribution of Ssec values
+
+THIS WAS DONE WITH TISSUE EXCLUSIVE GENES
 
@@
 sec_annot = read.csv("Sec_Annotation.csv")
-tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
-tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
+#tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
+#tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
 ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
@@
 sec_annot = fix.read.csv(sec_annot)
-tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
-tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
+#tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
+#tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
 ts_protein_brain = fix.read.csv(ts_protein_brain)
@@
 #### TSNS ####
-# ns tissue-specific -> ns tissue-specific
+# ns tissue-specific -> full
 ## BRAIN ##
@@
 
+with(compare, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
+abline(a = 0, b = 1)
+
 # Liver
@@
   labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+with(compare, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
+abline(a = 0, b = 1)
 
@@
 
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
+write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
+write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
+write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
+
 ```
@@
 
+with(compare.b, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
+abline(a = 0, b = 1)
+
 # Liver
@@
   labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+with(compare.l, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
+abline(a = 0, b = 1)
 
@@
 
+with(compare.b, qqplot(val[trt == "Not Secreted brain"], val[trt == "Secreted brain"]))
+abline(a = 0, b = 1)
+
 # Liver
@@
   labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+with(compare.l, qqplot(val[trt == "Not Secreted liver"], val[trt == "Secreted liver"]))
+abline(a = 0, b = 1)
 
@@
 ```
+
+```{r}
+
+# QQ plot
+
+sec = ts_Ssec_brain$Ssec
+nsec = tsns_Ssec_brain$Ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"]))
+abline(a = 0, b = 1)
+
+```

### 2022-05-23 SummaryStatistics.Rmd
haven't committed in a while pt2

---
title: "Summary Statistics"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Create tables and graphics of summary statistics for Thesis Proposal

```{r, echo = FALSE, message = FALSE}
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")
pheno = read.csv("Pheno_rQTL.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

pheno = fix.read.csv(pheno)

#libraries

library(tidyverse)

```
## Phenotypes ##
```{r, echo = FALSE}
# Phenotypes

clear = as.data.frame(cbind(rownames(pheno), as.numeric(pheno$ClearanceEtOH)))
clear$V2 = as.numeric(clear$V2)
clear = clear[order(clear$V2),]
clear$V2 = round(clear$V2, digits = 8)
clear$V1 = factor(clear$V1, levels = clear$V1)

auc = as.data.frame(cbind(rownames(pheno), as.numeric(pheno$meanauc2gkg)))
auc$V2 = as.numeric(auc$V2)
auc = auc[order(auc$V2),]
#auc$V2 = round(auc$V2, digits = 8)
auc$V1 = factor(auc$V1, levels = auc$V1)
```

```{r, echo = FALSE}

# boxplot for phenotype Alcohol Clearance

#caption_clear = str_wrap("The rate of alcohol clearance of each strain, averaged over multiple replicates, for recombinant-inbred HXB/BXH rat strains. Individuals were injected with 2 g/kg ethanol, and blood samples were taken at multiple intervals to track blood alcohol concentration. Broad Sense Heritability for this phenotype was calculated (81%).")

#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#pdf(file = "Barplot_Pheno_Clear.pdf", width = 12, height = 7)

# error bars
# http://www.sthda.com/english/wiki/ggplot2-barplots-quick-start-guide-r-software-and-data-visualization

ggplot(data = clear, aes(x = V1, y = V2)) +
  geom_bar(stat = "identity", fill = "green4") +
  theme_minimal() +
  xlab("Strain") + 
  ylab("Alcohol Clearance (mL/min/kg)") +
  theme(text = element_text(size = 20), 
        plot.title = element_text(size = 45, face="bold", margin = margin(10, 0, 10, 0)),
        axis.text.x = element_text(angle = 90)) +
  theme_minimal()

#dev.off()
```

The rate of alcohol clearance of each strain, averaged over multiple replicates, for recombinant-inbred HXB/BXH rat strains. Individuals were injected with 2 g/kg ethanol, and blood samples were taken at multiple intervals to track blood alcohol concentration. Broad Sense Heritability for this phenotype was calculated (81%).

```{r, echo = FALSE}

# boxplot for phenotype Acetate AUC

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
pdf(file = "Barplot_Pheno_AUC.pdf", width = 12, height = 7)

caption_auc = str_wrap("The averaged total area under the curve of the plot of blood acetate concentration over time of each strain for recombinant-inbred HXB/BXH rat strains. Individuals were injected with 2 g/kg ethanol, and blood samples were taken at multiple intervals to track blood acetate concentration. Broad Sense Heritability for this phenotype was calculated (66%).")

ggplot(data = auc, aes(x = V1, y = V2)) +
  geom_bar(stat = "identity", fill = "green4") +
  theme_minimal() +
  xlab("Strain") + 
  ylab("Acetate AUC from 0 to 400 minutes (mM*min)") +
  theme(text = element_text(size = 20), 
        plot.title = element_text(size = 45, face="bold", margin = margin(10, 0, 10, 0)),
        axis.text.x = element_text(angle = 90))

dev.off()

```

The averaged total area under the curve of the plot of blood acetate concentration over time of each strain for recombinant-inbred HXB/BXH rat strains. Individuals were injected with 2 g/kg ethanol, and blood samples were taken at multiple intervals to track blood acetate concentration. Broad Sense Heritability for this phenotype was calculated (66%).

## Broad Sense Heritability ##

```{r, echo = FALSE}
# Sec/NonSec code taken from "Secreted Proteins Report.1.Rmd"
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_brain_sec = read.csv("R_Squared_Brain_Sec.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")
fin.r.squared_liver_sec = read.csv("R_Squared_Liver_Sec.csv")
fin.r.squared_brain_notsec = read.csv("R_Squared_Brain_NotSec.csv")
fin.r.squared_liver_notsec = read.csv("R_Squared_Liver_NotSec.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_brain_sec = fix.read.csv(fin.r.squared_brain_sec)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)
fin.r.squared_liver_sec = fix.read.csv(fin.r.squared_liver_sec)
fin.r.squared_brain_notsec = fix.read.csv(fin.r.squared_brain_notsec)
fin.r.squared_liver_notsec = fix.read.csv(fin.r.squared_liver_notsec)

```

```{r, echo = FALSE}

# Data Frame Creation
gee = cbind(rep("not secreted", times = dim(fin.r.squared_brain_notsec)[1]), fin.r.squared_brain_notsec[,2])
gee2 = cbind(rep("secreted", times = dim(fin.r.squared_brain_sec)[1]), fin.r.squared_brain_sec[,2])

wil.b = rbind(gee, gee2)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("trt", "r.sq")

wil.b$trt = as.factor(wil.b$trt)
wil.b$r.sq = as.numeric(wil.b$r.sq)

gee = cbind(rep("not secreted", times = dim(fin.r.squared_liver_notsec)[1]), fin.r.squared_liver_notsec[,2])
gee2 = cbind(rep("secreted", times = dim(fin.r.squared_liver_sec)[1]), fin.r.squared_liver_sec[,2])

wil.l = rbind(gee, gee2)
wil.l = as.data.frame(wil.l)
colnames(wil.l) = c("trt", "r.sq")

wil.l$trt = as.factor(wil.l$trt)
wil.l$r.sq = as.numeric(wil.l$r.sq)

# ttest

ttest_brain = t.test(r.sq ~ trt, data = wil.b)
ttest_liver = t.test(r.sq ~ trt, data = wil.l)

```

```{r, echo = FALSE}

## Plot BSH Density Curves

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")
pdf(file = "BSH_brain.pdf", width = 12, height = 7)

ggplot(wil.b, aes(x = r.sq, color = trt)) + 
  geom_density(size = 2) + 
  labs(x ="R Squared", y = "Density") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("secreted", "not secreted"),
                            labels=c("Secreted: 657 genes", "Non-Secreted: 15,800 genes")) +
  theme(panel.background = element_rect(fill = "white"), panel.grid.major = element_line(color = "gray87"), legend.position = "top", text = element_text(size = 20))

dev.off()
```

CAPTION

```{r, echo = FALSE}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")
pdf(file = "BSH_liver.pdf", width = 12, height = 7)

ggplot(wil.l, aes(x = r.sq, color = trt)) + 
  geom_density(size = 2) + 
  labs(x ="R Squared", y = "Density") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("secreted", "not secreted"),
                            labels=c("Secreted: 818 genes", "Non-Secreted: 13,519 genes")) +
  theme(panel.background = element_rect(fill = "white"), panel.grid.major = element_line(color = "gray87"), legend.position = "top", text = element_text(size = 20))

dev.off()

```

CAPTION

```{r, echo = FALSE}
# TE Sec/NonSec code taken from "Tissue Specific Report.Rmd"
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

te.sec_brain = read.csv("TE.Secreted_Protein_Brain.csv")
te.sec_liver = read.csv("TE.Secreted_Protein_Liver.csv")
te.nsec_brain = read.csv("TE.NotSecreted_Protein_Brain.csv")
te.nsec_liver = read.csv("TE.NotSecreted_Protein_Liver.csv")

te.sec_brain = fix.read.csv(te.sec_brain)
te.sec_liver = fix.read.csv(te.sec_liver)
te.nsec_brain = fix.read.csv(te.nsec_brain)
te.nsec_liver = fix.read.csv(te.nsec_liver)

```

```{r, echo = FALSE}
# R squared for TE datasets

rsq.brain = sub(" ~ strain_factor_brain.*", "", fin.r.squared_brain$names_brain)
rsq.liver = sub(" ~ strain_factor_liver.*", "", fin.r.squared_liver$names_liver)

fin.r.squared_brain[,1] = rsq.brain
fin.r.squared_liver[,1] = rsq.liver

r.squared_brain_tesec = fin.r.squared_brain[fin.r.squared_brain$names_brain %in% colnames(te.sec_brain),]
r.squared_liver_tesec = fin.r.squared_liver[fin.r.squared_liver$names_liver %in% colnames(te.sec_liver),]

r.squared_brain_tens = fin.r.squared_brain[fin.r.squared_brain$names_brain %in% colnames(te.nsec_brain),]
r.squared_liver_tens = fin.r.squared_liver[fin.r.squared_liver$names_liver %in% colnames(te.nsec_liver),]

## Data Frame Creation
#brain

gee = cbind(rep("Tissue-Exclusive Non-Secreted", times = dim(r.squared_brain_tens)[1]), r.squared_brain_tens[,2])
gee2 = cbind(rep("Tissue-Exclusive Secreted", times = dim(r.squared_brain_tesec)[1]), r.squared_brain_tesec[,2])

wil.b_te = rbind(gee, gee2)
wil.b_te = as.data.frame(wil.b_te)
colnames(wil.b_te) = c("Dataset", "r.sq")

wil.b_te$Dataset = as.factor(wil.b_te$Dataset)
wil.b_te$r.sq = as.numeric(wil.b_te$r.sq)

# liver

gee = cbind(rep("Tissue-Exclusive Non-Secreted", times = dim(r.squared_brain_tens)[1]), r.squared_brain_tens[,2])
gee2 = cbind(rep("Tissue-Exclusive Secreted", times = dim(r.squared_liver_tesec)[1]), r.squared_liver_tesec[,2])

wil.l_te = rbind(gee, gee2)
wil.l_te = as.data.frame(wil.l_te)
colnames(wil.l_te) = c("Dataset", "r.sq")

wil.l_te$Dataset = as.factor(wil.l_te$Dataset)
wil.l_te$r.sq = as.numeric(wil.l_te$r.sq)

# ttest

ttest_tebrain = t.test(r.sq ~ Dataset, data = wil.b_te)
ttest_teliver = t.test(r.sq ~ Dataset, data = wil.l_te)

ttest_compbrain = t.test(wil.b_te$r.sq, wil.b$r.sq)

```

```{r, echo = FALSE}
# Plot BSH density curves

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
pdf(file = "BSH_te_brain.pdf", width = 12, height = 7)

ggplot(wil.b_te, aes(x = r.sq, color = Dataset)) + 
  geom_density(size = 2) + 
  labs(x ="R Squared", y = "Density") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("Tissue-Exclusive Secreted", "Tissue-Exclusive Non-Secreted"),
                            labels=c("Tissue-Exclusive Secreted: 174 genes", "Tissue-Exclusive Non-Secreted: 2,589 genes")) +
  theme(panel.background = element_rect(fill = "white"), panel.grid.major = element_line(color = "gray87"), legend.position = "top", text = element_text(size = 20))

dev.off()

```

caption

```{r, echo = FALSE}

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
pdf(file = "BSH_te_liver.pdf", width = 12, height = 7)

ggplot(wil.l_te, aes(x = r.sq, color = Dataset)) + 
  geom_density(size = 2) + 
  labs(x ="R Squared", y = "Density") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("Tissue-Exclusive Secreted", "Tissue-Exclusive Non-Secreted"),
                            labels=c("Tissue-Exclusive Secreted: 35 genes", "Tissue-Exclusive Non-Secreted: 371 genes")) +
  theme(panel.background = element_rect(fill = "white"), panel.grid.major = element_line(color = "gray87"), legend.position = "top", text = element_text(size = 20))

dev.off()

```

caption

## Ssec Scores

```{r}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)

```

```{r}

plot_brain = data.frame(class = c(rep("Tissue-Exclusive Secreted", times = dim(tesec_brain)[1]), rep("Tissue-Exclusive Non-Secreted", times = dim(tensec_brain)[1])), Ssec = c(tesec_brain[,2], tensec_brain[,2]))

plot_liver = data.frame(class = c(rep("Tissue-Exclusive Secreted", times = dim(tesec_liver)[1]), rep("Tissue-Exclusive Non-Secreted", times = dim(tensec_liver)[1])), Ssec = c(tesec_liver[,2], tensec_liver[,2]))

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
pdf(file = "Ssec_te_brain.pdf", width = 12, height = 7)

ggplot(plot_brain, aes(x = Ssec, color = class)) +
  geom_histogram(fill = "white") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("Tissue-Exclusive Secreted", "Tissue-Exclusive Non-Secreted")) +
  labs(title ="Ssec Values: Brain",x="Ssec Value", y = "", subtitle = "Tissue-Exclusive Secreted v. Tissue-Exclusive Non-Secreted") +
  theme_minimal()

dev.off()

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
pdf(file = "Ssec_te_liver.pdf", width = 12, height = 7)

ggplot(plot_liver, aes(x = Ssec, color = class)) +
  geom_histogram(fill = "white") +
  scale_color_manual(values = c("green4", "lawngreen"), name  ="",
                            breaks=c("Tissue-Exclusive Secreted", "Tissue-Exclusive Non-Secreted")) +
  labs(title ="Ssec Values: Liver",x="Ssec Value", y = "", subtitle = "Tissue-Exclusive Secreted v. Tissue-Exclusive Non-Secreted") +
  theme_minimal()

dev.off()

```

$$ Accuracy = 1 - \frac{|P - O|}  {O} $$

### 2022-05-23 Sussing Tissue Specific.Rmd
haven't committed in a while pt2

--- 2021-10-12 New Tissue Specific.Rmd
@@
 
-Here we are restricting our protein-coding RNA-seq data to those genes that are labeled as "tissue-specific" on the uniprot database.
-[Seldin method].
-[Definition of tissue-specific on uniprot].
-We will calculate Ssec scores for those labeled as tissue-specific on uniprot, comparing secreted to non-secreted.
-These will be labeled ts uniprot.
+Here we are restricting our protein-coding RNA-seq data to those genes that are labeled as "tissue-specific" on the uniprot database.We will calculate Ssec scores for those labeled as tissue-specific on uniprot, comparing secreted to non-secreted.
+These will be labeled as tissue-specific non-exclusive.
 Additionally, we will calculate these same scores for genes that are labeled tissue-specific AND are only found in the corresponding dataset (brain genes are not found in the liver dataset and vice versa).
-These will be labeled ts uniprot exclusive.
-Seldin: The Ssec rankings were then filtered ensuring that genes were
-primarily expressed by the origin tissue. This was accomplished
-by manual inspection of each gene in multi-tissue expression
-panels available via BioGPS (Su et al., 2002; Wu et al., 2009) in
-both mouse and humans (representative examples using
-BioGPS arrays shown in Figure S3). Genes not detected in
-BioGPS arrays were investigated for tissue specificity using the
-UniProt literature query for ‘‘tissue specificity.’’
-This was done after calculation of Ssec matrices. Manual inspection of top genes
-For this particular test, it does not appear that they focused on tissue-specific genes at all. Solely focused on the categories secreted and nonsecreted.They tested this with QQPlots
-What are BioGPS Arrays?
-Would it be useful at all to look at individuals instead of strain means?
-
-## CHECK INF!!!
-Do I want to try full dataset as target?
+These will be labeled tissue-specific exclusive.
 
@@
 
-The following analysis is comparing the Ssec distribution between secreted and non-secreted genes.
-The Wilcoxon-Mann-Whitney test shows a very significant difference in the liver dataset, though with the very small sample size of 8 genes, we cannot be sure of the accuracy of our calculations.
-The brain dataset continues to not be significantly different between secreted and non-secreted.
-
 ```{r, echo = FALSE}
@@
   geom_density() +
-  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+  labs(title="Brain Ssec Distribution: Tissue-Specific Exclusive Non-Secreted v. Tissue-Specific Exclusive Secreted", x = "Ssec", y = "Density")
 
@@
 
-ggplot(compare.l, aes(val, color = trt)) + 
-  geom_density() +
-  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
-
 #brain
@@
 wil.brain = coin::wilcox_test(val ~ trt, compare.b)
-wil.brain
 
@@
 wil.liver = coin::wilcox_test(val ~ trt, compare.l)
-wil.liver
 
@@
 
+The following analysis is comparing the Ssec distribution between secreted and non-secreted genes that were coded as tissue-specific by Uniprot and determined to be exclusive to brain/liver by direct comparison.
+The Wilcoxon-Mann-Whitney test shows that neither brain nor liver had significant differences between secreted and non-secreted (brain = `r coin::pvalue(wil.brain)`; liver = `r coin::pvalue(wil.liver)`).
+
+```{r, echo = FALSE}
+
+ggplot(compare.l, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: Tissue-Specific Exclusive Non-Secreted v. Tissue-Specific Exclusive Secreted", x = "Ssec", y = "Density")
+
+```
+
 ## Secreted TS Uniprot
-
-The following analysis is comparing the Ssec distribution between secreted and non-secreted genes.
-The Wilcoxon-Mann-Whitney test shows a highly significant difference between secreted and non-secreted for the liver dataset.
-The brain dataset, while not showing a significant p-value, is the closest to significance found in our analyses.
 
@@
 
-## Looking
+The following analysis is comparing the Ssec distribution between secreted and non-secreted genes that were coded as tissue-specific by Uniprot.
+This set excluded the extra step of determining exclusivity, so some genes may be shared between the two groups.
+The Wilcoxon-Mann-Whitney test shows a nearly significant p-value for brain (closest we have gotten so far: `r coin::pvalue(wil.brain)`), but an non-significant p-value for liver (`r coin::pvalue(wil.liver)`).
 
-```{r}
+```{r, echo = FALSE}
+
+ggplot(compare.l, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+
+```
+
+## Evaluating Non-Exclusive Tissue-Specific Datasets
+
+It seems that in the Uniprot database, tissue-specificity does not indicate that a protein is only found in a single tissue, but rather there is SOME KIND of specificity in where it is found.
+For example, the following gene is noted as tissue-specific in Uniprot, but was found in the liver dataset:
+
+ENSRNOG00000020889: not exclusive but tissue specific brain; “Minor constituent of the mineralized matrix of bone.
+Detected in calvaria, rib cartilage, liver, kidney, spleen, brain, lung, skeletal and heart muscle with highest expression in calvaria and approximately half the amount in kidney, liver and brain.”
+
+The following genes were the top genes as selected by the Ssec scores for the brain dataset, but were also found in the liver dataset:
+
+ENSRNOG00000003732: from top_brain; “Detected in brain (at protein level)” 
+ENSRNOG00000023337: from top_brain; “Expressed in the dorsal root ganglia” 
+ENSRNOG00000018962: from top_brain; “Expressed in the ventricle and atrium of adult rats.
+Also detected in the lung, kidney, liver, skeletal muscle, stomach and urinary bladder.
+Not detected in brain, colon, testis, spleen or thymus.
+Overexpressed in the ventricles in the case of hypertension and hypertrophy.” in the ensembl database, they say there is expression in brain with an expression score of 56.67, compared to 77.93 in the liver
+
+The following graph compares average reads for genes that are tissue-specific and NOT exclusive to brain to those same genes in liver.
+If tissue-specific annotation from Uniprot is useful, brain should have more reads for these genes than liver, as shown in the graph.
+The Wilcoxon-Mann_Whitney test affirms this.
+
+```{r, echo = FALSE}
 
@@
   geom_density() +
-  labs(title = "Average Reads of Brain and Liver Samples that are Tissue-Specific but NOT Exclusive to Brain/Liver", x = "Ssec", y = "Density")
+  labs(title = "Average Reads of Brain and Liver Samples that are Tissue-Specific but NOT Exclusive to Brain/Liver", x = "Avg Num Reads", y = "Density")
 
@@
 ```
-It seems that in the uniprot database, tissue-specicificity does not indicate that a protein is only found in a single tissue, rather that there is SOME KIND of specificity in where it is found.
-
-ENSRNOG00000003732: from top_brain; "Detected in brain (at protein level)"
-ENSRNOG00000023337: from top_brain; "Expressed in the dorsal root ganglia"
-ENSRNOG00000018962: from top_brain; "Expressed in the ventricle and atrium of adult rats.
-Also detected in the lung, kidney, liver, skeletal muscle, stomach and urinary bladder.
-Not detected in brain, colon, testis, spleen or thymus.
-Overexpressed in the ventricles in the case of hypertension and hypertrophy." in the ensembl database, they say there is expression in brain with an expression score of 56.67, compared to 77.93 in the liver
-
-ENSRNOG00000002811: from the Secreted TS Uniprot Exclusive Brain dataset; "Expressed in olfactory bulb, cerebellum and neuronal layers in hippocampus."
-
-ENSRNOG00000050761: from the Secreted TS Uniprot Exclusive Liver dataset; "Expressed by the liver and secreted in plasma."
-
-ENSRNOG00000012196: from the Secreted TS Uniprot Liver dataset; "Highly expressed in brain, kidney and heart (PubMed:11328816).
-Expressed at lower level in other tissues such as liver (PubMed:11328816).
-Expressed in intestine, kidney and liver (at protein level) (PubMed:11328816, PubMed:11330410).
-Localizes in the epithelia of the jejunum and ileum (PubMed:11330410)."
-
-ENSRNOG00000020889: not exclusive but tissue specific brain; "Minor constituent of the mineralized matrix of bone.
-Detected in calvaria, rib cartilage, liver, kidney, spleen, brain, lung, skeletal and heart muscle with highest expression in calvaria and approximately half the amount in kidney, liver and brain."

### 2022-05-23 Tissue Exclusive Flow Chart.Rmd
haven't committed in a while pt2

---
title: "Tissue-Exclusive Proteins Report"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Redo the flow chart for tissue exclusive proteins
2. Do a full report of tissue exclusive proteins

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")

protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")

notsec_brain = read.csv("NotSecreted_Brain.csv")
notsec_liver = read.csv("NotSecreted_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)

protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)

notsec_liver = fix.read.csv(notsec_liver)
notsec_brain = fix.read.csv(notsec_brain)

```

```{r}
# Tissue exclusive, secreted proteins
`%!in%` <- Negate(`%in%`)

te_names_brain = colnames(secreted_protein_brain) %!in% colnames(secreted_protein_liver)
te.sec_protein_brain = secreted_protein_brain[, te_names_brain]

te_names_liver = colnames(secreted_protein_liver) %!in% colnames(secreted_protein_brain)
te.sec_protein_liver = secreted_protein_liver[, te_names_liver]

# tissue exclusive, nonsecreted proteins

te_names_brain = colnames(notsec_brain) %!in% colnames(notsec_liver)
te.nsec_protein_brain = notsec_brain[, te_names_brain]

te_names_liver = colnames(notsec_liver) %!in% colnames(notsec_brain)
te.nsec_protein_liver = notsec_liver[, te_names_liver]

```

```{r, echo = FALSE, message = FALSE}
# OPTION 1

# flow chart
# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
library(DiagrammeR)

grViz(diagram = "digraph flowchart { 
      node [fontname = Arial, shape = oval, color = Lavender, style = filled, cex = 40]
      tab1 [label = '@@1', fontsize = 100]
      tab2 [label = '@@2', fontsize = 100]
      tab3 [label = '@@3', fontsize = 100]
      tab4 [label = '@@4', fontsize = 100]
      tab5 [label = '@@5', fontsize = 100]
      tab6 [label = '@@6', fontsize = 100]
      tab7 [label = '@@7', fontsize = 100]
      tab8 [label = '@@8', fontsize = 100]

      edge[minlen = 4, penwidth = 8, arrowsize = 4]
      tab1 -> tab2;
      tab2 -> tab3;
      tab3 -> tab4;
      tab4 -> tab5;
      tab5 -> tab6;
      tab4 -> tab7;
      tab7 -> tab8}

      [1]: 'Full Cleaned Data: B(132 individuals, 18,385 genes), L(132 individuals, 15,365 genes)'
      [2]: 'Means of Strains: B(45 strains, 18,385 genes), L(45 strains, 15,365 genes)'
      [3]: 'RI Strains Only: B(30, 18,385), L(30, 15,365)'
      [4]: 'Protein-Coding Genes Only: B(30, 16,457), L(30, 14,109)'
      [5]: 'Secreted Protein-Coding Genes: B(30, 657), L(30, 818)'
      [6]: 'Tissue-Exclusive Secreted Protein-Coding Genes: B(30, 174), L(30, 35)'
      [7]: 'Non-Secreted Protein-Coding Genes: B(30, 15,800), L(30, 13,591)'
      [8]: 'Tissue-Exclusive Non-Secreted Protein-Coding Genes: B(30, 2,589), L(30, 371)' 
      ", width = 1000, height = 400)

```

Labels are of the format Tissue dataset (number of strains, number of genes), unless otherwise specified. B indicates the brain dataset, while L indicates the liver dataset. The full cleaned data came in the form number of RNA reads per individual rat. The strain means of the reads were calculated by averaging the RNA reads per gene for all rats within a strain. The resulting dataset had data for 45 strains. Classically inbred strains were removed from the dataset, leaving only recombinant inbred strains, for which we have phenotype data. Then, all non-protein coding-genes were removed, as identified by _.   MORE

```{r}

# OPTION 2

# flow chart
# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
library(DiagrammeR)

grViz(diagram = "digraph flowchart { 
      node [fontname = Arial, shape = oval, color = Lavender, style = filled]
      tab1 [label = '@@1', fontsize = 100]
      tab2 [label = '@@2', fontsize = 100]
      tab3 [label = '@@3', fontsize = 100]
      tab4 [label = '@@4', fontsize = 100]
      tab5 [label = '@@5', fontsize = 100]
      tab6 [label = '@@6', fontsize = 100]

      edge[minlen = 4, penwidth = 8, arrowsize = 4]
      tab1 -> tab2;
      tab2 -> tab3;
      tab3 -> tab4;
      tab2 -> tab5;
      tab5 -> tab6;
}

      [1]: 'All RNA Reads: B(18,385 genes), L(15,365 genes)'
      [2]: 'Protein-Coding Genes Only: B(16,457), L(14,109)'
      [3]: 'Secreted Protein-Coding Genes: B(657), L(818)'
      [4]: 'Tissue-Exclusive Secreted Protein-Coding Genes: B(174), L(35)'
      [5]: 'Non-Secreted Protein-Coding Genes: B(15,800), L(13,591)'
      [6]: 'Tissue-Exclusive Non-Secreted Protein-Coding Genes: B(2,589), L(371)' 
      ", width = 1000, height = 400)

```

```{r}

library(Gmisc)
library(glue)
library(htmlTable)
library(grid)
library(magrittr)
```

```{r}

# https://cran.r-project.org/web/packages/Gmisc/vignettes/Grid-based_flowcharts.html

grid.newpage()
txt1 <-
"Full Cleaned Data
Brain (132 ind, 18,385 genes)
Liver (132 ind, 15,365 genes)"
boxFC = boxGrob(txt1, x = 0.15, y = .89)
txt2 = 
  "Means of Strains
Brain (45 strains, 18,385 genes)
Liver (45 strains, 15,365 genes)"
boxMS = boxGrob(txt2, x = 0.15, y = 0.65)
txt3 = 
  "RI Strains Only
Brain (30 strains, 18,385 genes)
Liver (30 strains, 15,365 genes)"
boxRI = boxGrob(txt3, x = 0.15, y = 0.40)
txt4 = 
  "Protein-Coding Genes Only
Brain (30 strains, 16,457 genes)
Liver (30 strains, 14,109 genes)"
boxPC = boxGrob(txt4, x = 0.15, y = 0.15)
txt5 = 
  "Secreted Protein-Coding Genes
Brain (30 strains, 657 genes)
Liver (30 strains, 818 genes)"
boxSe = boxGrob(txt5, x = 0.5, y = 0.89)
txt6 = 
  "Tissue-Exclusive Secreted 
  Protein-Coding Genes
Brain (30 strains, 174 genes)
Liver (30 strains, 35 genes)"
boxTES = boxGrob(txt6, x = 0.5, y = 0.65)
txt7 = 
  "Non-Secreted Protein-Coding Genes
Brain (30 strains, 15,800 genes)
Liver (30 strains, 13,591 genes)"
boxNSe = boxGrob(txt7, x = 0.5, y = 0.40)
txt8 = 
  "Tissue-Exclusive Non-Secreted 
  Protein-Coding Genes
Brain (30 strains, 2,589 genes)
Liver (30 strains, 371 genes)"
boxTENS = boxGrob(txt8, x = 0.5, y = 0.13)

connectGrob(boxFC, boxMS, "vertical")
connectGrob(boxMS, boxRI, "vertical")
connectGrob(boxRI, boxPC, "vertical")
connectGrob(boxPC, boxSe, "Z", "r")
connectGrob(boxSe, boxTES, "vertical")
connectGrob(boxPC, boxNSe, "Z", "r")
connectGrob(boxNSe, boxTENS, "vertical")

boxFC
boxMS
boxRI
boxPC
boxSe
boxTES
boxNSe
boxTENS

```

```{r}

grid.newpage()
txt1 =
"All Normalized Expression Values
Brain (18,385 genes)
Liver (15,365 genes)"
boxAR = boxGrob(txt1, x = 0.15, y = 0.5)
txt4 = 
  "Protein-Coding Genes Only
Brain (16,457 genes)
Liver (14,109 genes)"
boxPC = boxGrob(txt4, x = 0.35, y = 0.5)
txt5 = 
  "Secreted Protein-Coding Genes
Brain (657 genes)
Liver (818 genes)"
boxSe = boxGrob(txt5, x = 0.35, y = 0.70)
txt6 = 
  "Tissue-Exclusive, Secreted 
  Protein-Coding Genes
Brain (174 genes)
Liver (35 genes)"
boxTES = boxGrob(txt6, x = 0.55, y = 0.70)
txt7 = 
  "Non-Secreted Protein-Coding Genes
Brain (15,800 genes)
Liver (13,591 genes)"
boxNSe = boxGrob(txt7, x = 0.35, y = 0.30)
txt8 = 
  "Tissue-Exclusive, Non-Secreted 
  Protein-Coding Genes
Brain (2,589 genes)
Liver (371 genes)"
boxTENS = boxGrob(txt8, x = 0.55, y = 0.30)
txt9 = 
  "Tissue-Exclusive 
  Protein-Coding Genes
Brain (2,751 genes)
Liver (405 genes)"
boxTE = boxGrob(txt8, x = 0.55, y = 0.5)

connectGrob(boxAR, boxPC, "horizontal")
connectGrob(boxPC, boxSe, "vertical")
connectGrob(boxSe, boxTES, "horizontal")
connectGrob(boxPC, boxNSe, "vertical")
connectGrob(boxNSe, boxTENS, "horizontal")
connectGrob(boxPC, boxTE, "horizontal")

boxAR
boxPC
boxSe
boxTES
boxNSe
boxTENS
boxTE

```

### 2022-05-23 Server Code.R

@@ -1,0 +1,70 @@
+# Code for server
+
+# protein_brain, protein_liver, secreted_protein_brain, secreted_protein_liver
+# notsecreted_protein_brain, notsecreted_protein_liver
+
+### Ssec calculations
+
+target_liver = protein_liver
+
+brain.liv = bicorAndPvalue(secreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv$p))
+
+sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+
+write.csv(sec_Ssec_brain, "Sec-Full-Ssec_brain.csv")
+## LIVER ##
+
+target_brain = protein_brain
+
+liv.brain = bicorAndPvalue(secreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(liv.brain$p))
+
+sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_brain))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+write.csv(sec_Ssec_liver, "Sec-Full-Ssec_liver.csv")
+
+#### Not Secreted ####
+
+## BRAIN ##
+
+brain.liv.c = bicorAndPvalue(notsecreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+# brain.liv$p
+
+scores = rowSums(-log(brain.liv.c$p))
+
+nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
+  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  dplyr::select(-score) %>%
+  arrange(desc(Ssec))
+
+write.csv(nsec_Ssec_brain, "NSec-Full-Ssec_brain.csv")
+
+## LIVER ##
+
[…]

### 2022-05-23 EnrichR.Rmd
initial commit figuring out enrichr

---
title: "EnrichR Report"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Determine if Enrichr is a valid replacement for EGSEA for gene set analysis
2. If yes, summarize results of Enrichr

```{r, echo = FALSE, message = FALSE}

# https://cran.r-project.org/web/packages/enrichR/vignettes/enrichR.html

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to gene names

```

```{r, echo = FALSE}
library(enrichR)

setEnrichrSite("Enrichr") # human genes???
dbs = c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")

enriched = enrichr(sec_Ssec_brain[1:100,], dbs)

enriched

```

MSigDB
https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html
https://bioinformaticsbreakdown.com/how-to-gsea/

### 2022-05-24 Compare SecBrain to TESBrain.Rmd
consolidation

---
title: "Compare Heritability for Secreted and Tissue-Exclusive, Secreted Protein-Coding Genes in the Brain"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Graph Sec and TES brain heritability
2. Look at genes lost between Sec and TES with high heritability

```{r, echo = FALSE, message = FALSE}
# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
te.secreted_protein_brain = read.csv("TE.Secreted_Protein_Brain.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
te.secreted_protein_brain = fix.read.csv(te.secreted_protein_brain)

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

### SUBSET

names = substr(fin.r.squared_brain[,1], 1, 18)
fin.r.squared_brain[,1] = names

sec_herit = fin.r.squared_brain[fin.r.squared_brain[,1] %in% colnames(secreted_protein_brain),]
tesec_herit = fin.r.squared_brain[fin.r.squared_brain[,1] %in% colnames(te.secreted_protein_brain),]

`%!in%` <- Negate(`%in%`)

ntesec = sec_herit[,1][sec_herit[,1] %!in% tesec_herit[,1]]
ntesec_herit = fin.r.squared_brain[fin.r.squared_brain[,1] %in% ntesec,]

# sec v. tesec

gee = cbind(rep("secreted", times = dim(sec_herit)[1]), sec_herit[,2])
gee2 = cbind(rep("te secreted", times = dim(tesec_herit)[1]), tesec_herit[,2])
gee3 = cbind(rep("nte secreted", times = dim(ntesec_herit)[1]), ntesec_herit[,2])

wil.b = rbind(gee, gee2, gee3)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("trt", "r.sq")

wil.b$trt = as.factor(wil.b$trt)
wil.b$r.sq = as.numeric(wil.b$r.sq)

wil.b.1 = wil.b[wil.b[,1] == "te secreted" | wil.b[,1] == "nte secreted",]

wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b.1)

ngenes = c(dim(sec_herit)[1], dim(tesec_herit)[1], dim(ntesec_herit)[1])
names(ngenes) = c("Secreted", "Tissue-Exclusive Secreted", "Non-Tissue-Exclusive Secreted")

knitr::kable(ngenes, caption = "Number of Genes in each subset for Brain")

```

The density graph is comparing secreted protein-coding genes in the brain (secreted, green) to tissue-exclusive, secreted protein-coding genes in the brain (te secreted, blue) to non-tissue-exclusive, secreted protein-coding genes in the brain (nte secreted, red).

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Density R Squared
library(tidyverse)
ggplot(wil.b, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")

```

The histogram plot is comparing secreted protein-coding genes in the brain to tissue-exclusive, secreted protein-coding genes in the brain. It seems that the genes excluded in the tissue-exclusive set have higher heritability than the tissue-exclusive genes.

```{r, echo = FALSE}

c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

hgA <- hist(sec_herit[,2], plot = FALSE) # Save first histogram data
hgB <- hist(tesec_herit[,2], plot = FALSE) # Save 2nd histogram data

plot(hgA, col = c1, main = "Comparing Secreted to TE Secreted Brain", xlab = "R Squared") # Plot 1st histogram using a transparent color
plot(hgB, col = c2, add = TRUE) # Add 2nd histogram using different color

```

### 2022-05-24 SeldinReport.Rmd
consolidation

@@
 brain.liv = bicorAndPvalue(ts_protein_brain, target_liver, use = 'pairwise.complete.obs')
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+write.csv(brain.liv$p, "Bicor_brain.csv")
 # brain.liv$p
@@
 # brain.liv$p
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+write.csv(liv.brain$p, "Bicor_liver.csv")
 

### 2022-05-24 Partial Correlation Ssec Values.Rmd
just going in circles at this point

---
title: "Ssec Scores: Partial Correlation"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Compare theory behind partial midweight correlation and partial correlation
2. Calculate Partial Correlation, compare to biweight midcorrelation
3. Convert Partial Correlation to Ssec Scores
4. Look at distribution of Partial Correlation, QQ plots
5. Use permutation test from Seldin to check null distribution

Problem with Keller- no code, no explanation of how they computed PaCor analytically

Problem with partial correlation methods- don't like that there are more genes than subjects

```{r, echo = FALSE, message = FALSE}
library(WGCNA)
library(ppcor)
library(tidyverse)

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tes.brain = read.csv("TissueSpecific2_Protein_Brain.csv")
tes.liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tens.brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tens.liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

tes.brain = fix.read.csv(tes.brain)
tes.liver = fix.read.csv(tes.liver)
tens.brain = fix.read.csv(tens.brain)
tens.liver = fix.read.csv(tens.liver)

```
Wow what a cluster fuck
```{r, echo = FALSE, eval = FALSE}

# PaCor v. BiCor
# pcor(target, tes) - pull out tes genes?
target_liver = cbind(tes.liver, tens.liver)
pacor.brain_data = cbind(tes.brain, target_liver)
pacor.brain = pcor(pacor.brain_data, method = "spearman")

#########
library(psych)

pacor.brain = partial.r(pacor.brain_data, method = "spearman")

p.pacor.brain = corr.p(pacor.brain, n = 577, ci = FALSE)$p

fin.pacor.brain = p.pacor.brain[rownames(p.pacor.brain) %in% colnames(tes.brain), colnames(p.pacor.brain) %in% colnames(target_liver)]

scores = rowSums(-log(fin.pacor.brain))

pacor_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

###
sum = c()
for (i in colnames(tes.brain))
{
  pacor.brain = partial.r(cbind(tes.brain[,i], target_liver), method = "spearman")

  p.pacor.brain = corr.p(pacor.brain, n = 404, ci = FALSE)$p

  sum = c(sum, sum(p.pacor.brain[1,]))
}
```

These don't like that there are more genes than subjects :(

partial.r {psych}	R Documentation
Find the partial correlations for a set (x) of variables with set (y) removed.
Description
A straightforward application of matrix algebra to remove the effect of the variables in the y set from the x set. Input may be either a data matrix or a correlation matrix. Variables in x and y are specified by location. If x and y are not specified, then the effect of all variables are partialled from all the other correlations. May also be done using formula input which is more convenient when comparing results to regression models

Usage
partial.r(data, x, y, use="pairwise",method="pearson")
Arguments
data	
A data or correlation matrix

x	
The variable names or locations associated with the X set (or formula input)

y	
The variable names or locations associated with the Y set to be partialled from the X set

use	
How should we treat missing data? The default is pairwise complete.

Warning in cor.smooth(R) :
  Matrix was not positive definite, smoothing was done
In smc, smcs < 0 were set to .0

####

pcor {ppcor}	R Documentation
Partial correlation
Description
The function pcor can calculate the pairwise partial correlations for each pair of variables given others. In addition, it gives us the p value as well as statistic for each pair of variables.

Usage
pcor(x, method = c("pearson", "kendall", "spearman"))
Arguments
x	
a matrix or data fram.

method	
a character string indicating which partial correlation coefficient is to be computed. One of "pearson" (default), "kendall", or "spearman" can be abbreviated.

```{r, echo = FALSE}
library(GeneNet)

estimate = ggm.estimate.pcor(pacor.brain_data, method = "static", lambda = 0)
edge = network.test.edges(estimate)

n2kappa(30, p = 577)

pval = cor0.test(estimate, kappa = 29)

kappa2n(29, p = 2)

brain.liv = bicorAndPvalue(pacor.brain_data, pacor.brain_data, use = 'pairwise.complete.obs')

cor = cor2pcor(brain.liv$bicor)
cor2 = pcor.shrink(pacor.brain_data)

```

### 2022-05-24 egsea.Rmd
just going in circles at this point

---
title: "EGSEA Analysis"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Use the package EGSEA to find potential biological functions for gene sets
a. Import sets of target genes
b. Run EGSEA analysis

```{r, echo = FALSE}

library(EGSEA)

# load data

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

ts_Ssec_target_brain = read.csv("Target_TESec_Ssec_brain.csv")
ts_Ssec_target_liver = read.csv("Target_TESec_Ssec_liver.csv")
tsns_Ssec_target_brain = read.csv("Target_TENSec_Ssec_brain.csv")
tsns_Ssec_target_liver = read.csv("Target_TENSec_Ssec_liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")
bicor_brain = read.csv( "Bicor_brain.csv")
bicor_liver = read.csv( "Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

ts_Ssec_target_brain = fix.read.csv(ts_Ssec_target_brain)
ts_Ssec_target_liver = fix.read.csv(ts_Ssec_target_liver)
tsns_Ssec_target_brain = fix.read.csv(tsns_Ssec_target_brain)
tsns_Ssec_target_liver = fix.read.csv(tsns_Ssec_target_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

```

```{r, echo = FALSE}

library(EGSEAdata)

egsea.data("mouse")

```

```{r, echo = FALSE}

# top 4 Ssec values

top_brain = rownames(ts_Ssec_target_brain)[1:5]
top_liver = rownames(ts_Ssec_target_liver)[1:5]

corr_brain = matrix(NA, nrow = 5, ncol = 403)
corr_brain = as.data.frame(corr_brain)

for(i in 1:5)
{
  one = bicor_brain[top_brain[i] == rownames(bicor_brain),]
  corr_brain[i,] = one
}
rownames(corr_brain) = top_brain
colnames(corr_brain) = colnames(bicor_brain)

corr_liver = matrix(NA, nrow = 5, ncol = 2763)
corr_liver = as.data.frame(corr_liver)

for(i in 1:5)
{
  one = bicor_liver[top_liver[i] == rownames(bicor_liver),]
  corr_liver[i,] = one
}

rownames(corr_liver) = top_liver
colnames(corr_liver) = colnames(bicor_liver)

target_forbrain = corr_brain < 0.05
target_forliver = corr_liver < 0.05

num_target_brain = apply(target_forbrain, 1, sum)
num_target_liver = apply(target_forliver, 1, sum)

knitr::kable(num_target_brain, caption = "Number of Liver Target Genes for Each Top Ssec Brain")
knitr::kable(num_target_liver, caption = "Number of Brain Target Genes for Each Top Ssec Liver")

```

WHAT THE FUCK

```{r, echo = FALSE}
# get entrez gene ids

library(biomaRt)

ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")

get_attrib <- listAttributes(ensembl)
go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype"),
                            #"entrezgene_trans_name",
                            #"entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_tes.brain = get_entrez$entrezgene_id[get_entrez$ensembl_gene_id %in% rownames(ts_Ssec_target_brain)]

sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_tes.liver = get_entrez$entrezgene_id[get_entrez$ensembl_gene_id %in% rownames(ts_Ssec_target_liver)]

```

### 2022-06-20 EnrichR.Rmd
happy pride month

@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
+tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
+tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
+tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
+bicor_brain = read.csv("Bicor_brain.csv")
+bicor_liver = read.csv("Bicor_liver.csv")
+
+fix.read.csv = function(x)
+{
+  names = x[,1]
+  gee = x[,2:dim(x)[2]]
+  rownames(gee) = names
+  return(gee)
+}
+
+tesec_brain = fix.read.csv(tesec_brain)
+tesec_liver = fix.read.csv(tesec_liver)
+tensec_brain = fix.read.csv(tensec_brain)
+tensec_liver = fix.read.csv(tensec_liver)
+bicor_brain = fix.read.csv(bicor_brain)
+bicor_liver = fix.read.csv(bicor_liver)
 
@@
 
-setEnrichrSite("Enrichr") # human genes???
-dbs = c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")
-
-enriched = enrichr(sec_Ssec_brain[1:100,], dbs)
+#setEnrichrSite("Enrichr") # human genes???
+#dbs = c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")
+dbs = c("MSigDB_Computational", "MSigDB_Hallmark_2020", "MSigDB_Oncogenic_Signatures")
+fin = c()
+for(i in 1:174)
+{
+  names = colnames(bicor_brain[,bicor_brain[i,] < 0.05])
+  enriched = enrichr(tesec_brain[, 1], dbs)
+  if(dim(enriched$MSigDB_Computational)[1] != 0 | dim(enriched$MSigDB_Hallmark_2020)[1] != 0 | dim(enriched$MSigDB_Oncogenic_Signatures)[1] != 0 )
+  {
+    fin = c(fin, i)
+  }
+}
 

### 2022-06-20 MSigDB-SOP-2210-3.Rmd
happy pride month

---
title: "Gene Enrichment Analysis with MSigDB"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1.  Identify rat gene sets
2.  Convert Ensembl Gene IDs into Gene Symbol
3.  Use clusterProfiler to do pathway analysis

https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html

```{r, echo = FALSE}
library(msigdbr)

all_genes = msigdbr(species = "Rattus norvegicus")

```

```{r, echo = FALSE}

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
bicor_brain = read.csv("Bicor_brain.csv")
bicor_liver = read.csv("Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}
`%!in%` = Negate(`%in%`)

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to gene names

```

Fuck this shit

```{r, echo = FALSE}

library(biomaRt)
ensembl_version = "https://ensembl.org"

ensembl = useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "rnorvegicus_gene_ensembl", version = 109)
ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")

get_attrib <- listAttributes(ensembl)
#go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            #"entrezgene_trans_name",
                            "entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_tes.brain = get_entrez[get_entrez$ensembl_gene_id %in% rownames(tesec_brain),]

#sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
#GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_tes.liver = get_entrez[get_entrez$ensembl_gene_id %in% rownames(tesec_liver),]

# missing gene ids

missing_b = rownames(tesec_brain)[rownames(tesec_brain) %!in% entrez_tes.brain$ensembl_gene_id]

missing_l = rownames(tesec_liver)[rownames(tesec_liver) %!in% entrez_tes.liver$ensembl_gene_id]

```

### 2022-06-20 MSigDB.Rmd
happy pride month

---
title: "Gene Enrichment Analysis with MSigDB"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1.  Identify rat gene sets
2.  Convert Ensembl Gene IDs into Entrez Ids
3.  Report on missingness in Biomart
3.  Use clusterProfiler to do pathway analysis

For our gene set enrichment analysis (GSEA), we will be using the data resource MSigDB (Molecular Signatures Database), "a collection of annotated gene sets for use with GSEA software." This dataset was used in conjunction with the GSEA R package clusterProfiler. 
We identified our enrichment dataset via the function msigdbr from the msigdbr r package. The species was identified as "Rattus norvegicus." msigdbr has access to multiple collections, including hallmark gene sets (H), and eight different supplementary gene sets (C1 to C8). In this analysis, we did not exclude any of the collections.

```{r, echo = FALSE}
library(msigdbr)

all_genes = msigdbr(species = "Rattus norvegicus")

# https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html
# https://bioconductor.org/packages/release/bioc/manuals/clusterProfiler/man/clusterProfiler.pdf

```

```{r, echo = FALSE}

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
bicor_brain = read.csv("Bicor_brain.csv")
bicor_liver = read.csv("Bicor_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}
`%!in%` = Negate(`%in%`)

tesec_brain = fix.read.csv(tesec_brain)
tesec_liver = fix.read.csv(tesec_liver)
tensec_brain = fix.read.csv(tensec_brain)
tensec_liver = fix.read.csv(tensec_liver)
bicor_brain = fix.read.csv(bicor_brain)
bicor_liver = fix.read.csv(bicor_liver)

# top 5 TES genes for brain and liver
# the genes theyre correlated to
# convert to entrez ids

# top 5 genes for brain and liver

top.brain = tesec_brain[1:5,]
top.liver = tesec_liver[1:5,]

```
We began by identifying the top five brain and liver tissue-exclusive, secreted (TES) protein-coding genes, as identified by Ssec score. We then identified the target tissue transcripts that were most significantly associated with the top TES protein-coding genes (p-value < 0.05). These will be the gene sets for which we will identify annotations. This results in 138 liver genes that are associated with TES protein-coding genes from the brain, and 935 brain genes that are associated with TES protein-coding genes from the liver.

```{r, echo = FALSE}

# genes they are correlated to < 0.05

top.corr.brain = bicor_brain[rownames(bicor_brain) %in% rownames(top.brain),]
top.corr.liver = bicor_liver[rownames(bicor_liver) %in% rownames(top.liver),]

list.oflists = function(mat)
{
  out = list(c(), c(), c(), c(), c())
  for (i in 1:5)
  {
    save = colnames(mat[,mat[i,] < 0.05])
    out[[i]] = save
  }
  return(out)
}

brain.list = list.oflists(top.corr.brain)
liver.list = list.oflists(top.corr.liver)

combine.brain = unique(c(brain.list[[1]], brain.list[[2]], brain.list[[3]], brain.list[[4]], brain.list[[5]]))

combine.liver = unique(c(liver.list[[1]], liver.list[[2]], liver.list[[3]], liver.list[[4]], liver.list[[5]]))

num.brain = c(length(brain.list[[1]]), length(brain.list[[2]]), length(brain.list[[3]]), length(brain.list[[4]]), length(brain.list[[5]]))

num.liver = c(length(liver.list[[1]]), length(liver.list[[2]]), length(liver.list[[3]]), length(liver.list[[4]]), length(liver.list[[5]]))

num.target.brain = data.frame(top.brain[,1], num.brain)
num.target.liver = data.frame(top.liver[,1], num.liver)
colnames(num.target.brain) = c("TES PC from Brain", "Number of Target Genes in Liver")
colnames(num.target.liver) = c("TES PC from Liver", "Number of Target Genes in Brain")

knitr::kable(num.target.brain, caption = "Number of Target Genes in the Liver associated with TES Protein-Coding Genes from the Brain")

knitr::kable(num.target.liver, caption = "Number of Target Genes in the Brain associated with TES Protein-Coding Genes from the Liver")

```

Entrez gene IDs were identified via biomart version 104.

## Missingness Report

```{r, echo = FALSE, eval = FALSE}

library(biomaRt)
#ensembl_version = "https://ensembl.org"

#ensembl = useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "rnorvegicus_gene_ensembl", version = 109)
ensembl.miss = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl", version = 104)

get_attrib <- listAttributes(ensembl)
#go_attrib <- get_attrib[grep("GO ",get_attrib$description),]

# getting entrez ids
get_entrez = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            #"entrezgene_trans_name",
                            "entrezgene_id"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

get_uniprot = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype",
                            "uniprot_gn_id",
                            "name_1006",
                            "definition_1006","go_linkage_type","namespace_1003"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

entrez_brain.1= get_entrez[get_entrez$ensembl_gene_id %in% combine.brain,]

#sum(get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))
#GO_fin = subset(get_uniprot, get_uniprot$ensembl_gene_id %in% rownames(ts_Ssec_target_brain))

entrez_liver.1 = get_entrez[get_entrez$ensembl_gene_id %in% combine.liver,]

# missing gene ids

missing_b = combine.brain[combine.brain %!in% entrez_brain$ensembl_gene_id]
# 15 missing TES : "ENSRNOG00000047739" "ENSRNOG00000008563" "ENSRNOG00000009140" "ENSRNOG00000038437" "ENSRNOG00000034130" "ENSRNOG00000003887" "ENSRNOG00000007374" "ENSRNOG00000013441" "ENSRNOG00000003891" "ENSRNOG00000038980" "ENSRNOG00000013228" "ENSRNOG00000038150" "ENSRNOG00000049171" "ENSRNOG00000011074" "ENSRNOG00000050172"

# 28 missing target : "ENSRNOG00000003498" "ENSRNOG00000047079" "ENSRNOG00000053527" "ENSRNOG00000057548" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000049727" "ENSRNOG00000050499" "ENSRNOG00000050885" "ENSRNOG00000051227" "ENSRNOG00000052192" "ENSRNOG00000054524" "ENSRNOG00000061059" "ENSRNOG00000020805" "ENSRNOG00000003648" "ENSRNOG00000047785" "ENSRNOG00000056194" "ENSRNOG00000013410" "ENSRNOG00000049009" "ENSRNOG00000057089" "ENSRNOG00000059007" "ENSRNOG00000061217" "ENSRNOG00000003054" "ENSRNOG00000042197" "ENSRNOG00000047520" "ENSRNOG00000047821" "ENSRNOG00000048017"

missing_l = combine.liver[combine.liver %!in% entrez_liver$ensembl_gene_id]
# 12 missing TES : "ENSRNOG00000020805" "ENSRNOG00000050499" "ENSRNOG00000049727" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000013464" "ENSRNOG00000049515" "ENSRNOG00000013410" "ENSRNOG00000016535" "ENSRNOG00000033619" "ENSRNOG00000033245"

# 116 missing target : "ENSRNOG00000038437" "ENSRNOG00000001057" "ENSRNOG00000006591" "ENSRNOG00000007705" "ENSRNOG00000008475" "ENSRNOG00000009241" "ENSRNOG00000019857" "ENSRNOG00000021533" "ENSRNOG00000028576" "ENSRNOG00000038976" "ENSRNOG00000043358" "ENSRNOG00000045720" "ENSRNOG00000045960" "ENSRNOG00000046838" "ENSRNOG00000049878" "ENSRNOG00000052459" "ENSRNOG00000053258" "ENSRNOG00000055404" "ENSRNOG00000056130" "ENSRNOG00000057009" "ENSRNOG00000057250" "ENSRNOG00000057522" "ENSRNOG00000057724" "ENSRNOG00000061051" "ENSRNOG00000009140" "ENSRNOG00000005665" "ENSRNOG00000006226" "ENSRNOG00000007219" "ENSRNOG00000008544" "ENSRNOG00000009835" "ENSRNOG00000010720" "ENSRNOG00000011640" "ENSRNOG00000014847" "ENSRNOG00000024726" "ENSRNOG00000028279" "ENSRNOG00000042321" "ENSRNOG00000042502" "ENSRNOG00000042971" "ENSRNOG00000042985" "ENSRNOG00000043382" "ENSRNOG00000045545" "ENSRNOG00000045767" "ENSRNOG00000045837" "ENSRNOG00000046212" "ENSRNOG00000046272" "ENSRNOG00000047890" "ENSRNOG00000048172" "ENSRNOG00000048322" "ENSRNOG00000048765" "ENSRNOG00000050134" "ENSRNOG00000054121" "ENSRNOG00000055892" "ENSRNOG00000056186" "ENSRNOG00000058047" "ENSRNOG00000058276" "ENSRNOG00000058766" "ENSRNOG00000058984" "ENSRNOG00000059734" "ENSRNOG00000060719" "ENSRNOG00000062048" "ENSRNOG00000002953" "ENSRNOG00000022107" "ENSRNOG00000042211" "ENSRNOG00000047780" "ENSRNOG00000061301" "ENSRNOG00000038150" "ENSRNOG00000038980" "ENSRNOG00000003395" "ENSRNOG00000006051" "ENSRNOG00000009917" "ENSRNOG00000010749" "ENSRNOG00000011856" "ENSRNOG00000012477" "ENSRNOG00000019958" "ENSRNOG00000021183" "ENSRNOG00000022919" "ENSRNOG00000027929" "ENSRNOG00000029837" "ENSRNOG00000031734" "ENSRNOG00000032206" "ENSRNOG00000033894" "ENSRNOG00000034287" "ENSRNOG00000037852" "ENSRNOG00000038184" "ENSRNOG00000039050" "ENSRNOG00000042516" "ENSRNOG00000042939" "ENSRNOG00000046844" "ENSRNOG00000047864" "ENSRNOG00000048722" "ENSRNOG00000049435" "ENSRNOG00000050378" "ENSRNOG00000056555" "ENSRNOG00000057059" "ENSRNOG00000057228" "ENSRNOG00000059539" "ENSRNOG00000060482" "ENSRNOG00000060898" "ENSRNOG00000062013" "ENSRNOG00000000060" "ENSRNOG00000005114" "ENSRNOG00000006320" "ENSRNOG00000009388" "ENSRNOG00000009635" "ENSRNOG00000020821" "ENSRNOG00000021302" "ENSRNOG00000023404" "ENSRNOG00000024678" "ENSRNOG00000046157" "ENSRNOG00000047724" "ENSRNOG00000049802" "ENSRNOG00000050511" "ENSRNOG00000057758" "ENSRNOG00000057852" "ENSRNOG00000058416" "ENSRNOG00000060181"

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(entrez_brain.1, "EntrezIds_brain.csv")
write.csv(entrez_liver.1, "EntrezIds_liver.csv")

```

```{r, echo = FALSE, eval = FALSE}

# duplicates
adding = c()
dup = c()
for (i in entrez_tes.brain$ensembl_gene_id)
{
  if (i %in% adding)
  {
    dup = c(dup, i)
  }
  else
  {
    adding = c(adding, i)
  }

}

dup.entrez = entrez_tes.brain$entrezgene_id[entrez_tes.brain$ensembl_gene_id %in% dup]
# there are 4 NAs in the brain dataset
# there are 3 ensembl duplicates in brain dataset
# there are 5 ensembl duplicates in liver dataset

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
entrez_brain = read.csv("EntrezIds_brain.csv")
entrez_liver = read.csv("EntrezIds_liver.csv")

entrez_brain = fix.read.csv(entrez_brain)
entrez_liver = fix.read.csv(entrez_liver)

#msigdbr_t2g = msigdbr_df %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
#enricher(gene = gene_ids_vector, TERM2GENE = msigdbr_t2g, ...)

msigdbr_t2g = all_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()

library(clusterProfiler)

# pAdjustMethod, min/max (number?) of genes

#enricher(gene = target_genes, TERM2GENE = msigdbr_t2g, ...)

# brain.list, liver.list

out.brain = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = brain.list[[i]]
  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
  out.brain[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

out.liver = list(c(), c(), c(), c(), c())

for (i in 1:5)
{
  vect = liver.list[[i]]
  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
  entrez = na.omit(entrez)
  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
  out.liver[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
}

brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
names(brain.ids) = top.brain[,1]

knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")

liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
names(liver.ids) = top.liver[,1]

knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")

```

### 2022-06-20 Partial Correlation Ssec Values.Rmd
happy pride month

@@
 
-Problem with Keller- no code, no explanation of how they computed PaCor analytically
-
-Problem with partial correlation methods- don't like that there are more genes than subjects
+Lambda???
 
@@
 
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
@@
 ```
-Wow what a cluster fuck
-```{r, echo = FALSE, eval = FALSE}
-
-# PaCor v. BiCor
-# pcor(target, tes) - pull out tes genes?
-target_liver = cbind(tes.liver, tens.liver)
-pacor.brain_data = cbind(tes.brain, target_liver)
-pacor.brain = pcor(pacor.brain_data, method = "spearman")
-
-#########
-library(psych)
-
-pacor.brain = partial.r(pacor.brain_data, method = "spearman")
-
-p.pacor.brain = corr.p(pacor.brain, n = 577, ci = FALSE)$p
-
-fin.pacor.brain = p.pacor.brain[rownames(p.pacor.brain) %in% colnames(tes.brain), colnames(p.pacor.brain) %in% colnames(target_liver)]
-
-scores = rowSums(-log(fin.pacor.brain))
-
-pacor_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
-  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(target_liver))) %>%
-  dplyr::select(-score) %>%
-  arrange(desc(Ssec))
-
-###
-sum = c()
-for (i in colnames(tes.brain))
-{
-  pacor.brain = partial.r(cbind(tes.brain[,i], target_liver), method = "spearman")
-
-  p.pacor.brain = corr.p(pacor.brain, n = 404, ci = FALSE)$p
-
-  sum = c(sum, sum(p.pacor.brain[1,]))
-}
-```
-
-These don't like that there are more genes than subjects :(
-
-partial.r {psych}	R Documentation
-Find the partial correlations for a set (x) of variables with set (y) removed.
-Description
-A straightforward application of matrix algebra to remove the effect of the variables in the y set from the x set.
-Input may be either a data matrix or a correlation matrix.
-Variables in x and y are specified by location.
-If x and y are not specified, then the effect of all variables are partialled from all the other correlations.
-May also be done using formula input which is more convenient when comparing results to regression models
-
-Usage
-partial.r(data, x, y, use="pairwise",method="pearson")
-Arguments
-data	
-A data or correlation matrix
-
-x	
-The variable names or locations associated with the X set (or formula input)
-
-y	
-The variable names or locations associated with the Y set to be partialled from the X set
-
-use	
-How should we treat missing data? The default is pairwise complete.
-
-Warning in cor.smooth(R) :
-  Matrix was not positive definite, smoothing was done
-In smc, smcs < 0 were set to .0
-
-####
-
-pcor {ppcor}	R Documentation
-Partial correlation
-Description
-The function pcor can calculate the pairwise partial correlations for each pair of variables given others. In addition, it gives us the p value as well as statistic for each pair of variables.
-
-Usage
-pcor(x, method = c("pearson", "kendall", "spearman"))
-Arguments
-x	
-a matrix or data fram.
-
-method	
-a character string indicating which partial correlation coefficient is to be computed. One of "pearson" (default), "kendall", or "spearman" can be abbreviated.
 
@@
 
-estimate = ggm.estimate.pcor(pacor.brain_data, method = "static", lambda = 0)
+target_liver = cbind(tes.liver, tens.liver)
+
+pacor.ssec.brain = c()
+
+for (i in 1:174)
+{
+  test = data.frame(tes.brain[,i], target_liver)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_liver)
+
+  pacor.ssec.brain = c(pacor.ssec.brain, score) 
+}
+
+```
+
+```{r, echo = FALSE, eval = FALSE}
+
+# ARCHIVE BAD STUFF
+
+test = data.frame(tes.brain[,1], target_liver)
+estimate = ggm.estimate.pcor(test, method = "static") #, lambda = 0) # lambda??
 edge = network.test.edges(estimate)

### 2022-06-20 SeldinReport.Rmd
happy pride month

--- 2022-05-23 SeldinReport.Rmd
@@
 brain.liv = bicorAndPvalue(ts_protein_brain, target_liver, use = 'pairwise.complete.obs')
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(brain.liv$p, "Bicor_brain.csv")
 # brain.liv$p
@@
 # brain.liv$p
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(liv.brain$p, "Bicor_liver.csv")
 

### 2022-06-20 Tissue Exclusive Flow Chart.Rmd
happy pride month

@@
 Brain (657 genes)
-Liver (818 genes)"
+Liver (518 genes)"
 boxSe = boxGrob(txt5, x = 0.35, y = 0.70)
@@
   Protein-Coding Genes
-Brain (2,589 genes)
-Liver (371 genes)"
+Brain (2,577 genes)
+Liver (368 genes)"
 boxTENS = boxGrob(txt8, x = 0.55, y = 0.30)
@@
 Brain (2,751 genes)
-Liver (405 genes)"
+Liver (403 genes)"
 boxTE = boxGrob(txt8, x = 0.55, y = 0.5)

### 2022-07-01 MSigDB.Rmd
fixing problem with order of tes v. target

@@
 
-all_genes = msigdbr(species = "Rattus norvegicus")
+H_genes = msigdbr(species = "Rattus norvegicus", category = "H")
+C2_genes = msigdbr(species = "Rattus norvegicus", category = "C2")
+C5_genes = msigdbr(species = "Rattus norvegicus", category = "C5")
+C8_genes = msigdbr(species = "Rattus norvegicus", category = "C8")
 
@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
@@
 
+When retrieving Entrez IDs for our gene sets of interest, I found that many of the genes within these sets were no longer available within biomaRt.
+Looking at the Ensembl.org website, many of these ensembl IDs were retired via Ensembl.
+The version in which these ensembl IDs were available is version 104.
+This version was used for the enrichment analysis.
+However, there may be concerns using these retired ensembl IDs, especially where there is no longer a corresponding ID in the new version.
+Below is a table summarizing the number of genes missing, along with examples of pages from defunct ensembl ID webpages.
+
 ```{r, echo = FALSE, eval = FALSE}
@@
                  values = "protein_coding",
-                 mart = ensembl)
+                 mart = ensembl.miss)
 
@@
 
-missing_b = combine.brain[combine.brain %!in% entrez_brain$ensembl_gene_id]
+missing_b = combine.brain[combine.brain %!in% entrez_brain.1$ensembl_gene_id]
 # 15 missing TES : "ENSRNOG00000047739" "ENSRNOG00000008563" "ENSRNOG00000009140" "ENSRNOG00000038437" "ENSRNOG00000034130" "ENSRNOG00000003887" "ENSRNOG00000007374" "ENSRNOG00000013441" "ENSRNOG00000003891" "ENSRNOG00000038980" "ENSRNOG00000013228" "ENSRNOG00000038150" "ENSRNOG00000049171" "ENSRNOG00000011074" "ENSRNOG00000050172"
@@
 
-missing_l = combine.liver[combine.liver %!in% entrez_liver$ensembl_gene_id]
+missing_l = combine.liver[combine.liver %!in% entrez_liver.1$ensembl_gene_id]
 # 12 missing TES : "ENSRNOG00000020805" "ENSRNOG00000050499" "ENSRNOG00000049727" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000013464" "ENSRNOG00000049515" "ENSRNOG00000013410" "ENSRNOG00000016535" "ENSRNOG00000033619" "ENSRNOG00000033245"
@@
 ```
+
+```{r, echo = FALSE}
+
+missingness = data.frame(dataset = c("TES Brain", "Sig Target Genes Brain", "TES Liver", "Sig Target Genes Liver"), missing = c(15, 28, 12, 116), total = c(174, 138, 35, 930))
+
+knitr::kable(missingness, caption = "Missing Ensembl IDs within Datasets")
+
+```
+
+![Missing Gene: ENSRNOG00000050885](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/ENSRNOG00000050885_Ensembl.png)
+
+![Missing Gene: ENSRNOG00000013464](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/ENSRNOG00000013464_Ensembl.png)
 
@@
 
+## Enrichment Analysis
+
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 entrez_brain = read.csv("EntrezIds_brain.csv")
@@
 
-msigdbr_t2g = all_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_H = H_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_C2 = C2_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_C5 = C5_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_C8 = C8_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
 
@@
 
-# brain.list, liver.list
+# H
 
-out.brain = list(c(), c(), c(), c(), c())
+out.brain_H = list(c(), c(), c(), c(), c())
+database = msigdbr_t2g_H
 
@@
   entrez = na.omit(entrez)
-  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
-  out.brain[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_H[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
 }
 
-out.liver = list(c(), c(), c(), c(), c())
+out.liver_H = list(c(), c(), c(), c(), c())
 
@@
   entrez = na.omit(entrez)
-  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
-  out.liver[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_H[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
 }
 
-brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
-names(brain.ids) = top.brain[,1]
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
 
-knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
 
-liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
-names(liver.ids) = top.liver[,1]
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
 
-knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C2
+
+out.brain_C2 = list(c(), c(), c(), c(), c())
+database = msigdbr_t2g_C2
+
+for (i in 1:5)
+{
+  vect = brain.list[[i]]
+  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_C2[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+out.liver_C2 = list(c(), c(), c(), c(), c())
+
+for (i in 1:5)
+{
+  vect = liver.list[[i]]
+  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_C2[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C5
+
+out.brain_C5 = list(c(), c(), c(), c(), c())
+database = msigdbr_t2g_C5
+
+for (i in 1:5)
+{
+  vect = brain.list[[i]]
+  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_C5[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+out.liver_C5 = list(c(), c(), c(), c(), c())
+
+for (i in 1:5)
+{
+  vect = liver.list[[i]]
+  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_C5[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C8
+
+out.brain_C8 = list(c(), c(), c(), c(), c())
+database = msigdbr_t2g_C8
+
+for (i in 1:5)
+{
+  vect = brain.list[[i]]
+  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_C8[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+out.liver_C8 = list(c(), c(), c(), c(), c())
+
+for (i in 1:5)
+{
+  vect = liver.list[[i]]
+  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_C8[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+first.brain = rbind(out.brain_H[[1]], out.brain_C2[[1]], out.brain_C5[[1]], out.brain_C8[[1]])
+second.brain = rbind(out.brain_H[[2]], out.brain_C2[[2]], out.brain_C5[[2]], out.brain_C8[[2]])
+third.brain = rbind(out.brain_H[[3]], out.brain_C2[[3]], out.brain_C5[[3]], out.brain_C8[[3]])
+fourth.brain = rbind(out.brain_H[[4]], out.brain_C2[[4]], out.brain_C5[[4]], out.brain_C8[[4]])
+fifth.brain = rbind(out.brain_H[[5]], out.brain_C2[[5]], out.brain_C5[[5]], out.brain_C8[[5]])
+
+first.liver = rbind(out.liver_H[[1]], out.liver_C2[[1]], out.liver_C5[[1]], out.liver_C8[[1]])
+second.liver = rbind(out.liver_H[[2]], out.liver_C2[[2]], out.liver_C5[[2]], out.liver_C8[[2]])
+third.liver = rbind(out.liver_H[[3]], out.liver_C2[[3]], out.liver_C5[[3]], out.liver_C8[[3]])
+fourth.liver = rbind(out.liver_H[[4]], out.liver_C2[[4]], out.liver_C5[[4]], out.liver_C8[[4]])
+fifth.liver = rbind(out.liver_H[[5]], out.liver_C2[[5]], out.liver_C5[[5]], out.liver_C8[[5]])
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(first.brain, "Brain Enrichment_1.csv")
+#write.csv(second.brain, "Brain Enrichment_2.csv")
+#write.csv(third.brain, "Brain Enrichment_3.csv")
+#write.csv(fourth.brain, "Brain Enrichment_4.csv")
+#write.csv(fifth.brain, "Brain Enrichment_5.csv")
+
+#write.csv(first.liver, "liver Enrichment_1.csv")
+#write.csv(second.liver, "liver Enrichment_2.csv")
+#write.csv(third.liver, "liver Enrichment_3.csv")
+#write.csv(fourth.liver, "liver Enrichment_4.csv")
+#write.csv(fifth.liver, "liver Enrichment_5.csv")
 
 ```
+
+### Enrichment of Target Liver Protein-Coding Genes acted upon by TES Protein-Coding Genes from Brain
+
+```{r, echo = FALSE}
+
+knitr::kable(first.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+knitr::kable(second.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000020519")
+knitr::kable(third.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+knitr::kable(fourth.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000016567")
+knitr::kable(fifth.brain, caption = "Brain TES Protein-Coding Gene: ENSRNOG00000008176")
+
+```
+
+### Enrichment of Target Brain Protein-Coding Genes acted upon by TES Protein-Coding Genes from Liver
+
+```{r, echo = FALSE}
+
+knitr::kable(first.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000032737")
+knitr::kable(second.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000021029")
+knitr::kable(third.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000008422")
+knitr::kable(fourth.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000002964")
+knitr::kable(fifth.liver, caption = "Liver TES Protein-Coding Gene: ENSRNOG00000011973")
+
+```

### 2022-07-01 Partial Correlation Ssec Values.Rmd
fixing problem with order of tes v. target

@@
 1.  Compare theory behind partial midweight correlation and partial correlation
-2. Calculate Partial Correlation, compare to biweight midcorrelation
-3. Convert Partial Correlation to Ssec Scores
-4. Look at distribution of Partial Correlation, QQ plots
-5. Use permutation test from Seldin to check null distribution
+2.  Convert Partial Correlation to Ssec Scores
+3.  Look at distribution of Partial Correlation, QQ plots
+4.  Use permutation test from Seldin to check null distribution
 
-Lambda???
+### Theory
+
+Partial Correlation: Adjusts for confounders in correlation.
+These are calculated via the function ggm.estimate.pcor from the GeneNet package via shrinkage?
+according to methods established in Schafer and Strimmer (2005).
+These allow for calculation of correlation for datasets where there are more variables than subjects, as is common in genetics data.
+
+Biweight Midcorrelation: semi-parametric median-based correlation measure which is not susceptible to outliers.
+[I could do a sensitivity analysis to see how influential outliers could be within Ssec measure for pacor v.
+bicor?
+"Fast R Functions for Robust Correlations and Hierarchical Clustering"]
 
 ```{r, echo = FALSE, message = FALSE}
+#https://cran.r-project.org/web/packages/correlation/vignettes/types.html
+
+## PACKAGES
 library(WGCNA)
-library(ppcor)
 library(tidyverse)
 
+## DATA
 setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, eval = FALSE}
+### CALCULATION OF PACOR SSEC TES ###
+
 library(GeneNet)
 
-target_liver = cbind(tes.liver, tens.liver)
+## TES BRAIN ##
+
+target_liver = cbind(tes.liver, tens.liver) #merge instead
+target_liver = target_liver[rownames(tes.brain),]
 
@@
   test = data.frame(tes.brain[,i], target_liver)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate)
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_liver)
+
+  pacor.ssec.brain = c(pacor.ssec.brain, score)
+
+  if (i == 1)
+  {
+    corr.pacorssec.brain = this$pcor
+    pval.pacorssec.brain = this$pval
+    max.min.brain = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
+  } else {
+    corr.pacorssec.brain = cbind(corr.pacorssec.brain, this$pcor)
+    pval.pacorssec.brain = cbind(pval.pacorssec.brain, this$pval)
+    max.min.brain = cbind(max.min.brain, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
+  }
+}
+
+## TES LIVER ##
+
+target_brain = cbind(tes.brain, tens.brain)
+target_brain = target_brain[rownames(tes.liver),]
+
+pacor.ssec.liver = c()
+
+for (i in 1:35)
+{
+  test = data.frame(tes.liver[,i], target_brain)
+  estimate = ggm.estimate.pcor(test, method = "static") 
+  edge = network.test.edges(estimate)
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_brain)
+
+  pacor.ssec.liver = c(pacor.ssec.liver, score)
+
+  if (i == 1)
+  {
+    corr.pacorssec.liver = this$pcor
+    pval.pacorssec.liver = this$pval
+    max.min.liver = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
+  } else {
+    corr.pacorssec.liver = cbind(corr.pacorssec.liver, this$pcor)
+    pval.pacorssec.liver = cbind(pval.pacorssec.liver, this$pval)
+    max.min.liver = cbind(max.min.liver, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
+  }
+}
+
+pacor.ssec.brain.fin = data.frame(gene = colnames(tes.brain), ssec = pacor.ssec.brain)
+pacor.ssec.liver.fin = data.frame(gene = colnames(tes.liver), ssec = pacor.ssec.liver)
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+#write.csv(pacor.ssec.brain.fin, "TES_PaCor_Ssec_brain.csv")
+#write.csv(pacor.ssec.liver.fin, "TES_PaCor_Ssec_liver.csv")
+write.csv(corr.pacorssec.brain, "Pacor_Brain_TES.csv")
+write.csv(corr.pacorssec.liver, "Pacor_Liver_TES.csv")
+write.csv(pval.pacorssec.brain, "Pacor.pval_Brain_TES.csv")
+write.csv(pval.pacorssec.liver, "Pacor.pval_Liver_TES.csv")
+
+```
+
+```{r, echo = FALSE, eval = FALSE}
+### CALCULATION OF PACOR SSEC TENS ###
+
+## TENS BRAIN ##
+
+pacor.tens.ssec.brain = c()
+
+for (i in 1:2589)
+{
+  test = data.frame(tens.brain[,i], target_liver)
   estimate = ggm.estimate.pcor(test, method = "static")
@@
 
-  pacor.ssec.brain = c(pacor.ssec.brain, score) 
+  pacor.tens.ssec.brain = c(pacor.tens.ssec.brain, score) 
 }
+
+## TENS LIVER ##
+
+pacor.tens.ssec.liver = c()
+
+for (i in 1:368)
+{
+  test = data.frame(tens.liver[,i], target_brain)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_brain)
+
+  pacor.tens.ssec.liver = c(pacor.tens.ssec.liver, score) 
+}
+
+pacor.tens.ssec.brain.fin = data.frame(gene = colnames(tens.brain), ssec = pacor.tens.ssec.brain)
+pacor.tens.ssec.liver.fin = data.frame(gene = colnames(tens.liver), ssec = pacor.tens.ssec.liver)
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+#write.csv(pacor.tens.ssec.brain.fin, "TENS_PaCor_Ssec_brain.csv")
+write.csv(pacor.tens.ssec.liver.fin, "TENS_PaCor_Ssec_liver.csv")
+
+```
+
+```{r, echo = FALSE}
+### DATA ###
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+ts_Ssec_target_brain = read.csv("Target_TESec_Ssec_brain.csv")
+ts_Ssec_target_liver = read.csv("Target_TESec_Ssec_liver.csv")
+tsns_Ssec_target_brain = read.csv("Target_TENSec_Ssec_brain.csv")
+tsns_Ssec_target_liver = read.csv("Target_TENSec_Ssec_liver.csv")
+pacor.ssec.brain.fin = read.csv("TES_PaCor_Ssec_brain.csv")
+pacor.ssec.liver.fin = read.csv("TES_PaCor_Ssec_liver.csv")
+pacor.tens.ssec.brain.fin = read.csv("TENS_PaCor_Ssec_brain.csv")
+pacor.tens.ssec.liver.fin = read.csv("TENS_PaCor_Ssec_liver.csv")
+
+pacor.ssec.brain.fin = fix.read.csv(pacor.ssec.brain.fin)
+pacor.ssec.liver.fin = fix.read.csv(pacor.ssec.liver.fin)
+pacor.tens.ssec.brain.fin = fix.read.csv(pacor.tens.ssec.brain.fin)
+pacor.tens.ssec.liver.fin = fix.read.csv(pacor.tens.ssec.liver.fin)
+
+```
+
+### Pearson Correlation Comparison
+
+Ssec scores calculated with Pearson Correlation are fairly similar to Ssec scores calculated with biweight midcorrelation.
+Ssec scores calculated with partial correlation are significantly different from both pearson ssec scores and bicor ssec scores.
+
+```{r, echo = FALSE}
+
+# Calculating Pearson Correlation
+
+target_liver = cbind(tes.liver, tens.liver)
+target_brain = cbind(tes.brain, tens.brain)
+
+pearson.pval = function(ex, why)
+{
+  test = cor.test(ex, y = why, alternative = "two.sided", method = "pearson")
+  return(test$p.value)
+}
+
+pear.ssec.brain = c()
+
+for (i in 1:174)
+{
+
+  estimate = apply(target_liver, MARGIN = 2, FUN = pearson.pval, why = tes.brain[,i])
+
+  score = sum(-log(estimate))/length(target_liver)
+
+  pear.ssec.brain = c(pear.ssec.brain, score) 
+}
+
+pear.ssec.liver = c()
+
+for (i in 1:35)
+{
+
+  estimate = apply(target_brain, MARGIN = 2, FUN = pearson.pval, why = tes.liver[,i])
+
+  score = sum(-log(estimate))/length(target_brain)
+
+  pear.ssec.liver = c(pear.ssec.liver, score) 
+}
+
+```
+
+```{r, echo = FALSE}
+
+h1 = cbind(rep("Bicor", times = 174), ts_Ssec_target_brain$Ssec)
+h2 = cbind(rep("PaCor", times = 174), pacor.ssec.brain.fin$ssec)
+h3 = cbind(rep("Pearson", times = 174), pear.ssec.brain)
+
+compare = rbind(h1, h2, h3)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title = "Ssec Scores: Pearson v. PaCor v. Bicor in Brain", x = "Ssec", y = "Density")
+
+# liver
+
+h1 = cbind(rep("Bicor", times = 35), ts_Ssec_target_liver$Ssec)
+h2 = cbind(rep("PaCor", times = 35), pacor.ssec.liver.fin$ssec)
+h3 = cbind(rep("Pearson", times = 35), pear.ssec.liver)
+
+compare = rbind(h1, h2, h3)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Ssec Scores: Pearson v. PaCor v. Bicor in Liver", x = "Ssec", y = "Density")
+
+```
+
+### Direct Comparisons
+
+Looking at the ranked Ssec scores for those calculated with biweight midcorrelation and partial correlation, they appear to be flipped.
+The highest ssec scores for bicor are some of the lowest ssec scores for pacor (dark red, red, orange, yellow).
+The highest ssec scores for pacor are some the lowest ssec scores for bicor (light green, green, light blue, blue).
+
+![Comparison PaCor Ssec v. Bicor Ssec](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/PaCor v Bicor.1.png)
+$$ ... $$
+![Comparison PaCor Ssec v. Bicor Ssec](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/PaCor v Bicor.2.png)
+
+$$ ... $$
+![Comparison PaCor Ssec v. Bicor Ssec](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/PaCor v Bicor.3.png)
+
+```{r, echo = FALSE, eval = FALSE}
+pacor.ssec.brain.sort = pacor.ssec.brain.fin %>%
+  arrange(desc(ssec))
+
+pacor.ssec.liver.sort = pacor.ssec.liver.fin %>%
+  arrange(desc(ssec))
+
+comparison.liver = data.frame(Bicor_gene = ts_Ssec_target_liver$Gene_symbol, Bicor_Ssec = ts_Ssec_target_liver$Ssec, PaCor_gene = pacor.ssec.liver.sort$gene, Pacor_Ssec = pacor.ssec.liver.sort$ssec)
+
+comparison.brain = data.frame(Bicor_gene = ts_Ssec_target_brain$Gene_symbol, Bicor_Ssec = ts_Ssec_target_brain$Ssec, PaCor_gene = pacor.ssec.brain.sort$gene, Pacor_Ssec = pacor.ssec.brain.sort$ssec)
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+write.csv(comparison.brain, "Compare_Bicor_Pacor_SsecBrain_rank.csv")
+write.csv(comparison.liver, "Compare_Bicor_Pacor_SsecLiver_rank.csv")
+
+```
+
+### Comparing Correlation Values and P-values
+
+The following tables and graphs show very little variability in the partial correlation values, compared to biweight midcorrelation values. The p-values on the other hand, match much more closely.
+
+```{r, echo = FALSE}
+
+# compare correlation
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+corr.pacorssec.brain = read.csv("Pacor_Brain_TES.csv")
+corr.pacorssec.liver = read.csv("Pacor_Liver_TES.csv")
+pval.pacorssec.brain = read.csv("Pacor.pval_Brain_TES.csv")
+pval.pacorssec.liver = read.csv("Pacor.pval_Liver_TES.csv")
+
+corr.pacorssec.brain = fix.read.csv(corr.pacorssec.brain)
+corr.pacorssec.liver = fix.read.csv(corr.pacorssec.liver)
+pval.pacorssec.brain = fix.read.csv(pval.pacorssec.brain)
+pval.pacorssec.liver = fix.read.csv(pval.pacorssec.liver)
+
+bicor_brain = bicorAndPvalue(tes.brain, target_liver)
+bicor_liver = bicorAndPvalue(tes.liver, target_brain)
+
+#######################################################
+corr.pval = data.frame(max.correlation.value = c(0.006, 0.72), min.pvalue = c(4.8e-09, 6.2e-06))
+rownames(corr.pval) = c("Partial Correlation", "Biweight Midcorrelation")
+
+# CORRELATION
+thisone = sample(1:174, 1)
+thisone
+h1 = cbind(rep("Bicor", times = 403), bicor_brain$bicor[thisone,])
+h2 = cbind(rep("PaCor", times = 403), corr.pacorssec.brain[,thisone])
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Correlation Value Bicor v. PaCor, Random TES Protein-Coding Gene from Brain", x = "Correlation Value", y = "Density")
+
+thatone = sample(1:35, 1)
+thatone
+h1 = cbind(rep("Bicor", times = 2751), bicor_liver$bicor[thatone,])
+h2 = cbind(rep("PaCor", times = 2751), corr.pacorssec.liver[,thatone])
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Correlation Value Bicor v. PaCor, Random TES Protein-Coding Gene from Liver", x = "Correlation Value", y = "Density")
+
+# PVALUE
+
+h1 = cbind(rep("Bicor", times = 403), bicor_brain$p[thisone,])
+h2 = cbind(rep("PaCor", times = 403), pval.pacorssec.brain[,thisone])
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="P-Value Bicor v. PaCor, Random TES Protein-Coding Gene from Brain", x = "Correlation Value", y = "Density")
+
+h1 = cbind(rep("Bicor", times = 2751), bicor_liver$p[thatone,])
+h2 = cbind(rep("PaCor", times = 2751), pval.pacorssec.liver[,thatone])
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="P-Value Bicor v. PaCor, Random TES Protein-Coding Gene from Liver", x = "Correlation Value", y = "Density")
+
+```
+
+### Distribution of Ssec Scores
+
+```{r, echo = FALSE}
+
+### Compare distributions ###
+
+# TES PaCor v. TES Bicor
+#Brain
+h1 = cbind(rep("PaCor Brain", times = length(pacor.ssec.brain.fin$ssec)), pacor.ssec.brain.fin$ssec)
+h2 = cbind(rep("Bicor Brain", times = length(ts_Ssec_target_brain$Ssec)), ts_Ssec_target_brain$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="PaCor Ssec Values v. Bicor Ssec Values: Brain", x = "Ssec", y = "Density")
+
+# Liver
+h1 = cbind(rep("PaCor Liver", times = length(pacor.ssec.liver.fin$ssec)), pacor.ssec.liver.fin$ssec)
+h2 = cbind(rep("Bicor Liver", times = length(ts_Ssec_target_liver$Ssec)), ts_Ssec_target_liver$Ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="PaCor Ssec Values v. Bicor Ssec Values: Liver", x = "Ssec", y = "Density")
+
+# TES PaCor v. TENS PaCor
+#Brain
+h1 = cbind(rep("TES Brain", times = length(pacor.ssec.brain.fin$ssec)), pacor.ssec.brain.fin$ssec)
+h2 = cbind(rep("TENS Brain", times = length(pacor.tens.ssec.brain.fin$ssec)), pacor.tens.ssec.brain.fin$ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="TES v. TENS PaCor Ssec Value: Brain", x = "Ssec", y = "Density")
+
+# Liver
+h1 = cbind(rep("TES Liver", times = length(pacor.ssec.liver.fin$ssec)), pacor.ssec.liver.fin$ssec)
+h2 = cbind(rep("TENS Liver", times = length(pacor.tens.ssec.liver.fin$ssec)), pacor.tens.ssec.liver.fin$ssec)
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="TES v. TENS PaCor Ssec Value: Liver", x = "Ssec", y = "Density")
 
@@
 
-# ARCHIVE BAD STUFF
+# trying something
 
-test = data.frame(tes.brain[,1], target_liver)
-estimate = ggm.estimate.pcor(test, method = "static") #, lambda = 0) # lambda??
-edge = network.test.edges(estimate)
+quant = function(x)
+{
+  out = c()
+  cdf = ecdf(x)
+  for (i in x)
+  {
+    gee = cdf(i)
+    out = c(out, gee)
+  }
+  return(out)
+}
 
-n2kappa(30, p = 577)
+pacor.brain = quant(compare$val[compare$trt == "PaCor Brain"])
+bicor.brain = quant(compare$val[compare$trt == "Bicor Brain"])
 
-pval = cor0.test(estimate, kappa = 29)
+compare.1 = compare
+compare.1[,2] = c(pacor.brain, bicor.brain)
 
-kappa2n(29, p = 2)
-
-brain.liv = bicorAndPvalue(pacor.brain_data, pacor.brain_data, use = 'pairwise.complete.obs')
-
-cor = cor2pcor(brain.liv$bicor)
-cor2 = pcor.shrink(pacor.brain_data)
+with(compare.1, qqplot(val[trt == "Bicor Brain"], val[trt == "PaCor Brain"]))
+abline(a = 0, b = 1)
 
 ```
+
+### QQ Plots
+
+#### PACOR SSEC SCORE
+
+```{r, echo = FALSE, message = FALSE}
+
+### QQ PLOTS ###
+
+## BRAIN
+
+sec = pacor.ssec.brain.fin$ssec
+nsec = pacor.tens.ssec.brain.fin$ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
+abline(a = 0, b = 1)
+
+## LIVER
+
+sec = pacor.ssec.liver.fin$ssec
+nsec = pacor.tens.ssec.liver.fin$ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
+abline(a = 0, b = 1)
+
+```
+
+#### BICOR SSEC SCORE
+
+From QQ Plots.Rmd
+
+```{r, echo = FALSE, message = FALSE}
+
+## BRAIN
+
+sec = ts_Ssec_target_brain$Ssec
+nsec = tsns_Ssec_target_brain$Ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
+abline(a = 0, b = 1)
+
+## LIVER
+
+sec = ts_Ssec_target_liver$Ssec
+nsec = tsns_Ssec_target_liver$Ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
+abline(a = 0, b = 1)
+
+```

### 2022-07-01 SeldinReport.Rmd
fixing problem with order of tes v. target

--- 2022-05-23 SeldinReport.Rmd
@@
 
-#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 protein_liver = read.csv("Liver_WGCNA.csv")
@@
 ```{r, echo = FALSE, message = FALSE}
-# Calculate Ssec values: tissue-specific -> ns tissue-specific
+# Calculate Ssec values: tissue-specific -> s tissue-specific
 #### TS ####
@@
 target_liver = cbind(ts_protein_liver, tsns_protein_liver)
+target_liver = target_liver[rownames(ts_protein_brain),]
 
 brain.liv = bicorAndPvalue(ts_protein_brain, target_liver, use = 'pairwise.complete.obs')
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(brain.liv$p, "Bicor_brain.csv")
 # brain.liv$p
@@
 target_brain = cbind(ts_protein_brain, tsns_protein_brain)
+target_brain = target_brain[rownames(ts_protein_liver),]
 
@@
 # brain.liv$p
+
+setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(liv.brain$p, "Bicor_liver.csv")
 

### 2022-07-01 Update PaCor FU.Rmd
fixing problem with order of tes v. target

--- 2022-06-20 Partial Correlation Ssec Values.Rmd
@@
----
-title: "Ssec Scores: Partial Correlation"
-author: "Angela Yoder"
-date: "`r Sys.Date()`"
----
+I hate onedrive. I thought I fixed this problem this little shit.
 
-**The goals of this program are:**
+```{r, echo = FALSE}
 
-1. Compare theory behind partial midweight correlation and partial correlation
-2. Calculate Partial Correlation, compare to biweight midcorrelation
-3. Convert Partial Correlation to Ssec Scores
-4. Look at distribution of Partial Correlation, QQ plots
-5. Use permutation test from Seldin to check null distribution
+# Calculating Pearson Correlation
 
-Lambda???
+target_liver = cbind(tes.liver, tens.liver)
+target_brain = cbind(tes.brain, tens.brain)
 
-```{r, echo = FALSE, message = FALSE}
-library(WGCNA)
-library(ppcor)
-library(tidyverse)
-
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
-
-tes.brain = read.csv("TissueSpecific2_Protein_Brain.csv")
-tes.liver = read.csv("TissueSpecific2_Protein_Liver.csv")
-tens.brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
-tens.liver = read.csv("NSTissueSpecific_Protein_Liver.csv")
-
-fix.read.csv = function(x)
+pearson.pval = function(ex, why)
 {
-  names = x[,1]
-  gee = x[,2:dim(x)[2]]
-  rownames(gee) = names
-  return(gee)
+  test = cor.test(ex, y = why, alternative = "two.sided", method = "pearson")
+  return(test$p.value)
 }
 
-tes.brain = fix.read.csv(tes.brain)
-tes.liver = fix.read.csv(tes.liver)
-tens.brain = fix.read.csv(tens.brain)
-tens.liver = fix.read.csv(tens.liver)
+pear.ssec.brain = c()
+
+for (i in 1:174)
+{
+
+  estimate = apply(target_liver, MARGIN = 2, FUN = pearson.pval, why = tes.brain[,i])
+
+  score = sum(-log(estimate))/length(target_liver)
+
+  pear.ssec.brain = c(pear.ssec.brain, score) 
+}
+
+pear.ssec.liver = c()
+
+for (i in 1:35)
+{
+
+  estimate = apply(target_brain, MARGIN = 2, FUN = pearson.pval, why = tes.liver[,i])
+
+  score = sum(-log(estimate))/length(target_brain)
+
+  pear.ssec.liver = c(pear.ssec.liver, score) 
+}
 
@@
 ```{r, echo = FALSE}
-library(GeneNet)
 
-target_liver = cbind(tes.liver, tens.liver)
+h1 = cbind(rep("Bicor", times = 174), ts_Ssec_target_brain$Ssec)
+h2 = cbind(rep("PaCor", times = 174), pacor.ssec.brain.fin$ssec)
+h3 = cbind(rep("Pearson", times = 174), pear.ssec.brain)
 
-pacor.ssec.brain = c()
+compare = rbind(h1, h2, h3)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
 
-for (i in 1:174)
-{
-  test = data.frame(tes.brain[,i], target_liver)
-  estimate = ggm.estimate.pcor(test, method = "static")
-  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="TItle", x = "Ssec", y = "Density")
 
-  # calculate ssec
+# liver
 
-  this = edge[edge$node1 == 1,]
-  score = sum(-log(this$pval))/length(target_liver)
+h1 = cbind(rep("Bicor", times = 35), ts_Ssec_target_liver$Ssec)
+h2 = cbind(rep("PaCor", times = 35), pacor.ssec.liver.fin$ssec)
+h3 = cbind(rep("Pearson", times = 35), pear.ssec.liver)
 
-  pacor.ssec.brain = c(pacor.ssec.brain, score) 
-}
+compare = rbind(h1, h2, h3)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="TItle", x = "Ssec", y = "Density")
+
+```
+
+```{r, echo = FALSE}
+pacor.ssec.brain.sort = pacor.ssec.brain.fin %>%
+  arrange(desc(ssec))
+
+pacor.ssec.liver.sort = pacor.ssec.liver.fin %>%
+  arrange(desc(ssec))
+
+comparison.liver = data.frame(Bicor_gene = ts_Ssec_target_liver$Gene_symbol, Bicor_Ssec = ts_Ssec_target_liver$Ssec, PaCor_gene = pacor.ssec.liver.sort$gene, Pacor_Ssec = pacor.ssec.liver.sort$ssec)
+
+comparison.brain = data.frame(Bicor_gene = ts_Ssec_target_brain$Gene_symbol, Bicor_Ssec = ts_Ssec_target_brain$Ssec, PaCor_gene = pacor.ssec.brain.sort$gene, Pacor_Ssec = pacor.ssec.brain.sort$ssec)
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+write.csv(comparison.brain, "Compare_Bicor_Pacor_SsecBrain_rank.csv")
+write.csv(comparison.liver, "Compare_Bicor_Pacor_SsecLiver_rank.csv")
 
@@
 
-# ARCHIVE BAD STUFF
+# compare correlation
 
-test = data.frame(tes.brain[,1], target_liver)
-estimate = ggm.estimate.pcor(test, method = "static") #, lambda = 0) # lambda??
-edge = network.test.edges(estimate)
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-n2kappa(30, p = 577)
+corr.pacorssec.brain = read.csv("Pacor_Brain_TES.csv")
+corr.pacorssec.liver = read.csv("Pacor_Liver_TES.csv")
 
-pval = cor0.test(estimate, kappa = 29)
+corr.pacorssec.brain = fix.read.csv(corr.pacorssec.brain)
+corr.pacorssec.liver = fix.read.csv(corr.pacorssec.liver)
 
-kappa2n(29, p = 2)
+bicor_brain = bicorAndPvalue(tes.brain, target_liver)
+bicor_liver = bicorAndPvalue(tes.liver, target_brain)
 
-brain.liv = bicorAndPvalue(pacor.brain_data, pacor.brain_data, use = 'pairwise.complete.obs')
+#######################################################
+# CORRELATION
+thisone = sample(1:174, 1)
+thisone
+h1 = cbind(rep("Bicor", times = 403), bicor_brain$bicor[thisone,])
+h2 = cbind(rep("PaCor", times = 403), corr.pacorssec.brain[,thisone])
 
-cor = cor2pcor(brain.liv$bicor)
-cor2 = pcor.shrink(pacor.brain_data)
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Correlation Value Bicor v. PaCor", x = "Correlation Value", y = "Density")
+
+# PVALUE
+thisone = sample(1:174, 1)
+thisone
+h1 = cbind(rep("Bicor", times = 403), bicor_brain$p[thisone,])
+h2 = cbind(rep("PaCor", times = 403), corr.pacorssec.brain[,thisone])
+
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
+
+ggplot(compare, aes(val, color = trt)) + 
+  geom_density() +
+  labs(title="Correlation Value Bicor v. PaCor", x = "Correlation Value", y = "Density")
 

### 2022-07-12 MSigDB.Rmd
Getting there

--- 2022-06-20 MSigDB.Rmd
@@
 
-all_genes = msigdbr(species = "Rattus norvegicus")
+H_genes = msigdbr(species = "Rattus norvegicus", category = "H")
+C2_genes = msigdbr(species = "Rattus norvegicus", category = "C2") # HAS KEGG genes
+C5_genes = msigdbr(species = "Rattus norvegicus", category = "C5") #GO genes
+#C8_genes = msigdbr(species = "Rattus norvegicus", category = "C8")
 
@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-tesec_brain = read.csv("TE.Sec_Ssec_brain.csv")
-tesec_liver = read.csv("TE.Sec_Ssec_liver.csv")
-tensec_brain = read.csv("TE.NSec_Ssec_brain.csv")
-tensec_liver = read.csv("TE.NSec_Ssec_liver.csv")
+tesec_brain = read.csv("Target_TESec_Ssec_brain.csv")
+tesec_liver = read.csv("Target_TESec_Ssec_liver.csv")
+tensec_brain = read.csv("Target_TENSec_Ssec_brain.csv")
+tensec_liver = read.csv("Target_TENSec_Ssec_liver.csv")
 bicor_brain = read.csv("Bicor_brain.csv")
@@
 
-# top 5 TES genes for brain and liver
-# the genes theyre correlated to
-# convert to entrez ids
+# top Ssec genes for brain and liver FROM: SeldinReport_fix.Rmd
 
-# top 5 genes for brain and liver
-
-top.brain = tesec_brain[1:5,]
-top.liver = tesec_liver[1:5,]
+top.brain = tesec_brain[1:6,]
+top.liver = tesec_liver[1,]
 
@@
 
-list.oflists = function(mat)
+list.oflists = function(mat, num)
 {
-  out = list(c(), c(), c(), c(), c())
-  for (i in 1:5)
+  out = list(c(), c(), c(), c(), c(), c())
+  out = out[1:num]
+  if(num > 1)
+  {
+    for (i in 1:num)
     {
@@
     }
+  }
+  else
+  {
+    save = colnames(mat[,mat[1,] < 0.05])
+    out[[1]] = save
+  }
   return(out)
@@
 
-brain.list = list.oflists(top.corr.brain)
-liver.list = list.oflists(top.corr.liver)
+brain.list = list.oflists(top.corr.brain, nrow(top.brain))
+liver.list = list.oflists(top.corr.liver, nrow(top.liver))
 
-combine.brain = unique(c(brain.list[[1]], brain.list[[2]], brain.list[[3]], brain.list[[4]], brain.list[[5]]))
+combine.brain = unique(c(brain.list[[1]], brain.list[[2]], brain.list[[3]], brain.list[[4]], brain.list[[5]], brain.list[[6]]))
 
-combine.liver = unique(c(liver.list[[1]], liver.list[[2]], liver.list[[3]], liver.list[[4]], liver.list[[5]]))
+combine.liver = liver.list[[1]]
 
-num.brain = c(length(brain.list[[1]]), length(brain.list[[2]]), length(brain.list[[3]]), length(brain.list[[4]]), length(brain.list[[5]]))
+num.brain = c(length(brain.list[[1]]), length(brain.list[[2]]), length(brain.list[[3]]), length(brain.list[[4]]), length(brain.list[[5]]), length(brain.list[[6]]))
 
-num.liver = c(length(liver.list[[1]]), length(liver.list[[2]]), length(liver.list[[3]]), length(liver.list[[4]]), length(liver.list[[5]]))
+num.liver = c(length(liver.list[[1]]))
 
@@
 ## Missingness Report
+
+When retrieving Entrez IDs for our gene sets of interest, I found that many of the genes within these sets were no longer available within biomaRt.
+Looking at the Ensembl.org website, many of these ensembl IDs were retired via Ensembl.
+The version in which these ensembl IDs were available is version 104.
+This version was used for the enrichment analysis.
+However, there may be concerns using these retired ensembl IDs, especially where there is no longer a corresponding ID in the new version.
+Below is a table summarizing the number of genes missing, along with examples of pages from defunct ensembl ID webpages.
 
@@
 
-get_uniprot = getBM(attributes=c("ensembl_gene_id",
-                              "external_gene_name",
-                              "description",
-                              "chromosome_name",
-                              "start_position",
-                              "end_position",
-                              "gene_biotype",
-                            "uniprot_gn_id",
-                            "name_1006",
-                            "definition_1006","go_linkage_type","namespace_1003"),
-                 filters = 'biotype',
-                 values = "protein_coding",
-                 mart = ensembl)
-
 entrez_brain.1= get_entrez[get_entrez$ensembl_gene_id %in% combine.brain,]
@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 write.csv(entrez_brain.1, "EntrezIds_brain.csv")
@@
 
-msigdbr_t2g = all_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_H = H_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_C2 = C2_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+msigdbr_t2g_C5 = C5_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
+#msigdbr_t2g_C8 = C8_genes %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
 
@@
 
-# brain.list, liver.list
+# H
 
-out.brain = list(c(), c(), c(), c(), c())
+out.brain_H = list(c(), c(), c(), c(), c(), c())
+database = msigdbr_t2g_H
 
-for (i in 1:5)
+for (i in 1:6)
 {
@@
   entrez = na.omit(entrez)
-  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
-  out.brain[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_H[[i]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
 }
 
-out.liver = list(c(), c(), c(), c(), c())
+out.liver_H = list(c())
 
-for (i in 1:5)
-{
-  vect = liver.list[[i]]
+  vect = liver.list[[1]]
   entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
   entrez = na.omit(entrez)
-  listt = enricher(gene = entrez, TERM2GENE = msigdbr_t2g, pAdjustMethod = "fdr")
-  out.liver[[i]] = data.frame(ID = listt$ID, Description = listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_H[[1]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C2
+
+out.brain_C2 = list(c(), c(), c(), c(), c(), c())
+database = msigdbr_t2g_C2
+
+for (i in 1:6)
+{
+  vect = brain.list[[i]]
+  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_C2[[i]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
 }
 
-brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
-names(brain.ids) = top.brain[,1]
+out.liver_C2 = list(c())
 
-knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+  vect = liver.list[[1]]
+  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_C2[[1]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
 
-liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
-names(liver.ids) = top.liver[,1]
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
 
-knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C5
+
+out.brain_C5 = list(c(), c(), c(), c(), c(), c())
+database = msigdbr_t2g_C5
+
+for (i in 1:6)
+{
+  vect = brain.list[[i]]
+  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.brain_C5[[i]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+}
+
+out.liver_C5 = list(c())
+
+  vect = liver.list[[1]]
+  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+  entrez = na.omit(entrez)
+  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+  out.liver_C5[[1]] = data.frame(ID = listt$ID, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = listt$geneID)
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+# C8
+
+#out.brain_C8 = list(c(), c(), c(), c(), c())
+#database = msigdbr_t2g_C8
+
+#for (i in 1:5)
+#{
+#  vect = brain.list[[i]]
+#  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
+#  entrez = na.omit(entrez)
+#  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+#  out.brain_C8[[i]] = data.frame(ID = listt$ID, Description = #listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = #listt$geneID)
+#}
+
+#out.liver_C8 = list(c(), c(), c(), c(), c())
+
+#for (i in 1:5)
+#{
+#  vect = liver.list[[i]]
+#  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
+#  entrez = na.omit(entrez)
+#  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
+#  out.liver_C8[[i]] = data.frame(ID = listt$ID, Description = #listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = #listt$geneID)
+#}
+
+#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
+#names(brain.ids) = top.brain[,1]
+
+#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+
+#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
+#names(liver.ids) = top.liver[,1]
+
+#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+
+#first.brain = rbind(out.brain_H[[1]], out.brain_C2[[1]], out.brain_C5[[1]], out.brain_C8[[1]])
+#second.brain = rbind(out.brain_H[[2]], out.brain_C2[[2]], out.brain_C5[[2]], out.brain_C8[[2]])
+#third.brain = rbind(out.brain_H[[3]], out.brain_C2[[3]], out.brain_C5[[3]], out.brain_C8[[3]])
+#fourth.brain = rbind(out.brain_H[[4]], out.brain_C2[[4]], out.brain_C5[[4]], out.brain_C8[[4]])
+#fifth.brain = rbind(out.brain_H[[5]], out.brain_C2[[5]], out.brain_C5[[5]], out.brain_C8[[5]])
+
+#first.liver = rbind(out.liver_H[[1]], out.liver_C2[[1]], out.liver_C5[[1]], out.liver_C8[[1]])
+#second.liver = rbind(out.liver_H[[2]], out.liver_C2[[2]], out.liver_C5[[2]], out.liver_C8[[2]])
+#third.liver = rbind(out.liver_H[[3]], out.liver_C2[[3]], out.liver_C5[[3]], out.liver_C8[[3]])
+#fourth.liver = rbind(out.liver_H[[4]], out.liver_C2[[4]], out.liver_C5[[4]], out.liver_C8[[4]])
+#fifth.liver = rbind(out.liver_H[[5]], out.liver_C2[[5]], out.liver_C5[[5]], out.liver_C8[[5]])
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(first.brain, "Brain Enrichment_1.csv")
+#write.csv(second.brain, "Brain Enrichment_2.csv")
+#write.csv(third.brain, "Brain Enrichment_3.csv")
+#write.csv(fourth.brain, "Brain Enrichment_4.csv")
+#write.csv(fifth.brain, "Brain Enrichment_5.csv")
+
+#write.csv(first.liver, "liver Enrichment_1.csv")
+#write.csv(second.liver, "liver Enrichment_2.csv")
+#write.csv(third.liver, "liver Enrichment_3.csv")
+#write.csv(fourth.liver, "liver Enrichment_4.csv")
+#write.csv(fifth.liver, "liver Enrichment_5.csv")
 
 ```
+
+### Enrichment of Target Liver Protein-Coding Genes acted upon by TES Protein-Coding Genes from Brain
+
+#### HALLMARK GENE SET
+
+```{r, echo = FALSE}
+
+knitr::kable(out.brain_H[[1]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+knitr::kable(out.brain_H[[2]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+knitr::kable(out.brain_H[[3]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+knitr::kable(out.brain_H[[4]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+knitr::kable(out.brain_H[[5]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+knitr::kable(out.brain_H[[6]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+
+```
+
+#### C2 GENE SET
+
+```{r, echo = FALSE}
+# Search KEGG
+
+these_1 = grep("KEGG", out.brain_C2[[1]][,1], value = TRUE)
+these_2 = grep("KEGG", out.brain_C2[[2]][,1], value = TRUE)
+these_3 = grep("KEGG", out.brain_C2[[3]][,1], value = TRUE)
+these_4 = grep("KEGG", out.brain_C2[[4]][,1], value = TRUE)
+these_5 = grep("KEGG", out.brain_C2[[5]][,1], value = TRUE)
+these_6 = grep("KEGG", out.brain_C2[[6]][,1], value = TRUE)
+
+knitr::kable(out.brain_C2[[1]][out.brain_C2[[1]][,1] %in% these_1,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+knitr::kable(out.brain_C2[[2]][out.brain_C2[[2]][,1] %in% these_2,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+knitr::kable(out.brain_C2[[3]][out.brain_C2[[3]][,1] %in% these_3,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+knitr::kable(out.brain_C2[[4]][out.brain_C2[[4]][,1] %in% these_4,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+knitr::kable(out.brain_C2[[5]][out.brain_C2[[5]][,1] %in% these_5,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+knitr::kable(out.brain_C2[[6]][out.brain_C2[[6]][,1] %in% these_6,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+
+```
+
+#### C5 GENE SET
+
+```{r, echo = FALSE}
+
+these_1 = grep("GOBP", out.brain_C5[[1]][,1], value = TRUE)
+these_2 = grep("GOBP", out.brain_C5[[2]][,1], value = TRUE)
+these_3 = grep("GOBP", out.brain_C5[[3]][,1], value = TRUE)
+these_4 = grep("GOBP", out.brain_C5[[4]][,1], value = TRUE)
+these_5 = grep("GOBP", out.brain_C5[[5]][,1], value = TRUE)
+these_6 = grep("GOBP", out.brain_C5[[6]][,1], value = TRUE)
+
+knitr::kable(out.brain_C5[[1]][out.brain_C5[[1]][,1] %in% these_1,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+knitr::kable(out.brain_C5[[2]][out.brain_C5[[2]][,1] %in% these_2,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+knitr::kable(out.brain_C5[[3]][out.brain_C5[[3]][,1] %in% these_3,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+knitr::kable(out.brain_C5[[4]][out.brain_C5[[4]][,1] %in% these_4,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+knitr::kable(out.brain_C5[[5]][out.brain_C5[[5]][,1] %in% these_5,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+knitr::kable(out.brain_C5[[6]][out.brain_C5[[6]][,1] %in% these_6,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+
+```
+
+### Enrichment of Target Brain Protein-Coding Genes acted upon by TES Protein-Coding Genes from Liver
+
+#### HALLMARK GENE SET
+
+```{r, echo = FALSE}
+
+knitr::kable(out.liver_H[[1]], caption = "Hallmark Liver TES Protein-Coding Gene: ENSRNOG00000047481")
+
+```
+
+#### C2 GENE SET
+
+```{r, echo = FALSE}
+# search KEGG
+
+these_1 = grep("KEGG", out.liver_C2[[1]][,1], value = TRUE)
+
+knitr::kable(out.liver_C2[[1]][out.liver_C2[[1]][,1] %in% these_1,], caption = "C2: KEGG Liver TES Protein-Coding Gene: ENSRNOG00000047481")
+
+```
+
+#### C5 GENE SET
+
+```{r, echo = FALSE}
+
+these_1 = grep("GOBP", out.liver_C5[[1]][,1], value = TRUE)
+
+knitr::kable(out.liver_C5[[1]][out.liver_C5[[1]][,1] %in% these_1,], caption = "C5: GO Biological Processes Liver TES Protein-Coding Gene: ENSRNOG00000047481")
+
+```

### 2022-07-12 Partial Correlation Ssec Values.Rmd
Getting there

@@
 ## DATA
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
@@
 
-## TES BRAIN ##
-
-target_liver = cbind(tes.liver, tens.liver) #merge instead
+target_liver = cbind(tes.liver, tens.liver) 
 target_liver = target_liver[rownames(tes.brain),]
-
-pacor.ssec.brain = c()
-
-for (i in 1:174)
-{
-  test = data.frame(tes.brain[,i], target_liver)
-  estimate = ggm.estimate.pcor(test, method = "static")
-  edge = network.test.edges(estimate)
-
-  # calculate ssec
-
-  this = edge[edge$node1 == 1,]
-  score = sum(-log(this$pval))/length(target_liver)
-
-  pacor.ssec.brain = c(pacor.ssec.brain, score)
-
-  if (i == 1)
-  {
-    corr.pacorssec.brain = this$pcor
-    pval.pacorssec.brain = this$pval
-    max.min.brain = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
-  } else {
-    corr.pacorssec.brain = cbind(corr.pacorssec.brain, this$pcor)
-    pval.pacorssec.brain = cbind(pval.pacorssec.brain, this$pval)
-    max.min.brain = cbind(max.min.brain, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
-  }
-}
-
-## TES LIVER ##
 
@@
 
-pacor.ssec.liver = c()
-
-for (i in 1:35)
+partial.correlation = function(x, y)
 {
-  test = data.frame(tes.liver[,i], target_brain)
+  test = data.frame(x, y)
   estimate = ggm.estimate.pcor(test, method = "static") 
@@
 
-  # calculate ssec
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(y)
 
-  this = edge[edge$node1 == 1,]
-  score = sum(-log(this$pval))/length(target_brain)
+  return(score)
 
-  pacor.ssec.liver = c(pacor.ssec.liver, score)
+}
 
-  if (i == 1)
-  {
-    corr.pacorssec.liver = this$pcor
-    pval.pacorssec.liver = this$pval
-    max.min.liver = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
-  } else {
-    corr.pacorssec.liver = cbind(corr.pacorssec.liver, this$pcor)
-    pval.pacorssec.liver = cbind(pval.pacorssec.liver, this$pval)
-    max.min.liver = cbind(max.min.liver, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
-  }
-}
+pacor.ssec.brain = apply(tes.brain, 2, partial.correlation, y = target_liver)
+pacor.ssec.liver = apply(tes.liver, 2, partial.correlation, y = target_brain)
 
@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-#write.csv(pacor.ssec.brain.fin, "TES_PaCor_Ssec_brain.csv")
-#write.csv(pacor.ssec.liver.fin, "TES_PaCor_Ssec_liver.csv")
-write.csv(corr.pacorssec.brain, "Pacor_Brain_TES.csv")
-write.csv(corr.pacorssec.liver, "Pacor_Liver_TES.csv")
-write.csv(pval.pacorssec.brain, "Pacor.pval_Brain_TES.csv")
-write.csv(pval.pacorssec.liver, "Pacor.pval_Liver_TES.csv")
+write.csv(pacor.ssec.brain.fin, "TES_PaCor_Ssec_brain.csv")
+write.csv(pacor.ssec.liver.fin, "TES_PaCor_Ssec_liver.csv")
+#write.csv(corr.pacorssec.brain, "Pacor_Brain_TES.csv")
+#write.csv(corr.pacorssec.liver, "Pacor_Liver_TES.csv")
+#write.csv(pval.pacorssec.brain, "Pacor.pval_Brain_TES.csv")
+#write.csv(pval.pacorssec.liver, "Pacor.pval_Liver_TES.csv")
 
@@
 
-## TENS BRAIN ##
-
-pacor.tens.ssec.brain = c()
-
-for (i in 1:2589)
-{
-  test = data.frame(tens.brain[,i], target_liver)
-  estimate = ggm.estimate.pcor(test, method = "static")
-  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
-
-  # calculate ssec
-
-  this = edge[edge$node1 == 1,]
-  score = sum(-log(this$pval))/length(target_liver)
-
-  pacor.tens.ssec.brain = c(pacor.tens.ssec.brain, score) 
-}
-
-## TENS LIVER ##
-
-pacor.tens.ssec.liver = c()
-
-for (i in 1:368)
-{
-  test = data.frame(tens.liver[,i], target_brain)
-  estimate = ggm.estimate.pcor(test, method = "static")
-  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
-
-  # calculate ssec
-
-  this = edge[edge$node1 == 1,]
-  score = sum(-log(this$pval))/length(target_brain)
-
-  pacor.tens.ssec.liver = c(pacor.tens.ssec.liver, score) 
-}
+pacor.tens.ssec.brain = apply(tens.brain, 2, partial.correlation, y = target_liver)
+pacor.tens.ssec.liver = apply(tens.liver, 2, partial.correlation, y = target_brain)
 
@@
 
-setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-#write.csv(pacor.tens.ssec.brain.fin, "TENS_PaCor_Ssec_brain.csv")
+write.csv(pacor.tens.ssec.brain.fin, "TENS_PaCor_Ssec_brain.csv")
 write.csv(pacor.tens.ssec.liver.fin, "TENS_PaCor_Ssec_liver.csv")
@@
 ```
+
+```{r, echo = FALSE, eval = FALSE}
+
+target_liver = cbind(tes.liver, tens.liver) 
+target_liver = target_liver[rownames(tes.brain),]
+
+target_brain = cbind(tes.brain, tens.brain)
+target_brain = target_brain[rownames(tes.liver),]
+
+pacor.ssec.liver = c()
+
+for (i in 1:35)
+{
+  test = data.frame(tes.liver[,i], target_brain)
+  estimate = ggm.estimate.pcor(test, method = "static") 
+  edge = network.test.edges(estimate)
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_brain)
+
+  pacor.ssec.liver = c(pacor.ssec.liver, score)
+
+  if (i == 1)
+  {
+    corr.pacorssec.liver = this$pcor
+    pval.pacorssec.liver = this$pval
+    max.min.liver = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
+  } else {
+    corr.pacorssec.liver = cbind(corr.pacorssec.liver, this$pcor)
+    pval.pacorssec.liver = cbind(pval.pacorssec.liver, this$pval)
+    max.min.liver = cbind(max.min.liver, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
+  }
+}
+
+pacor.ssec.brain = c()
+
+for (i in 1:174)
+{
+  test = data.frame(tes.brain[,i], target_liver)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate)
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_liver)
+
+  pacor.ssec.brain = c(pacor.ssec.brain, score)
+
+  if (i == 1)
+  {
+    corr.pacorssec.brain = this$pcor
+    pval.pacorssec.brain = this$pval
+    max.min.brain = c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval))
+  } else {
+    corr.pacorssec.brain = cbind(corr.pacorssec.brain, this$pcor)
+    pval.pacorssec.brain = cbind(pval.pacorssec.brain, this$pval)
+    max.min.brain = cbind(max.min.brain, c(max(this$pcor), min(this$pcor), max(this$pval), min(this$pval)))
+  }
+}
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+write.csv(corr.pacorssec.brain, "Pacor_Brain_TES.csv")
+write.csv(corr.pacorssec.liver, "Pacor_Liver_TES.csv")
+write.csv(pval.pacorssec.brain, "Pacor.pval_Brain_TES.csv")
+write.csv(pval.pacorssec.liver, "Pacor.pval_Liver_TES.csv")
+
+############################################################
+
+pacor.tens.ssec.brain = c()
+
+for (i in 1:2589)
+{
+  test = data.frame(tens.brain[,i], target_liver)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_liver)
+
+  pacor.tens.ssec.brain = c(pacor.tens.ssec.brain, score) 
+}
+
+## TENS LIVER ##
+
+pacor.tens.ssec.liver = c()
+
+for (i in 1:368)
+{
+  test = data.frame(tens.liver[,i], target_brain)
+  estimate = ggm.estimate.pcor(test, method = "static")
+  edge = network.test.edges(estimate) # pvalue may be fdr... :P, but i guess it doesn't matter??
+
+  # calculate ssec
+
+  this = edge[edge$node1 == 1,]
+  score = sum(-log(this$pval))/length(target_brain)
+
+  pacor.tens.ssec.liver = c(pacor.tens.ssec.liver, score) 
+}
+
+```

### 2022-07-12 SeldinReport.Rmd
Getting there

--- 2022-06-20 SeldinReport.Rmd
@@
 
+```{r, echo = FALSE, message = FALSE, eval = FALSE}
 ## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group All Target Genes
 
-```{r, echo = FALSE, message = FALSE}
 # Calculate Ssec values: tissue-specific -> full
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, eval = FALSE}
 # Visualizations
@@
 target_liver = cbind(ts_protein_liver, tsns_protein_liver)
+target_liver = target_liver[rownames(ts_protein_brain),]
 
@@
 
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
 setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
-write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
-write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
-write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
+#write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
+#write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
+#write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
+#write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
 
@@
 
-with(compare.b, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
-abline(a = 0, b = 1)
+#with(compare.b, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
+#abline(a = 0, b = 1)
 
@@
 
-with(compare.l, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
+#with(compare.l, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
+#abline(a = 0, b = 1)
+
+```
+
+```{r, echo = FALSE}
+
+## BRAIN
+
+sec = ts_Ssec_brain$Ssec
+nsec = tsns_Ssec_brain$Ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
+abline(a = 0, b = 1)
+
+## LIVER
+
+sec = ts_Ssec_liver$Ssec
+nsec = tsns_Ssec_liver$Ssec
+
+cdf = ecdf(sec)
+
+quant = c()
+for (i in sec)
+{
+  gee = cdf(i)
+  quant = c(quant, gee)
+}
+
+vals.nsec = quantile(nsec, probs = quant)
+
+this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
+colnames(this) = c("trt", "val")
+this = as.data.frame(this)
+this$val = as.numeric(this$val)
+
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
 abline(a = 0, b = 1)
@@
 
-```{r}
+```{r, eval = FALSE, echo = FALSE}
 # checking secreted
@@
 
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 # annotation
@@
 
-```{r, echo = FALSE, warning = FALSE}
+```{r, echo = FALSE, warning = FALSE, eval = FALSE}
 
@@
 
-# still want to compare protein_brain, protein_liver- server
-
-```{r, echo = FALSE}
+```{r, echo = FALSE, eval = FALSE}
 
@@
 ```
+
 # compare secreted to non-secreted
 
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 
@@
 
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 

### 2022-07-12 SeldinReport_fix.Rmd
Getting there

--- 2022-07-12 SeldinReport.Rmd
@@
 
-1. Report on Ssec scores from "Seldin Methods"
-  a. Find distribution of Ssec values
-  b. Find distribution of correlation p-values
-  c. Find a null distribution of Ssec values
+1. Calculate Ssec values using biweight midcorrelation for TES Protein-Coding Genes
+  a. Plot the distribution of Ssec values, comparing TES to TENS
+  b. Plot the QQ plots of Ssec values, comparing TES to TENS
+  c. Find null distribution of Ssec values
+2. Select TES protein-coding genes to evaluate via enrichment analysis in MsigDB.Rmd
+  a. Plot target genes for the top TES protein-coding genes
 
-THIS WAS DONE WITH TISSUE EXCLUSIVE GENES
+This report will document the results of the fix implemented to the biweight midcorrelation calculation of Ssec scores.
+Liver and brain datasets were verified to be in the same strain order in this report.
+
+## Density Plots
+
+Below are the density plots comparing the distributions of tissue-exclusive, secreted (TES) protein-coding genes and tissue-exclusive, non-secreted (TENS) protein-coding genes for both brain and liver.
+The plots for brain are fairly similar, however there is a slight bump at the right hand side of the plot where TES protein-coding genes have a higher density.
 
@@
 ```
-
-Previously compared Secreted Genes to Tissue-Specific Secreted Genes.
-For both of these categories, they were compared to the full set of target tissue genes.
-There was not a very significant difference between the two peaks.
-Below are the graphs for Tissue-Specific Secreted Genes and Tissue-Specific Non-Secreted Genes.
-These were both compared to full set of target genes.
-These also do not show a significant difference between the Ssec scores of Tissue-Specific Secreted and Tissue-Specific Non-Secreted.
-
-```{r, echo = FALSE, message = FALSE, eval = FALSE}
-## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group All Target Genes
-
-# Calculate Ssec values: tissue-specific -> full
-#### TS ####
-## BRAIN ##
-
-brain.liv = bicorAndPvalue(ts_protein_brain, protein_liver, use = 'pairwise.complete.obs')
-# brain.liv$p
-
-scores = rowSums(-log(brain.liv$p))
-
-ts_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
-  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(protein_liver))) %>%
-  dplyr::select(-score) %>%
-  arrange(desc(Ssec))
-
-#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
-top_brain = head(ts_Ssec_brain[,1])
-bot_brain = tail(ts_Ssec_brain[,1])
-
-## LIVER ##
-
-liv.brain = bicorAndPvalue(ts_protein_liver, protein_brain, use = 'pairwise.complete.obs')
-# brain.liv$p
-
-scores = rowSums(-log(liv.brain$p))
-
-ts_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
-  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(protein_brain))) %>%
-  dplyr::select(-score) %>%
-  arrange(desc(Ssec))
-
-#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
-top_liver = head(ts_Ssec_liver[,1])
-bot_liver = tail(ts_Ssec_liver[,1])
-
-#### TSNS ####
-# ns tissue-specific -> full
-## BRAIN ##
-
-brain.liv = bicorAndPvalue(tsns_protein_brain, protein_liver, use = 'pairwise.complete.obs')
-# brain.liv$p
-
-scores = rowSums(-log(brain.liv$p))
-
-tsns_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
-  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(protein_liver))) %>%
-  dplyr::select(-score) %>%
-  arrange(desc(Ssec))
-
-#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
-#top_brain = head(tsns_Ssec_brain[,1])
-#bot_brain = tail(ts_Ssec_brain[,1])
-
-## LIVER ##
-
-liv.brain = bicorAndPvalue(tsns_protein_liver, protein_brain, use = 'pairwise.complete.obs')
-# brain.liv$p
-
-scores = rowSums(-log(liv.brain$p))
-
-tsns_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
-  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(protein_brain))) %>%
-  dplyr::select(-score) %>%
-  arrange(desc(Ssec))
-
-#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
-#top_liver = head(ts_Ssec_liver[,1])
-#bot_liver = tail(ts_Ssec_liver[,1])
-
-```
-
-```{r, echo = FALSE, eval = FALSE}
-# Visualizations
-#hist(Ssec_brain$Ssec)
-#hist(ts_Ssec_brain$Ssec)
-
-# Ssec distribtions TSNS v. TS, brain and liver
-#Brain
-h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
-h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
-
-compare = rbind(h1, h2)
-colnames(compare) = c("trt", "val")
-compare = as.data.frame(compare)
-compare$trt = as.factor(compare$trt)
-compare$val = as.numeric(compare$val)
-
-ggplot(compare, aes(val, color = trt)) + 
-  geom_density() +
-  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
-
-with(compare, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
-abline(a = 0, b = 1)
-
-# Liver
-h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
-h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
-
-compare = rbind(h1, h2)
-colnames(compare) = c("trt", "val")
-compare = as.data.frame(compare)
-compare$trt = as.factor(compare$trt)
-compare$val = as.numeric(compare$val)
-
-ggplot(compare, aes(val, color = trt)) + 
-  geom_density() +
-  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
-
-with(compare, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
-abline(a = 0, b = 1)
-
-```
-
-The graphs below modify the comparison slightly.
-Instead of comparing the tissue specific secreted and tissue specific nonsecreted to the full dataset as target genes, our target genes are ALL tissue specific genes in the target tissue (secreted and non-secreted).
-This hopefully eliminates the overlap with the non-tissue-specific genes and doesn't limit the association to just secreted genes.
-There is a very slight difference in the density graphs for these.
-The Tissue-Specific Secreted have a VERY slightly higher Ssec score on average than the Tissue-Specific Non-Secreted.
-
-## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group Tissue-Specific Target Genes
 
@@
 
-#write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
-#write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
-#write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
-#write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
+write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
+write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
+write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
+write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
 
@@
 #Brain
-h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
-h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
+h1 = cbind(rep("TENS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
+h2 = cbind(rep("TES brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
 
@@
   geom_density() +
-  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+  labs(title="Brain Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")
 
-#with(compare.b, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
-#abline(a = 0, b = 1)
+with(compare.b, qqplot(val[trt == "TENS brain"], val[trt == "TES brain"]))
+abline(a = 0, b = 1)
+
+```
+
+## Histograms
+
+```{r, echo = FALSE}
+
+c1 = rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
+c2 = rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")
+
+sec = hist(ts_Ssec_brain$Ssec, plot = FALSE) 
+nsec = hist(tsns_Ssec_brain$Ssec, plot = FALSE)
+
+plot(nsec, col = c1) 
+plot(sec, col = c2, add = TRUE)
+
+sec = hist(ts_Ssec_liver$Ssec, plot = FALSE) 
+nsec = hist(tsns_Ssec_liver$Ssec, plot = FALSE)
+
+plot(nsec, col = c1) 
+plot(sec, col = c2, add = TRUE)
+
+```
+
+The plots for liver differ more dramatically, likely due to the small sample size for TES protein-coding genes in the liver. We continue to see a small bump in the lower right that may be promising.
+
+```{r, echo = FALSE}
 
 # Liver
-h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
-h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
+h1 = cbind(rep("TENS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
+h2 = cbind(rep("TES liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
 
@@
   geom_density() +
-  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+  labs(title="Liver Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")
 
-#with(compare.l, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
-#abline(a = 0, b = 1)
+with(compare.l, qqplot(val[trt == "TENS liver"], val[trt == "TES liver"]))
+abline(a = 0, b = 1)
 
@@
 ```{r, echo = FALSE}
+
+qqplotAnnot = function(x, y,
+                       probs=c(0.001, 0.01, 0.1, 0.9, 0.99, 0.999),
+                       ...)
+{
+  qqplot(
+    x,
+    y,
+    pch=16, cex=0.5,
+    ...
+  )
+  abline(0, 1, col="red")
+
+  q_y = quantile(y, probs=probs, na.rm=TRUE)
+  q_x = quantile(x, probs=probs, na.rm=TRUE)
+
+  points(q_x, q_y, col="grey", cex=0.6)
+  text(q_x, q_y, label=paste0(probs*100, "%"), pos=1, cex=0.8, col="grey")
+}
+
+qqplotAnnot(tsns_Ssec_brain$Ssec, ts_Ssec_brain$Ssec,
+            main = "QQ-plot, Liver x adipose",
+            xlab = "Non-secreted factor Ssec (-ln p)",
+            ylab = "Secreted factor Ssec (-ln p)"
+)
+
+qqplotAnnot(tsns_Ssec_liver$Ssec, ts_Ssec_liver$Ssec,
+            main = "QQ-plot, Liver x adipose",
+            xlab = "Non-secreted factor Ssec (-ln p)",
+            ylab = "Secreted factor Ssec (-ln p)"
+)
+
+```
+
+## QQ Plots
+
+The following are the QQ plots comparing the quantiles of Ssec scores for TES protein-coding genes and TENS protein-coding genes. 
+In the plot for brain, we can see that the points follow the line fairly closely, with three points slightly above the null line.
+In the plot for liver, there is a pretty substantial dip of points below the null line, with no points above the line.
+
+```{r, echo = FALSE}
+################### QQ PLOTS #########################
 
@@
 
-```{r, echo = FALSE, message = FALSE}
-
-#brain
-
-wil.brain = coin::wilcox_test(val ~ trt, compare.b)
-
-# liver
-
-wil.liver = coin::wilcox_test(val ~ trt, compare.l)
-
-```
-
-```{r, eval = FALSE, echo = FALSE}
-# checking secreted
-
-checking = sec_annot[sec_annot$ensembl_gene_id == top_brain[4],] # extracellular matrix
-
-# ENSRNOG00000005770: "The WNT signalling modulator, Wise, is expressed in an interaction-dependent manner during hair-follicle cycling."
-
-```
-A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
-We can try comparing the Tissue-Specific Secreted to the Ssec values of all of the origin tissue proteins.
-I need to create that dataset in the server; attempts to complete in chunks have been unsuccessful.
-
-```{r, echo = FALSE}
-## Pvalue for top and bottom Ssec scores, liver and brain
-# brain
-plotting = melt(brain.liv$p, id.vars = "refseq")
-cheese =  as.character(plotting$Var1) %in% c(top_brain, bot_brain)
-limit = plotting[cheese, ]
-
-ggplot(limit, aes(value, color = Var1)) + 
-  geom_density() +
-  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Genes", x = "P-Value", y = "Density")
-
-# liver
-plotting = melt(liv.brain$p, id.vars = "refseq")
-limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]
-
-ggplot(limit, aes(value, color = Var1)) + 
-  geom_density() +
-  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Genes", x = "P-Value", y = "Density")
-
-```
-
-The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores.
-We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores.
-These graphs do not appear as well-defined as the graphs from previous reports where the all target tissue genes were used as the reference.
-[Attempted to put pvalues on log scale but lost the distinction between top and bottom Ssec scores; Tried to include a line for a null distribution, but was having trouble with ggplot]
-
-```{r, echo = FALSE}
-# Correlation coefficients top and bottom Ssec scores, liver and brain
-# brain
-magn = brain.liv$bicor
-
-plotting = melt(magn, id.vars = "refseq")
-limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])
-
-ggplot(limit, aes(value, color = Var1)) + 
-  geom_density() +
-  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Genes", x = "Correlation", y = "Density")
-
-# liver
-magn = liv.brain$bicor
-plotting = melt(magn, id.vars = "refseq")
-limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])
-
-ggplot(limit, aes(value, color = Var1)) + 
-  geom_density() +
-  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Genes", x = "Correlation", y = "Density")
-
-```
-
-Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation.
-The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients.
+## Permutation Test
 
@@
 Below that, a protein that had the lowest Ssec value is far BELOW the 95% confidence interval (0.974 to 0.996).
-What this means has yet to be determined.
 
-```{r, echo = FALSE, eval = FALSE}
-# annotation
-
-anno_brain = sec_annot[sec_annot$ensembl_gene_id %in% top_brain,]
-anno_liver = sec_annot[sec_annot$ensembl_gene_id %in% top_liver,]
-
-```
-```{r, eval = FALSE, echo = FALSE, warning = FALSE}
+```{r, echo = FALSE, warning = FALSE, eval = FALSE}
 
@@
 # repeat 1000- compare observed value to mean of permuted values
-target_protein = protein_liver
-perm_test = matrix(NA, ncol = 1000, nrow = 174)
+
+permy = function(target_protein, tes_protein_ssec, tes_protein_protein)
+{
+  perm_test = matrix(NA, ncol = 1000, nrow = nrow(tes_protein_ssec))
   perm_test = as.data.frame(perm_test)
-rownames(perm_test) = ts_Ssec_brain[,1]
+  rownames(perm_test) = tes_protein_ssec[,1]
 
@@
 
-brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')
+    brain.liv = bicorAndPvalue(tes_protein_protein, target_protein, use = 'pairwise.complete.obs')
 
@@
 
-Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
+    Ssec_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
      #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(protein_liver))) %>%
+     mutate(Ssec = score / length(colnames(target_protein))) %>%
      dplyr::select(-score)
@@
 
-perm_test[,i] = Ssec_brain_perm[,2]
+    perm_test[,i] = Ssec_perm[,2]
   }
+
+  return(perm_test)
+}
+
+perm_test.b = permy(target_liver, ts_Ssec_brain, ts_protein_brain)
+perm_test.l = permy(target_brain, ts_Ssec_liver, ts_protein_liver)
+
+conf.int = function(x)
+{
+  se = sd(x)/sqrt(length(x))
+  upper = mean(x) + (se * 1.96)
+  lower = mean(x) - (se * 1.96)
+
+  return(c(lower, upper))
+}
+
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#write.csv(perm_test, "Permutation_Seldin_Report.csv")
+
 ```
@@
 
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
-perm_test = read.csv("Permutation_Seldin_Report.csv")
-perm_test= fix.read.csv(perm_test)
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#perm_test = read.csv("Permutation_Seldin_Report.csv")
+#perm_test= fix.read.csv(perm_test)
 
-test_top_b = perm_test[rownames(perm_test) == top_brain[1],]
+test_top_b = perm_test.b[rownames(perm_test.b) == top_brain[1],]
 true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == top_brain[1],]
-hist(as.numeric(test_top_b), main = "Null Ssec Distribution for Top Gene: ENSRNOG00000021020", xlab = "Ssec Score")
+
+hist(as.numeric(test_top_b), main = "Null Ssec Distribution for Top Gene: Brain", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_top_b))[1], 2), round(conf.int(as.numeric(test_top_b))[2],2)), xlab = "Ssec Score", xlim = c(0, true$Ssec))
 abline(v = true, col = "red")
 
-test_bot_b = perm_test[rownames(perm_test) == bot_brain[1],]
+test_bot_b = perm_test.b[rownames(perm_test.b) == bot_brain[1],]
 true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == bot_brain[1],]
-hist(as.numeric(test_bot_b), main = "Null Ssec Distribution for Bottom Gene: ENSRNOG00000026419", xlab = "Ssec Score")
+hist(as.numeric(test_bot_b), main = "Null Ssec Distribution for Bottom Gene: Brain", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_bot_b))[1], 2), round(conf.int(as.numeric(test_bot_b))[2], 2)), xlab = "Ssec Score")
 abline(v = true, col = "red")
 
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
-write.csv(perm_test, "Permutation_Seldin_Report.csv")
+test_top_l = perm_test.l[rownames(perm_test.l) == top_liver[1],]
+true = ts_Ssec_liver[ts_Ssec_liver$Gene_symbol == top_liver[1],]
+hist(as.numeric(test_top_l), main = "Null Ssec Distribution for Top Gene: Liver", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_top_l))[1], 2), round(conf.int(as.numeric(test_top_l))[2], 2)), xlab = "Ssec Score")
+abline(v = true, col = "red")
+
+test_bot_l = perm_test.l[rownames(perm_test.l) == bot_liver[1],]
+true = ts_Ssec_liver[ts_Ssec_liver$Gene_symbol == bot_liver[1],]
+hist(as.numeric(test_bot_l), main = "Null Ssec Distribution for Bottom Gene: Liver", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_bot_l))[1], 2), round(conf.int(as.numeric(test_bot_l))[2]), 2), xlab = "Ssec Score")
+abline(v = true, col = "red")
 
@@
 
-```{r, echo = FALSE, eval = FALSE}
+## Selecting Relavent TES Protein-Coding Genes
 
+Because our QQ plots are less than satisfactory, we will determine our top Ssec scores by selecting all TES protein-coding genes that have a higher Ssec score than the 95th percentile of Ssec scores for TENS protein-coding genes in the respective tissues.
+This results in the following six genes for brain and one gene for liver.
+
+```{r, echo = FALSE}
+
+brain_quant = quantile(tsns_Ssec_brain$Ssec, probs = 0.95)
+liver_quant = quantile(tsns_Ssec_liver$Ssec, probs = 0.95)
+
+selected_brain = ts_Ssec_brain[ts_Ssec_brain$Ssec > brain_quant,]
+selected_liver = ts_Ssec_liver[ts_Ssec_liver$Ssec > liver_quant,]
+
+### Plots ###
+
+# target genes
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
 setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+bicor.brain = read.csv("Bicor_brain.csv")
+bicor.liver = read.csv("Bicor_liver.csv")
+bicor.brain = fix.read.csv(bicor.brain)
+bicor.liver = fix.read.csv(bicor.liver)
 
-full_brain = read.csv("Full_Brain_asso_TS.TSNS_Ssec.csv")
-full_liver = read.csv("Full_Liver_asso_TS.TSNS_Ssec.csv")
+s.brain_1 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[1], ]
+top.target.brain_1 = colnames(s.brain_1)[s.brain_1[1,] %in% sort(as.numeric(s.brain_1))[1:4]]
 
-# Visualizations
-#hist(Ssec_brain$Ssec)
-#hist(ts_Ssec_brain$Ssec)
+s.brain_2 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[2], ]
+top.target.brain_2 = colnames(s.brain_2)[s.brain_2[1,] %in% sort(as.numeric(s.brain_2))[1:4]]
 
-# Ssec distribtions TSNS v. TS, brain and liver
-#Brain
-h1 = cbind(rep("Full brain", times = length(full_brain$Ssec)), full_brain$Ssec)
-h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
+s.brain_3 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[3], ]
+top.target.brain_3 = colnames(s.brain_3)[s.brain_3[1,] %in% sort(as.numeric(s.brain_3))[1:4]]
 
-compare.b = rbind(h1, h2)
-colnames(compare.b) = c("trt", "val")
-compare.b = as.data.frame(compare.b)
-compare.b$trt = as.factor(compare.b$trt)
-compare.b$val = as.numeric(compare.b$val)
+s.brain_4 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[4], ]
+top.target.brain_4 = colnames(s.brain_4)[s.brain_4[1,] %in% sort(as.numeric(s.brain_4))[1:4]]
 
-ggplot(compare.b, aes(val, color = trt)) + 
-  geom_density() +
-  labs(title="Brain Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+s.brain_5 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[5], ]
+top.target.brain_5 = colnames(s.brain_5)[s.brain_5[1,] %in% sort(as.numeric(s.brain_5))[1:4]]
 
-# Liver
-h1 = cbind(rep("Full liver", times = length(full_liver$Ssec)), full_liver$Ssec)
-h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
+s.brain_6 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[6], ]
+top.target.brain_6 = colnames(s.brain_6)[s.brain_6[1,] %in% sort(as.numeric(s.brain_6))[1:4]]
 
-compare.l = rbind(h1, h2)
-colnames(compare.l) = c("trt", "val")
-compare.l = as.data.frame(compare.l)
-compare.l$trt = as.factor(compare.l$trt)
-compare.l$val = as.numeric(compare.l$val)
-
-ggplot(compare.l, aes(val, color = trt)) + 
-  geom_density() +
-  labs(title="Liver Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
+s.liver_1 = bicor.liver[rownames(bicor.liver) == selected_liver$Gene_symbol[1], ]
+top.target.liver_1 = colnames(s.liver_1)[s.liver_1[1,] %in% sort(as.numeric(s.liver_1))[1:4]]
 
@@
 
-# compare secreted to non-secreted
+```{r, echo = FALSE}
+## SUMMARY TABLE
 
-```{r, echo = FALSE, eval = FALSE}
+names_brain = c("Cntn4", "Tnn", "Bmp3", "Wnt3", "Lamc3", "Cbln1")
+names_liver = c("Kiss1")
 
-#### Secreted ####
+s.brain_1top = s.brain_1[s.brain_1 < 0.05]
+s.brain_2top = s.brain_2[s.brain_2 < 0.05]
+s.brain_3top = s.brain_3[s.brain_3 < 0.05]
+s.brain_4top = s.brain_4[s.brain_4 < 0.05]
+s.brain_5top = s.brain_5[s.brain_5 < 0.05]
+s.brain_6top = s.brain_6[s.brain_6 < 0.05]
+
+numtarget_brain = c(s.brain_1top, s.brain_2top, s.brain_3top, s.brain_4top, s.brain_5top, s.brain_6top)
+
+s.liver_1top = s.liver_1[s.liver_1 < 0.05]
+
+fin_brain = data.frame(names_brain, selected_brain[,2], numtarget_brain)
+
+fin_liver = data.frame(names_liver, selected_liver[,2], s.liver_1top)
+
+knitr::kable(fin_brain, caption = "TES Protein-Coding Genes in the Brain that have higher Ssec scores than the 95th quantile of TENS protein-Coding Genes from the Brain")
+
+knitr::kable(fin_liver, caption = "TES Protein-Coding Genes in the Liver that have higher Ssec scores than the 95th quantile of TENS Protein-Coding Genes fromt the Liver")
+
+```
+
+```{r, echo = FALSE}
+
+# plotting RNA expression
+
+better.plot = function (protein, selected, target, top, tissue, not.tissue)
+{
+  par(mfrow = c(2,2))
+  for (i in 1:4)
+  {
+    plot(protein[, colnames(protein) == selected], target[, colnames(target) == top[i]], xlab = paste("TES", tissue, selected), ylab = paste("Target", not.tissue, top[i]))
+  }
+}
+```
+
+To give a rough estimate of the influence these TES protein-coding genes have in the target tissue, we have plotted the expression of the top TES protein-coding gene against the expression of the four most significantly correlated target genes for each of our top genes.
+We hope to see defined slope that will indicate the influence these genes have.
+
+### Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[1], target_liver, top.target.brain_1, "brain", "liver")
+
+```
+
+### Second Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[2], target_liver, top.target.brain_2, "brain", "liver")
+
+```
+
+### Third Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[3], target_liver, top.target.brain_3, "brain", "liver")
+
+```
+
+### Fourth Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[4], target_liver, top.target.brain_4, "brain", "liver")
+
+```
+
+### Fifth Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[5], target_liver, top.target.brain_5, "brain", "liver")
+
+```
+
+### Sixth Highest Ssec Score TES Brain and its 4 most correlated liver genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_brain, selected_brain$Gene_symbol[6], target_liver, top.target.brain_6, "brain", "liver")
+
+```
+
+### Highest Ssec Score TES Liver and its 4 most correlated brain genes
+
+```{r, echo = FALSE}
+better.plot(ts_protein_liver, selected_liver$Gene_symbol[1], target_brain, top.target.liver_1, "liver", "brain")
+
+```
+
+#### Appendix: TES v. All target tissue genes
+
+```{r, echo = FALSE, message = FALSE, eval = FALSE}
+## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group All Target Genes
+
+# Calculate Ssec values: tissue-specific -> full
+#### TS ####
 ## BRAIN ##
-target_liver = cbind(ts_protein_liver, tsns_protein_liver)
+protein_liver = protein_liver[rownames(ts_protein_brain),]
 
-brain.liv = bicorAndPvalue(secreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+brain.liv = bicorAndPvalue(ts_protein_brain, protein_liver, use = 'pairwise.complete.obs')
 # brain.liv$p
@@
 
-sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+ts_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
   #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
   dplyr::select(-score) %>%
@@
 #ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
-top_brain = head(sec_Ssec_brain[,1])
-bot_brain = tail(sec_Ssec_brain[,1])
+top_brain = head(ts_Ssec_brain[,1])
+bot_brain = tail(ts_Ssec_brain[,1])
 
@@
 
-target_brain = cbind(ts_protein_brain, tsns_protein_brain)
+protein_brain = protein_brain[rownames(ts_protein_liver),]
 
-liv.brain = bicorAndPvalue(secreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
+liv.brain = bicorAndPvalue(ts_protein_liver, protein_brain, use = 'pairwise.complete.obs')
 # brain.liv$p
@@
 
-sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+ts_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
   #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(target_brain))) %>%
+  mutate(Ssec = score / length(colnames(protein_brain))) %>%
   dplyr::select(-score) %>%
@@
 #ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
-top_liver = head(sec_Ssec_liver[,1])
-bot_liver = tail(sec_Ssec_liver[,1])
+top_liver = head(ts_Ssec_liver[,1])
+bot_liver = tail(ts_Ssec_liver[,1])
 
-#### Not Secreted ####
-
+#### TSNS ####
+# ns tissue-specific -> full
 ## BRAIN ##
 
-brain.liv.c = bicorAndPvalue(notsecreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
+brain.liv = bicorAndPvalue(tsns_protein_brain, protein_liver, use = 'pairwise.complete.obs')
 # brain.liv$p
 
-scores = rowSums(-log(brain.liv.c$p))
+scores = rowSums(-log(brain.liv$p))
 
-nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
+tsns_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
   #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(target_liver))) %>%
+  mutate(Ssec = score / length(colnames(protein_liver))) %>%
   dplyr::select(-score) %>%
@@
 
-liv.brain.c = bicorAndPvalue(notsecreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
+liv.brain = bicorAndPvalue(tsns_protein_liver, protein_brain, use = 'pairwise.complete.obs')
 # brain.liv$p
 
-scores = rowSums(-log(liv.brain.c$p))
+scores = rowSums(-log(liv.brain$p))
 
-nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
+tsns_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
   #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
-  mutate(Ssec = score / length(colnames(target_brain))) %>%
+  mutate(Ssec = score / length(colnames(protein_brain))) %>%
   dplyr::select(-score) %>%
@@
 
-####### Visualizations
+```
 
+```{r, echo = FALSE, eval = FALSE}
+# Visualizations
+#hist(Ssec_brain$Ssec)
+#hist(ts_Ssec_brain$Ssec)
+
+# Ssec distribtions TSNS v. TS, brain and liver
 #Brain
-h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
-h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)
+h1 = cbind(rep("TENS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
+h2 = cbind(rep("TES brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)
 
-compare.b = rbind(h1, h2)
-colnames(compare.b) = c("trt", "val")
-compare.b = as.data.frame(compare.b)
-compare.b$trt = as.factor(compare.b$trt)
-compare.b$val = as.numeric(compare.b$val)
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
 
-ggplot(compare.b, aes(val, color = trt)) + 
+ggplot(compare, aes(val, color = trt)) + 
   geom_density() +
-  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
-
-with(compare.b, qqplot(val[trt == "Not Secreted brain"], val[trt == "Secreted brain"]))
-abline(a = 0, b = 1)
+  labs(title="Brain Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")
 
 # Liver
-h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
-h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)
+h1 = cbind(rep("TENS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
+h2 = cbind(rep("TES liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)
 
-compare.l = rbind(h1, h2)
-colnames(compare.l) = c("trt", "val")
-compare.l = as.data.frame(compare.l)
-compare.l$trt = as.factor(compare.l$trt)
-compare.l$val = as.numeric(compare.l$val)
+compare = rbind(h1, h2)
+colnames(compare) = c("trt", "val")
+compare = as.data.frame(compare)
+compare$trt = as.factor(compare$trt)
+compare$val = as.numeric(compare$val)
 
-ggplot(compare.l, aes(val, color = trt)) + 
+ggplot(compare, aes(val, color = trt)) + 
   geom_density() +
-  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")
-
-with(compare.l, qqplot(val[trt == "Not Secreted liver"], val[trt == "Secreted liver"]))
-abline(a = 0, b = 1)
-
-#brain
-
-wil.brain = coin::wilcox_test(val ~ trt, compare.b)
-wil.brain
-
-# liver
-
-wil.liver = coin::wilcox_test(val ~ trt, compare.l)
-wil.liver
-
-# :'(
+  labs(title="Liver Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")
 
@@
 
-# random brain secreted genes
-
-sample(secreted_protein_brain, size = 4, replace = FALSE)
-
-#ENSRNOG00000012655: extracellular matrix
-# ENSRNOG00000053272: endo reticulum, extracellular space, cytoplasm, perinuclear region
-
-checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000012655",]
-checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000001806",]
-checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000008519",]
-
-checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000021020",]
-
-# random liber secreted genes
-
-sample(secreted_protein_liver, size = 4, replace = FALSE)
-
-# ENSRNOG00000058444: nucleus, chromosome, secreted
-# ENSRNOG00000048642: same
-
-```
-
-```{r, echo = FALSE, eval = FALSE}
-
-# QQ plot
-
-sec = ts_Ssec_brain$Ssec
-nsec = tsns_Ssec_brain$Ssec
+sec = ts_Ssec_liver$Ssec
+nsec = tsns_Ssec_liver$Ssec
 
@@
 
-with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"]))
+with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
 abline(a = 0, b = 1)

### 2022-07-17 MSigDB.Rmd
pre-remapping

@@
 The species was identified as "Rattus norvegicus." msigdbr has access to multiple collections, including hallmark gene sets (H), and eight different supplementary gene sets (C1 to C8).
-In this analysis, we did not exclude any of the collections.
+In this analysis, we focused on the Hallmark Sets (H), KEGG Pathways (part of C2), and GO: Biological Processes (part of C5).
 
@@
 ```
-We began by identifying the top five brain and liver tissue-exclusive, secreted (TES) protein-coding genes, as identified by Ssec score.
+The top Tissue-Exclusive, Secreted (TES) protein-coding genes were determined in SeldinReport_fix.Rmd by determining which TES protein-coding genes had Ssec scores higher than the 95th percentile of Ssec scores from tissue-exclusive, non-secreted (TENS) protein-coding genes.
 We then identified the target tissue transcripts that were most significantly associated with the top TES protein-coding genes (p-value < 0.05).
 These will be the gene sets for which we will identify annotations.
-This results in 138 liver genes that are associated with TES protein-coding genes from the brain, and 935 brain genes that are associated with TES protein-coding genes from the liver.
 
@@
 
-knitr::kable(num.target.brain, caption = "Number of Target Genes in the Liver associated with TES Protein-Coding Genes from the Brain")
+#knitr::kable(num.target.brain, caption = "Number of Target Genes in the Liver associated with TES Protein-Coding Genes from the Brain")
 
-knitr::kable(num.target.liver, caption = "Number of Target Genes in the Brain associated with TES Protein-Coding Genes from the Liver")
+#knitr::kable(num.target.liver, caption = "Number of Target Genes in the Brain associated with TES Protein-Coding Genes from the Liver")
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+summ.brain = read.csv("Summary_Table_topSsec_brain.csv")
+summ.liver = read.csv("Summary_Table_topSsec_liver.csv")
+
+knitr::kable(summ.brain, caption = "Summary of Top Ssec Scores for Brain")
+knitr::kable(summ.liver, caption = "Summary of Top Ssec Score for Liver")
 
@@
 
-# 28 missing target : "ENSRNOG00000003498" "ENSRNOG00000047079" "ENSRNOG00000053527" "ENSRNOG00000057548" "ENSRNOG00000033130" "ENSRNOG00000042543" "ENSRNOG00000046529" "ENSRNOG00000049727" "ENSRNOG00000050499" "ENSRNOG00000050885" "ENSRNOG00000051227" "ENSRNOG00000052192" "ENSRNOG00000054524" "ENSRNOG00000061059" "ENSRNOG00000020805" "ENSRNOG00000003648" "ENSRNOG00000047785" "ENSRNOG00000056194" "ENSRNOG00000013410" "ENSRNOG00000049009" "ENSRNOG00000057089" "ENSRNOG00000059007" "ENSRNOG00000061217" "ENSRNOG00000003054" "ENSRNOG00000042197" "ENSRNOG00000047520" "ENSRNOG00000047821" "ENSRNOG00000048017"
+# 66 missing target :  ENSRNOG00000031343, ENSRNOG00000037928, ENSRNOG00000046284, ENSRNOG00000046386, ENSRNOG00000047103, ENSRNOG00000052003, ENSRNOG00000056194, ENSRNOG00000057089, ENSRNOG00000058753, ENSRNOG00000058824, ENSRNOG00000020628, ENSRNOG00000030575, ENSRNOG00000037495, ENSRNOG00000042197, ENSRNOG00000043176, ENSRNOG00000045991, ENSRNOG00000046572, ENSRNOG00000047308, ENSRNOG00000047560, ENSRNOG00000048017, ENSRNOG00000049009, ENSRNOG00000049299, ENSRNOG00000049315, ENSRNOG00000050243, ENSRNOG00000053899, ENSRNOG00000055193, ENSRNOG00000055375, ENSRNOG00000055620, ENSRNOG00000058460, ENSRNOG00000059007, ENSRNOG00000059199, ENSRNOG00000060451, ENSRNOG00000013464, ENSRNOG00000033245, ENSRNOG00000046696, ENSRNOG00000047079, ENSRNOG00000047535, ENSRNOG00000048070, ENSRNOG00000048182, ENSRNOG00000050066, ENSRNOG00000050118, ENSRNOG00000050981, ENSRNOG00000056285, ENSRNOG00000057734, ENSRNOG00000058352, ENSRNOG00000060328, ENSRNOG00000003648, ENSRNOG00000042083, ENSRNOG00000057548, ENSRNOG00000061431, ENSRNOG00000033130, ENSRNOG00000042543, ENSRNOG00000046529, ENSRNOG00000049727, ENSRNOG00000050499, ENSRNOG00000003498, ENSRNOG00000039931, ENSRNOG00000040195, ENSRNOG00000047843, ENSRNOG00000050232, ENSRNOG00000050885, ENSRNOG00000051227, ENSRNOG00000054524, ENSRNOG00000058034, ENSRNOG00000013410, ENSRNOG00000046011
 
@@
 
-# 116 missing target : "ENSRNOG00000038437" "ENSRNOG00000001057" "ENSRNOG00000006591" "ENSRNOG00000007705" "ENSRNOG00000008475" "ENSRNOG00000009241" "ENSRNOG00000019857" "ENSRNOG00000021533" "ENSRNOG00000028576" "ENSRNOG00000038976" "ENSRNOG00000043358" "ENSRNOG00000045720" "ENSRNOG00000045960" "ENSRNOG00000046838" "ENSRNOG00000049878" "ENSRNOG00000052459" "ENSRNOG00000053258" "ENSRNOG00000055404" "ENSRNOG00000056130" "ENSRNOG00000057009" "ENSRNOG00000057250" "ENSRNOG00000057522" "ENSRNOG00000057724" "ENSRNOG00000061051" "ENSRNOG00000009140" "ENSRNOG00000005665" "ENSRNOG00000006226" "ENSRNOG00000007219" "ENSRNOG00000008544" "ENSRNOG00000009835" "ENSRNOG00000010720" "ENSRNOG00000011640" "ENSRNOG00000014847" "ENSRNOG00000024726" "ENSRNOG00000028279" "ENSRNOG00000042321" "ENSRNOG00000042502" "ENSRNOG00000042971" "ENSRNOG00000042985" "ENSRNOG00000043382" "ENSRNOG00000045545" "ENSRNOG00000045767" "ENSRNOG00000045837" "ENSRNOG00000046212" "ENSRNOG00000046272" "ENSRNOG00000047890" "ENSRNOG00000048172" "ENSRNOG00000048322" "ENSRNOG00000048765" "ENSRNOG00000050134" "ENSRNOG00000054121" "ENSRNOG00000055892" "ENSRNOG00000056186" "ENSRNOG00000058047" "ENSRNOG00000058276" "ENSRNOG00000058766" "ENSRNOG00000058984" "ENSRNOG00000059734" "ENSRNOG00000060719" "ENSRNOG00000062048" "ENSRNOG00000002953" "ENSRNOG00000022107" "ENSRNOG00000042211" "ENSRNOG00000047780" "ENSRNOG00000061301" "ENSRNOG00000038150" "ENSRNOG00000038980" "ENSRNOG00000003395" "ENSRNOG00000006051" "ENSRNOG00000009917" "ENSRNOG00000010749" "ENSRNOG00000011856" "ENSRNOG00000012477" "ENSRNOG00000019958" "ENSRNOG00000021183" "ENSRNOG00000022919" "ENSRNOG00000027929" "ENSRNOG00000029837" "ENSRNOG00000031734" "ENSRNOG00000032206" "ENSRNOG00000033894" "ENSRNOG00000034287" "ENSRNOG00000037852" "ENSRNOG00000038184" "ENSRNOG00000039050" "ENSRNOG00000042516" "ENSRNOG00000042939" "ENSRNOG00000046844" "ENSRNOG00000047864" "ENSRNOG00000048722" "ENSRNOG00000049435" "ENSRNOG00000050378" "ENSRNOG00000056555" "ENSRNOG00000057059" "ENSRNOG00000057228" "ENSRNOG00000059539" "ENSRNOG00000060482" "ENSRNOG00000060898" "ENSRNOG00000062013" "ENSRNOG00000000060" "ENSRNOG00000005114" "ENSRNOG00000006320" "ENSRNOG00000009388" "ENSRNOG00000009635" "ENSRNOG00000020821" "ENSRNOG00000021302" "ENSRNOG00000023404" "ENSRNOG00000024678" "ENSRNOG00000046157" "ENSRNOG00000047724" "ENSRNOG00000049802" "ENSRNOG00000050511" "ENSRNOG00000057758" "ENSRNOG00000057852" "ENSRNOG00000058416" "ENSRNOG00000060181"
+# 34 missing target : "ENSRNOG00000007374" "ENSRNOG00000047739" "ENSRNOG00000000060" "ENSRNOG00000006051" "ENSRNOG00000008125" "ENSRNOG00000011369" "ENSRNOG00000012057" "ENSRNOG00000012086" "ENSRNOG00000012961" "ENSRNOG00000024149" "ENSRNOG00000025745" "ENSRNOG00000028114" "ENSRNOG00000029046" "ENSRNOG00000032206" "ENSRNOG00000038184" "ENSRNOG00000039197" "ENSRNOG00000040213" "ENSRNOG00000042602" "ENSRNOG00000042609" "ENSRNOG00000042788" "ENSRNOG00000042939" "ENSRNOG00000043120" "ENSRNOG00000047075" "ENSRNOG00000047740" "ENSRNOG00000048172" "ENSRNOG00000048358" "ENSRNOG00000049878" "ENSRNOG00000050511" "ENSRNOG00000051471" "ENSRNOG00000055889" "ENSRNOG00000056130" "ENSRNOG00000057009" "ENSRNOG00000057522" "ENSRNOG00000061301"
 
@@
 
-missingness = data.frame(dataset = c("TES Brain", "Sig Target Genes Brain", "TES Liver", "Sig Target Genes Liver"), missing = c(15, 28, 12, 116), total = c(174, 138, 35, 930))
+missingness = data.frame(dataset = c("TES Brain", "Sig Target Genes Brain", "TES Liver", "Sig Target Genes Liver"), missing = c(15, 66, 12, 34), total = c(174, 182, 35, 329))
 
 knitr::kable(missingness, caption = "Missing Ensembl IDs within Datasets")
+
+missing.tes.liver = c("ENSRNOG00000020805", "ENSRNOG00000050499", "ENSRNOG00000049727", "ENSRNOG00000033130", "ENSRNOG00000042543", "ENSRNOG00000046529", "ENSRNOG00000013464", "ENSRNOG00000049515", "ENSRNOG00000013410", "ENSRNOG00000016535", "ENSRNOG00000033619", "ENSRNOG00000033245")
 
@@
 
+Of most concern were the 12 missing Ensembl IDs from the TES protein-coding genes in the liver (34% of total missing).
+We investigated further and found that these genes are still available in Ensembl databases, under different names.
+
 ```{r, echo = FALSE, eval = FALSE}
 
-# duplicates
-adding = c()
-dup = c()
-for (i in entrez_tes.brain$ensembl_gene_id)
-{
-  if (i %in% adding)
-  {
-    dup = c(dup, i)
-  }
-  else
-  {
-    adding = c(adding, i)
-  }
+knitr::kable(missing.tes.liver, caption = "Missing TES Liver Ensembl IDs")
 
-}
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-dup.entrez = entrez_tes.brain$entrezgene_id[entrez_tes.brain$ensembl_gene_id %in% dup]
-# there are 4 NAs in the brain dataset
-# there are 3 ensembl duplicates in brain dataset
-# there are 5 ensembl duplicates in liver dataset
+miss.tes.liver = read.csv("Missing_Ensembl_TES_Liver.txt")
+
+knitr::kable(miss.tes.liver, caption = "Ensembl ID History Converter")
 
 ```
+
+Additionally, significant target genes brain were found to have a high degree of missingness, with 36% missing.
 
@@
 ```
+
+### Cursory Googling
+
+TES.brain.2.Hallmark: Hallmark_KRAS_SIGNALING_DN
+  https://www.gsea-msigdb.org/gsea/msigdb/cards/HALLMARK_KRAS_SIGNALING_DN.html
+  Genes down-regulated by KRAS activation
+  K-RAS protein: part of signalling pathway that give instructions for the cell to grow and divide or mature and take on specialized functions
+  KRAS + alcohol linked to pancreatic cancer
+
+TES.brain.5.Hallmark: HALLMARK_FATTY_ACID_METABOLISM
+  https://www.gsea-msigdb.org/gsea/msigdb/cards/HALLMARK_FATTY_ACID_METABOLISM
+  Genes encoding proteins involved in metabolism of fatty acids
+  https://pubmed.ncbi.nlm.nih.gov/87483/#:~:text=Progressive%20alteration%20of%20the%20mitochondria,production%2C%20which%20results%20in%20ketonemia.
+"Progressive alteration of the mitochondria, which occurs during chronic alcohol consumption, decreases fatty acid oxidation by interfering with citric acid cycle activity.
+This block is partially compensated for by increased ketone body production, which results in ketonemia."
+
+TES.brain.5.Hallmark: HALLMARK_INTERFERON_GAMMA_RESPONSE
+  https://www.gsea-msigdb.org/gsea/msigdb/cards/HALLMARK_INTERFERON_GAMMA_RESPONSE.html
+  Genes up-regulated in response to IFNG
+  INFG protein: secreted as a part of innate and adaptive immune systems
+  Cytokines and lipid peroxidation in alcoholics with chronic hepatitis C virus infection: https://academic.oup.com/alcalc/article/43/2/137/122851
+
+TES.brain.5.C2-KEGG: KEGG_PPAR_SIGNALING_PATHWAY
+  https://www.gsea-msigdb.org/gsea/msigdb/cards/KEGG_PPAR_SIGNALING_PATHWAY
+  Peroxisome proliferator-activated receptors- nuclear hormone receptors activated by fatty acids
+  liver pathway: https://www.genome.jp/kegg/pathway/hsa/hsa03320.html
+  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4308472/ Peroxisome proliferator-activated receptors alpha and gamma are linked with alcohol consumption in mice and withdrawal and dependence in humans
+
+TES.liver.1.Hallmark: HALLMARK_KRAS_SIGNALLING_DN
+  * see TES.brain.2.Hallmark
+
+TES.liver.1.C2-KEGG: KEGG_NEUROACTIVE_LIGAND_RECEPTOR_INTERACTION
+  pathway: https://www.genome.jp/kegg/pathway/hsa/hsa04080.html

### 2022-07-17 SeldinReport_fix.Rmd
getting there

@@
 #setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
 
-write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
-write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
-write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
-write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
+#write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
+#write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
+#write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
+#write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")
 
@@
 
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+te.sec.b_table = read.csv("eQTL_BrainTESec_added_table.csv")
+te.sec.l_table = read.csv("eQTL_LiverTESec_added_table.csv")
+te.sec.b_table = fix.read.csv(te.sec.b_table)
+te.sec.l_table = fix.read.csv(te.sec.l_table)
+
 names_brain = c("Cntn4", "Tnn", "Bmp3", "Wnt3", "Lamc3", "Cbln1")
 names_liver = c("Kiss1")
+
+eqtl.brain = te.sec.b_table[te.sec.b_table$RNAseq.Gene.Name %in% names_brain,]
+eqtl.brain = eqtl.brain[selected_brain$Gene_symbol,]
+
+eqtl.liver = te.sec.l_table[te.sec.l_table$RNAseq.Gene.Name %in% names_liver,]
 
@@
 
-numtarget_brain = c(s.brain_1top, s.brain_2top, s.brain_3top, s.brain_4top, s.brain_5top, s.brain_6top)
+numtarget_brain = c(length(s.brain_1top), length(s.brain_2top), length(s.brain_3top), length(s.brain_4top), length(s.brain_5top), length(s.brain_6top))
 
@@
 
-fin_brain = data.frame(names_brain, selected_brain[,2], numtarget_brain)
+fin_brain = data.frame(Gene = selected_brain$Gene_symbol, Gene_Symbol = names_brain, Ssec_Score = selected_brain[,2], Number_Target_Genes = numtarget_brain, eQTL_FDR = eqtl.brain$fdr, eQTL_pval = eqtl.brain$pval)
 
-fin_liver = data.frame(names_liver, selected_liver[,2], s.liver_1top)
+fin_liver = data.frame(Gene = selected_liver$Gene_symbol, Gene_Symbol = names_liver, Ssec_score = selected_liver[,2], Number_Target_Genes = length(s.liver_1top), eQTL_FDR = eqtl.liver$fdr, eQTL_pval = eqtl.liver$pval)
 
@@
 knitr::kable(fin_liver, caption = "TES Protein-Coding Genes in the Liver that have higher Ssec scores than the 95th quantile of TENS Protein-Coding Genes fromt the Liver")
+
+#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+write.csv(fin_brain, "Summary_Table_topSsec_brain.csv")
+write.csv(fin_liver, "Summary_Table_topSsec_liver.csv")
 

### 2022-07-27 MSigDB.Rmd
Will be removing Missingness Report and Remapping

@@
 date: "`r Sys.Date()`"
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-```{r, echo = FALSE}
+```{r}
 library(msigdbr)
@@
 
-```{r, echo = FALSE}
+```{r}
 
@@
 
+summ.brain = fix.read.csv(summ.brain)
+summ.liver = fix.read.csv(summ.liver)
+
 knitr::kable(summ.brain, caption = "Summary of Top Ssec Scores for Brain")
@@
 library(biomaRt)
-#ensembl_version = "https://ensembl.org"
 
 #ensembl = useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "rnorvegicus_gene_ensembl", version = 109)
-ensembl.miss = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
+ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
 ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl", version = 104)
@@
                  values = "protein_coding",
-                 mart = ensembl.miss)
+                 mart = ensembl)
 
 entrez_brain.1= get_entrez[get_entrez$ensembl_gene_id %in% combine.brain,]
+
+entrez_topbrain = get_entrez[get_entrez$ensembl_gene_id %in% top.brain$Gene_symbol,]
+entrez_topliver = get_entrez[get_entrez$ensembl_gene_id %in% top.liver$Gene_symbol,]
 
@@
 setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+write.csv(entrez_topbrain, "EntrezIds_topbrain.csv")
+write.csv(entrez_topliver, "EntrezIds_topliver.csv")
 write.csv(entrez_brain.1, "EntrezIds_brain.csv")
@@
 
-missing.tes.liver = c("ENSRNOG00000020805", "ENSRNOG00000050499", "ENSRNOG00000049727", "ENSRNOG00000033130", "ENSRNOG00000042543", "ENSRNOG00000046529", "ENSRNOG00000013464", "ENSRNOG00000049515", "ENSRNOG00000013410", "ENSRNOG00000016535", "ENSRNOG00000033619", "ENSRNOG00000033245")
+#missing.tes.liver = c("ENSRNOG00000020805", "ENSRNOG00000050499", "ENSRNOG00000049727", "ENSRNOG00000033130", "ENSRNOG00000042543", "ENSRNOG00000046529", "ENSRNOG00000013464", "ENSRNOG00000049515", "ENSRNOG00000013410", "ENSRNOG00000016535", "ENSRNOG00000033619", "ENSRNOG00000033245")
 
@@
 ![Missing Gene: ENSRNOG00000013464](C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data/ENSRNOG00000013464_Ensembl.png)
-
-Of most concern were the 12 missing Ensembl IDs from the TES protein-coding genes in the liver (34% of total missing).
-We investigated further and found that these genes are still available in Ensembl databases, under different names.
 
@@
 
-Additionally, significant target genes brain were found to have a high degree of missingness, with 36% missing.
+## Remapping Missing Ensembl IDs
+
+Due to this missingness, efforts were made to identify new Ensembl IDs for these missing genes.
+Updated Ensembl IDs for the missing IDs were retrieved via Ensembl.org's ID History Converter, whose output was subsetted by updates to Version 105.
+A decent portion of our missing Ensembl IDs had corresponding IDs in the new version, however, many were divided into multiple IDs.
+These were not able to be converted one-to-one due to uncertainty about which version of the new IDs our detected genes would map to.
+This resulted in only one target gene from the brain and one target gene from the liver having a viable one-to-one match to a Version 105 Ensembl ID.
+
+```{r}
+
+## we got this from http://uswest.ensembl.org/Rattus_norvegicus/Tools/IDMapper?db=core;expand_form=true;tl=veVyFBLEW339Hw8k-8445349
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+
+remap.tesbrain = read.csv("remapping_TESbrain.csv")
+remap.tesliver = read.csv("remapping_TESliver.csv")
+remap.targetbrain = read.csv("remapping_targetbrain.csv")
+remap.targetliver = read.csv("remapping_targetliver.csv")
+
+remap.tesbrain$Releases = sub("\\:.*", "", remap.tesbrain$Releases)
+remap.targetbrain$Releases = sub("\\:.*", "", remap.targetbrain$Releases)
+remap.tesliver$Releases = sub("\\:.*", "", remap.tesliver$Releases)
+remap.targetliver$Releases = sub("\\:.*", "", remap.targetliver$Releases)
+
+remap.tesbrain.105 = remap.tesbrain[remap.tesbrain$Releases == 105, ]
+remap.targetbrain.105 = remap.targetbrain[remap.targetbrain$Releases == 105, ]
+remap.tesliver.105 = remap.tesliver[remap.tesliver$Releases == 105, ]
+remap.targetliver.105 = remap.targetliver[remap.targetliver$Releases == 105, ]
+
+```
+
+```{r}
+
+# summary
+
+summ.remap = data.frame(dataset = missingness$dataset, total.genes = missingness$total, missing.genes = missingness$missing, num.remapped = c(nrow(remap.tesbrain.105), nrow(remap.targetbrain.105), nrow(remap.tesliver.105), nrow(remap.targetliver.105)), num.unique.remapped = c(length(unique(remap.tesbrain.105$Requested.ID)), length(unique(remap.targetbrain.105$Requested.ID)), length(unique(remap.tesliver.105$Requested.ID)), length(unique(remap.targetliver.105$Requested.ID))))
+
+#knitr::kable(summ.remap)
+
+```
+
+```{r}
+
+# usable remapping
+
+only.one = function(x)
+{
+  out = c()
+  for (i in 1:length(x))
+  {
+    if (x[i] %in% x[-i])
+    {
+      out = out
+    }
+    else
+    {
+      out = c(out, x[i])
+    }
+  }
+  return(out)
+}
+
+usable.tesbrain = remap.tesbrain.105[remap.tesbrain.105$Requested.ID %in% only.one(remap.tesbrain.105$Requested.ID),]
+usable.targetbrain = remap.targetbrain.105[remap.targetbrain.105$Requested.ID %in% only.one(remap.targetbrain.105$Requested.ID),]
+usable.tesliver = remap.tesliver.105[remap.tesliver.105$Requested.ID %in% only.one(remap.tesliver.105$Requested.ID), ]
+usable.targetliver = remap.targetliver.105[remap.targetliver.105$Requested.ID %in% only.one(remap.targetliver.105$Requested.ID),]
+
+one.to.one.match = c(length(usable.tesbrain$Requested.ID), length(usable.targetbrain$Requested.ID), length(usable.tesliver$Requested.ID), length(usable.targetliver$Requested.ID))
+
+summ.remap = data.frame(summ.remap, one.to.one.match)
+
+knitr::kable(summ.remap)
+
+```
+
+```{r, echo = FALSE, eval = FALSE}
+
+# getting entrez IDs for these Ensembl IDs
+# replace ensembl in combine brain and then rerun
+
+replace.id = function(final, replace)
+{
+  for(i in 1:length(replace$Requested.ID))
+  {
+    final[final == replace$Requested.ID[i]] = replace$Matched.ID.s[i]
+  }
+  return(final)
+}
+
+combine.brain.remap = replace.id(combine.brain, usable.targetbrain)
+combine.liver.remap = replace.id(combine.liver, usable.targetliver)
+
+# run biomart to get entrez IDS
+
+library(biomaRt)
+
+ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
+
+get_entrez = getBM(attributes=c("ensembl_gene_id",
+                              "external_gene_name",
+                              "description",
+                              "chromosome_name",
+                              "start_position",
+                              "end_position",
+                              "gene_biotype",
+                            #"entrezgene_trans_name",
+                            "entrezgene_id"),
+                 filters = 'biotype',
+                 values = "protein_coding",
+                 mart = ensembl)
+
+entrez_brain.1= get_entrez[get_entrez$ensembl_gene_id %in% combine.brain,]
+
+entrez_liver.1 = get_entrez[get_entrez$ensembl_gene_id %in% combine.liver,]
+
+# compare to other dataframe
+
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+entrez_brain = read.csv("EntrezIds_brain.csv")
+entrez_liver = read.csv("EntrezIds_liver.csv")
+
+entrez_brain = fix.read.csv(entrez_brain)
+entrez_liver = fix.read.csv(entrez_liver)
+
+```
 
@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+Summary of enrichment analysis was complicated by the R function enricher not including fold enrichment in its output.
+I have included a summary table of the number of enriched terms for each top TES protein-coding gene in this analysis.
+At the end of the document, there are also plots provided in the r package enrichplot, though they would only correspond to a single TES protein-coding gene's analysis within a single gene set.
+I am unsure how useful they would be in describing these relationships.
+
+```{r, message = FALSE, warning = FALSE}
 
@@
 
-#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
-#names(brain.ids) = top.brain[,1]
+```
 
-#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+```{r}
+# looking for correlations to TES also
 
-#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
-#names(liver.ids) = top.liver[,1]
+setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
+entrez_topbrain = read.csv("EntrezIds_topbrain.csv")
+entrez_topliver = read.csv("EntrezIds_topliver.csv")
 
-#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+database = msigdbr_t2g_H
+two.H = enricher(gene = c(entrez_topbrain$entrezgene_id[2], 315323, 171154), TERM2GENE = database, pAdjustMethod = "fdr")
+#two.H$geneID
 
-# C8
+five.H1 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[5], 50549, 84029, 25598, 29171)), TERM2GENE = database, pAdjustMethod = "fdr")
+#five.H1$geneID
 
-#out.brain_C8 = list(c(), c(), c(), c(), c())
-#database = msigdbr_t2g_C8
+five.H2 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[5], 266708, 29187, 309178, 171371)), TERM2GENE = database, pAdjustMethod = "fdr")
+#five.H2$geneID
 
-#for (i in 1:5)
-#{
-#  vect = brain.list[[i]]
-#  entrez = entrez_brain$entrezgene_id[entrez_brain$ensembl_gene_id %in% vect]
-#  entrez = na.omit(entrez)
-#  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
-#  out.brain_C8[[i]] = data.frame(ID = listt$ID, Description = #listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = #listt$geneID)
-#}
+database = msigdbr_t2g_C2
+five.c2 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[5], 50549, 25598, 29171)), TERM2GENE = database, pAdjustMethod = "fdr")
+#five.c2$geneID
 
-#out.liver_C8 = list(c(), c(), c(), c(), c())
+six.c2 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[6], 291283, 307092, 289533, 63867)), TERM2GENE = database, pAdjustMethod = "fdr")
+#six.c2$geneID
 
-#for (i in 1:5)
-#{
-#  vect = liver.list[[i]]
-#  entrez = entrez_liver$entrezgene_id[entrez_liver$ensembl_gene_id %in% vect]
-#  entrez = na.omit(entrez)
-#  listt = enricher(gene = entrez, TERM2GENE = database, pAdjustMethod = "fdr")
-#  out.liver_C8[[i]] = data.frame(ID = listt$ID, Description = #listt$Description, pvalue = listt$pvalue, p.adjust = listt$p.adjust, geneID = #listt$geneID)
-#}
+database = msigdbr_t2g_C5
+one.c5 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[1], 296366, 312102, 301229, 289057)), TERM2GENE = database, pAdjustMethod = "fdr")
+#one.c5$geneID
 
-#brain.ids = c(paste(out.brain[[1]]$ID, collapse = ", "), paste(out.brain[[2]]$ID, collapse = ", "), paste(out.brain[[3]]$ID, collapse = ", "), paste(out.brain[[4]]$ID, collapse = ", "), paste(out.brain[[5]]$ID, collapse = ", "))
-#names(brain.ids) = top.brain[,1]
+two.c5 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[2], 29300, 117033, 25098)), TERM2GENE = database, pAdjustMethod = "fdr")
+#two.c5$geneID
+# "311862/117033"
 
-#knitr::kable(brain.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Brain", colnames = "")
+three.c5 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[3], 24833, 408236, 81806, 171154)), TERM2GENE = database, pAdjustMethod = "fdr")
+#three.c5$geneID
 
-#liver.ids = c(paste(out.liver[[1]]$ID, collapse = ", "), paste(out.liver[[2]]$ID, collapse = ", "), paste(out.liver[[3]]$ID, collapse = ", "), paste(out.liver[[4]]$ID, collapse = ", "), paste(out.liver[[5]]$ID, collapse = ", "))
-#names(liver.ids) = top.liver[,1]
+six.c5 = enricher(gene = c(c(entrez_topbrain$entrezgene_id[6], 291283, 311387, 307092)), TERM2GENE = database, pAdjustMethod = "fdr")
+#six.c5$geneID
 
-#knitr::kable(liver.ids, caption = "Enrichment of Target Liver Genes Influenced by TES Protein-Coding Genes from the Liver", colnames = "")
+database = msigdbr_t2g_H
+liver.H = enricher(gene = c(c(entrez_topliver$entrezgene_id, 24294, 689619, 79208, 29471, 24898, 25412, 25546, 25041, 29361)), TERM2GENE = database, pAdjustMethod = "fdr")
+#liver.H$geneID
 
-#first.brain = rbind(out.brain_H[[1]], out.brain_C2[[1]], out.brain_C5[[1]], out.brain_C8[[1]])
-#second.brain = rbind(out.brain_H[[2]], out.brain_C2[[2]], out.brain_C5[[2]], out.brain_C8[[2]])
-#third.brain = rbind(out.brain_H[[3]], out.brain_C2[[3]], out.brain_C5[[3]], out.brain_C8[[3]])
-#fourth.brain = rbind(out.brain_H[[4]], out.brain_C2[[4]], out.brain_C5[[4]], out.brain_C8[[4]])
-#fifth.brain = rbind(out.brain_H[[5]], out.brain_C2[[5]], out.brain_C5[[5]], out.brain_C8[[5]])
+################# LOOK INTO THIS ONE #####################
+database = msigdbr_t2g_C2
+liver.c2 = enricher(gene = c(c(entrez_topliver$entrezgene_id, 65187, 24409, 289606, 25153, 24182, 24410, 140675, 25033, 29595, 116506, 24316, 81721, 24922, 29471, 25674, 24418, 29234, 170796, 25041)), TERM2GENE = database, pAdjustMethod = "fdr")
+#liver.c2$geneID
 
-#first.liver = rbind(out.liver_H[[1]], out.liver_C2[[1]], out.liver_C5[[1]], out.liver_C8[[1]])
-#second.liver = rbind(out.liver_H[[2]], out.liver_C2[[2]], out.liver_C5[[2]], out.liver_C8[[2]])
-#third.liver = rbind(out.liver_H[[3]], out.liver_C2[[3]], out.liver_C5[[3]], out.liver_C8[[3]])
-#fourth.liver = rbind(out.liver_H[[4]], out.liver_C2[[4]], out.liver_C5[[4]], out.liver_C8[[4]])
-#fifth.liver = rbind(out.liver_H[[5]], out.liver_C2[[5]], out.liver_C5[[5]], out.liver_C8[[5]])
-
-#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
-#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
-#write.csv(first.brain, "Brain Enrichment_1.csv")
-#write.csv(second.brain, "Brain Enrichment_2.csv")
-#write.csv(third.brain, "Brain Enrichment_3.csv")
-#write.csv(fourth.brain, "Brain Enrichment_4.csv")
-#write.csv(fifth.brain, "Brain Enrichment_5.csv")
-
-#write.csv(first.liver, "liver Enrichment_1.csv")
-#write.csv(second.liver, "liver Enrichment_2.csv")
-#write.csv(third.liver, "liver Enrichment_3.csv")
-#write.csv(fourth.liver, "liver Enrichment_4.csv")
-#write.csv(fifth.liver, "liver Enrichment_5.csv")
+database = msigdbr_t2g_C5
+liver.c5 = enricher(gene = c(c(entrez_topliver$entrezgene_id, 65187, 24409, 689043, 289606, 114032, 60337, 50599, 170899, 24410, 140675, 117018, 307206, 24316, 81721, 24922, 29161, 294388, 29701, 25674, 170739, 24602, 78956, 24418, 307540, 25041)), TERM2GENE = database, pAdjustMethod = "fdr")
+#liver.c5$geneID
 
@@
 
+```{r}
+
+hallmark.brain = c(nrow(out.brain_H[[1]]), nrow(out.brain_H[[2]]), nrow(out.brain_H[[3]]), nrow(out.brain_H[[4]]), nrow(out.brain_H[[5]]), nrow(out.brain_H[[6]]))
+
+these_1k = grep("KEGG", out.brain_C2[[1]][,1], value = TRUE)
+these_2k = grep("KEGG", out.brain_C2[[2]][,1], value = TRUE)
+these_3k = grep("KEGG", out.brain_C2[[3]][,1], value = TRUE)
+these_4k = grep("KEGG", out.brain_C2[[4]][,1], value = TRUE)
+these_5k = grep("KEGG", out.brain_C2[[5]][,1], value = TRUE)
+these_6k = grep("KEGG", out.brain_C2[[6]][,1], value = TRUE)
+
+kegg.brain = c(nrow(out.brain_C2[[1]][out.brain_C2[[1]][,1] %in% these_1k,]), 
+               nrow(out.brain_C2[[2]][out.brain_C2[[2]][,1] %in% these_2k,]), 
+               nrow(out.brain_C2[[3]][out.brain_C2[[3]][,1] %in% these_3k,]), 
+               nrow(out.brain_C2[[4]][out.brain_C2[[4]][,1] %in% these_4k,]), 
+               nrow(out.brain_C2[[5]][out.brain_C2[[5]][,1] %in% these_5k,]), 
+               nrow(out.brain_C2[[6]][out.brain_C2[[6]][,1] %in% these_6k,]))
+
+these_1g = grep("GOBP", out.brain_C5[[1]][,1], value = TRUE)
+these_2g = grep("GOBP", out.brain_C5[[2]][,1], value = TRUE)
+these_3g = grep("GOBP", out.brain_C5[[3]][,1], value = TRUE)
+these_4g = grep("GOBP", out.brain_C5[[4]][,1], value = TRUE)
+these_5g = grep("GOBP", out.brain_C5[[5]][,1], value = TRUE)
+these_6g = grep("GOBP", out.brain_C5[[6]][,1], value = TRUE)
+
+go.brain = c(nrow(out.brain_C5[[1]][out.brain_C5[[1]][,1] %in% these_1g,]), 
+               nrow(out.brain_C5[[2]][out.brain_C5[[2]][,1] %in% these_2g,]), 
+               nrow(out.brain_C5[[3]][out.brain_C5[[3]][,1] %in% these_3g,]), 
+               nrow(out.brain_C5[[4]][out.brain_C5[[4]][,1] %in% these_4g,]), 
+               nrow(out.brain_C5[[5]][out.brain_C5[[5]][,1] %in% these_5g,]), 
+               nrow(out.brain_C5[[6]][out.brain_C5[[6]][,1] %in% these_6g,]))
+
+summary.brain = data.frame(gene.symbol = summ.brain$Gene_Symbol, hallmark.terms = hallmark.brain, kegg.terms = kegg.brain, go.bio.process.terms = go.brain, sig.target.genes = summ.brain$Number_Target_Genes)
+
+knitr::kable(summary.brain, caption = "Number of Significant Terms for Each Top Brain TES Protein-Coding Genes")
+
+```
+
 #### HALLMARK GENE SET
 
-```{r, echo = FALSE}
+```{r}
 
-knitr::kable(out.brain_H[[1]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000005652")
-knitr::kable(out.brain_H[[2]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002548")
-knitr::kable(out.brain_H[[3]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002381")
-knitr::kable(out.brain_H[[4]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000003845")
-knitr::kable(out.brain_H[[5]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000059507")
-knitr::kable(out.brain_H[[6]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+#knitr::kable(out.brain_H[[1]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+knitr::kable(out.brain_H[[2]], caption = "2. Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+#knitr::kable(out.brain_H[[3]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+#knitr::kable(out.brain_H[[4]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+knitr::kable(out.brain_H[[5]], caption = "5. Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+#knitr::kable(out.brain_H[[6]], caption = "Hallmark Brain TES Protein-Coding Gene: ENSRNOG00000000010")
 
@@
 
-```{r, echo = FALSE}
+```{r}
 # Search KEGG
 
-these_1 = grep("KEGG", out.brain_C2[[1]][,1], value = TRUE)
-these_2 = grep("KEGG", out.brain_C2[[2]][,1], value = TRUE)
-these_3 = grep("KEGG", out.brain_C2[[3]][,1], value = TRUE)
-these_4 = grep("KEGG", out.brain_C2[[4]][,1], value = TRUE)
-these_5 = grep("KEGG", out.brain_C2[[5]][,1], value = TRUE)
-these_6 = grep("KEGG", out.brain_C2[[6]][,1], value = TRUE)
-
-knitr::kable(out.brain_C2[[1]][out.brain_C2[[1]][,1] %in% these_1,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000005652")
-knitr::kable(out.brain_C2[[2]][out.brain_C2[[2]][,1] %in% these_2,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002548")
-knitr::kable(out.brain_C2[[3]][out.brain_C2[[3]][,1] %in% these_3,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002381")
-knitr::kable(out.brain_C2[[4]][out.brain_C2[[4]][,1] %in% these_4,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000003845")
-knitr::kable(out.brain_C2[[5]][out.brain_C2[[5]][,1] %in% these_5,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000059507")
-knitr::kable(out.brain_C2[[6]][out.brain_C2[[6]][,1] %in% these_6,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+#knitr::kable(out.brain_C2[[1]][out.brain_C2[[1]][,1] %in% these_1k,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+#knitr::kable(out.brain_C2[[2]][out.brain_C2[[2]][,1] %in% these_2k,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+#knitr::kable(out.brain_C2[[3]][out.brain_C2[[3]][,1] %in% these_3k,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+#knitr::kable(out.brain_C2[[4]][out.brain_C2[[4]][,1] %in% these_4k,], caption = "C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+knitr::kable(out.brain_C2[[5]][out.brain_C2[[5]][,1] %in% these_5k,], caption = "5. C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+knitr::kable(out.brain_C2[[6]][out.brain_C2[[6]][,1] %in% these_6k,], caption = "6. C2: KEGG Brain TES Protein-Coding Gene: ENSRNOG00000000010")
 
@@
 
-```{r, echo = FALSE}
+```{r}
 
-these_1 = grep("GOBP", out.brain_C5[[1]][,1], value = TRUE)
-these_2 = grep("GOBP", out.brain_C5[[2]][,1], value = TRUE)
-these_3 = grep("GOBP", out.brain_C5[[3]][,1], value = TRUE)
-these_4 = grep("GOBP", out.brain_C5[[4]][,1], value = TRUE)
-these_5 = grep("GOBP", out.brain_C5[[5]][,1], value = TRUE)
-these_6 = grep("GOBP", out.brain_C5[[6]][,1], value = TRUE)
-
-knitr::kable(out.brain_C5[[1]][out.brain_C5[[1]][,1] %in% these_1,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000005652")
-knitr::kable(out.brain_C5[[2]][out.brain_C5[[2]][,1] %in% these_2,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002548")
-knitr::kable(out.brain_C5[[3]][out.brain_C5[[3]][,1] %in% these_3,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002381")
-knitr::kable(out.brain_C5[[4]][out.brain_C5[[4]][,1] %in% these_4,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000003845")
-knitr::kable(out.brain_C5[[5]][out.brain_C5[[5]][,1] %in% these_5,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000059507")
-knitr::kable(out.brain_C5[[6]][out.brain_C5[[6]][,1] %in% these_6,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000000010")
+knitr::kable(out.brain_C5[[1]][out.brain_C5[[1]][,1] %in% these_1g,], caption = "1. C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000005652")
+knitr::kable(out.brain_C5[[2]][out.brain_C5[[2]][,1] %in% these_2g,], caption = "2. C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002548")
+knitr::kable(out.brain_C5[[3]][out.brain_C5[[3]][,1] %in% these_3g,], caption = "3. C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000002381")
+#knitr::kable(out.brain_C5[[4]][out.brain_C5[[4]][,1] %in% these_4g,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000003845")
+#knitr::kable(out.brain_C5[[5]][out.brain_C5[[5]][,1] %in% these_5g,], caption = "C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000059507")
+knitr::kable(out.brain_C5[[6]][out.brain_C5[[6]][,1] %in% these_6g,], caption = "6. C5: GO Biological Process Brain TES Protein-Coding Gene: ENSRNOG00000000010")
 
@@
 
+```{r}
+
+hallmark.liver = nrow(out.liver_H[[1]])
+
+these_1k = grep("KEGG", out.liver_C2[[1]][,1], value = TRUE)
+kegg.liver = nrow(out.liver_C2[[1]][out.liver_C2[[1]][,1] %in% these_1k,])
+
+these_1g = grep("GOBP", out.liver_C5[[1]][,1], value = TRUE)
+go.liver = nrow(out.liver_C5[[1]][out.liver_C5[[1]][,1] %in% these_1g,])
+
+summary.liver = data.frame(gene.symbol = summ.liver$Gene_Symbol, hallmark.terms = hallmark.liver, kegg.terms = kegg.liver, go.bio.process.terms = go.liver, sig.target.genes = summ.liver$Number_Target_Genes)
+
+knitr::kable(summary.liver, caption = "Number of Significant Terms for Each Top Liver TES Protein-Coding Genes")
+
+```
+
 #### HALLMARK GENE SET
 
-```{r, echo = FALSE}
+```{r}
 
@@
 
-```{r, echo = FALSE}
+```{r}
 # search KEGG
 
-these_1 = grep("KEGG", out.liver_C2[[1]][,1], value = TRUE)
-
-knitr::kable(out.liver_C2[[1]][out.liver_C2[[1]][,1] %in% these_1,], caption = "C2: KEGG Liver TES Protein-Coding Gene: ENSRNOG00000047481")
+knitr::kable(out.liver_C2[[1]][out.liver_C2[[1]][,1] %in% these_1k,], caption = "C2: KEGG Liver TES Protein-Coding Gene: ENSRNOG00000047481")
 
@@
 
-```{r, echo = FALSE}
+```{r}
 
-these_1 = grep("GOBP", out.liver_C5[[1]][,1], value = TRUE)
+knitr::kable(out.liver_C5[[1]][out.liver_C5[[1]][,1] %in% these_1g,], caption = "C5: GO Biological Processes Liver TES Protein-Coding Gene: ENSRNOG00000047481")
 
-knitr::kable(out.liver_C5[[1]][out.liver_C5[[1]][,1] %in% these_1,], caption = "C5: GO Biological Processes Liver TES Protein-Coding Gene: ENSRNOG00000047481")
+```
+
+### Other Plots
+
+```{r}
+
+library(enrichplot)
+barplot(listt)
+cnetplot(listt) 
+dotplot(listt)
+#goplot(listt) #DAG
 
